---
title: "Day 07 - multi-scale"
format: html
---

```{r}
ggplot2::theme_set(ggplot2::theme_minimal())
```


## Idea

show distribution per region --> scale ~ variation, scale ~ region

work with a subset of the KPIs (default to traffic and punctuality)

## distribution ~ ridgeline

### data preparation

```{r}
# load data
punc <- fs::dir_ls("data", regexp = "-punc-2024") |> 
  purrr::map(.f = ~ readr::read_csv(.x, show_col_types = FALSE)) |> 
  dplyr::bind_rows()

# label regions
punc <- punc |> mutate(REGION = case_when(
   grepl("^S", APT)   ~ "BRA"
  ,grepl("^E|L", APT) ~ "EUR"
  ,.default = NA
))


# helper functions to aggregate the data
add_dly_early_late_groupings <- function(.punc_df){
  df <- .punc_df |>  
    dplyr::mutate(
      EARLY        = rowSums(across(.cols = `(-INF,-60]`:`(-20,-15]`), na.rm = TRUE)
      ,EARLY_M15M05 = (`(-15,-10]` + `(-10,-5]`)
      ,EARLY_M05M00 =  `(-5,0]`                 
      ,LATE_P00P05  =  `(0,5)`                   
      ,LATE_P05P15  = (`[5,10)` + `[10,15)`)     
      ,LATE         = rowSums(across(.cols = `[15,20)`:`[60,INF)`), na.rm = TRUE) 
      ,WITHIN_M05P05= (`(-5,0]` + `(0,5)`)
      ,WITHIN_M15P15= (`(-15,-10]`+`(-10,-5]`+`(-5,0]`+`(0,5)`+`[5,10)`+`[10,15)`)
    )
}

daily_punc <- punc |> 
  add_dly_early_late_groupings() |> 
  dplyr::select(APT, DATE, PHASE, REGION, EARLY:LATE) |> 
  dplyr::group_by(REGION, APT, DATE, PHASE) |> 
  dplyr::mutate(N_VALID = rowSums(dplyr::across(EARLY:LATE)))  |> 
  dplyr::ungroup()

daily_punc <- daily_punc |> 
  dplyr::mutate(PUNC = rowSums(dplyr::across(EARLY_M15M05:LATE_P05P15)) / N_VALID)
```


df  # one row per airport per time-unit (day/week/year)
# region   : "BRA" / "EUR"
# airport  : airport code (ICAO)
# period   : Date (daily), or week start date, or year (numeric)
# kpi      : numeric value (e.g., arrival punctuality % or delay minutes)
# movements: optional numeric weights (e.g., flights)




### exploratory graph - ridgeline

```{r}
library(ggplot2)
```



```{r}
daily_punc |> # dplyr::filter(PHASE == "ARR") |> 
  ggplot(aes(x = PUNC, colour = PHASE, weight = N_VALID)) +
  geom_density(adjust = 1) +
  facet_wrap(~ APT, scales = "free_y")
```

## using ggridge

```{r}
library(ggridges)

daily_punc |> 
ggplot(aes(x = PUNC, y = APT, fill = PHASE, weight = N_VALID)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  facet_wrap(~ REGION, scale = "free_y")
```  

```{r}
punc_med <- daily_punc |> 
  group_by(REGION, APT, PHASE)  |> 
  reframe(P50 = median(PUNC, na.rm = TRUE)) |> 
group_by(REGION) |>
  mutate(
    y = as.numeric(factor(APT)),
    y0 = y - 0.18,
    y1 = y + 0.9
  ) |> 
  ungroup()


p <- daily_punc |>   
 ggplot(aes(x = PUNC, y = APT, fill = PHASE)) +
  geom_density_ridges(aes(weight = N_VALID), scale = 1) +
  facet_wrap(~ REGION, scales = "free_y")

p
```

```{r}
# started with plotting a point for the 50th percentile
p + 
  geom_point(data = punc_med, aes(x = P50, y = APT, color = PHASE))

# vline is global
# thus draw a "segment"
p +
  geom_segment(
    data = punc_med, 
    aes(x = P50, xend = P50, y = y0, yend = y1, color = PHASE))
```


```{r}
p <- daily_punc |>   
 ggplot(aes(x = PUNC, y = APT, fill = PHASE)) +
  geom_density_ridges(
  alpha = 0.6, scale = 1, rel_min_height = 0.01,
  quantile_lines = TRUE,
  quantiles = 0.5
)

p + 
  facet_wrap(~ REGION, scale = "free_y")
```


## using ggdist

```{r}
library(ggdist)

daily_punc |> 
ggplot(aes(y = APT, x = PUNC, fill = PHASE)) +
  ggdist::stat_halfeye() +
  facet_wrap(~ REGION, scale = "free_y")
```

```{r}
daily_punc |> 
ggplot(aes(y = APT, x = PUNC, fill = PHASE)) +
  ggdist::stat_halfeye(alpha = 0.6, position = "identity") +
  facet_wrap(~REGION, scale = "free_y")
```


```{r}

```

```{r}
airports <- tibble::tribble(
  ~REGION, ~APT,  ~lon,   ~lat,
  "BRA", "SBSV", -38.33, -12.91,
  "BRA", "SBSP", -46.66, -23.63,
  "BRA", "SBRJ", -43.17, -22.81,
  "BRA", "SBPA", -51.17, -29.99,
  "BRA", "SBKP", -47.14, -23.01,
  "BRA", "SBGR", -46.47, -23.43,
  "BRA", "SBGL", -43.24, -22.81,
  "BRA", "SBCT", -49.17, -25.53,
  "BRA", "SBCF", -43.97, -19.63,
  "BRA", "SBBR", -47.92, -15.87,
  "EUR", "LSZH",   8.55,  47.46,
  "EUR", "LPPT",  -9.13,  38.77,
  "EUR", "LFPG",   2.55,  49.01,
  "EUR", "LEMD",  -3.56,  40.47,
  "EUR", "LEBL",   2.08,  41.30,
  "EUR", "EHAM",   4.76,  52.31,
  "EUR", "EGLL",  -0.45,  51.47,
  "EUR", "EGKK",  -0.19,  51.15,
  "EUR", "EDDM",  11.79,  48.35,
  "EUR", "EDDF",   8.57,  50.05
)


map_kpi <- daily_punc |>
  filter(PHASE == "ARR") |>
  group_by(REGION, APT) |>
  summarise(P50_ARR = median(PUNC, na.rm = TRUE), .groups = "drop")

airports_map <- airports |>
  left_join(map_kpi, by = c("REGION", "APT"))

p_map <- ggplot(airports_map, aes(lon, lat)) +
  geom_point(aes(size = P50_ARR), alpha = 0.8) +
  geom_text(aes(label = APT), nudge_y = 0.8, size = 3) +
  facet_wrap(~ REGION, scales = "free") 
#+
#  coord_equal()

```


adding map and ridgelines does not work :( ~no nice story


```{r}
airport_layout <- tibble::tribble(
  ~APT,   ~row, ~col,
  # Brazil (north → south)
  "SBSV",  1,    2,
  "SBBR",  2,    2,
  "SBCF",  3,    3,
  "SBSP",  4,    4,
  "SBGR",  4,    5,
  "SBRJ",  5,    5,
  "SBGL",  5,    6,
  "SBCT",  6,    4,
  "SBPA",  7,    4,
  "SBKP",  4,    6,

  # Europe (west → east, north → south)
  "LPPT",  2,    1,
  "LEMD",  3,    2,
  "LEBL",  3,    3,
  "LFPG",  2,    4,
  "EGLL",  2,    5,
  "EGKK",  3,    5,
  "EHAM",  1,    4,
  "EDDF",  1,    5,
  "EDDM",  2,    6,
  "LSZH",  3,    6
)

daily_punc_geo <- daily_punc |>
  left_join(airport_layout, by = "APT")

```

```{r}
ggplot(
  daily_punc_geo,
  aes(x = PUNC, y = PHASE, fill = PHASE)
) +
  geom_density_ridges(scale = 0.9, alpha = 0.7) +
  facet_grid(row ~ col) + # , labeller = labeller(APT = label_value)) +
  theme_void()
```



```{r}
library(sf)
library(ggplot2)
library(cowplot)

world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

map_bg <- ggplot() +
  geom_sf(
    data = world |> dplyr::filter(admin %in% c("Brazil")),
    fill = "grey95", colour = "grey50"
  ) +
  geom_sf(
    data = world |> dplyr::filter(continent == "Europe"),
    fill = "grey95", colour = "grey50"
  ) +
  coord_sf(xlim = c(-80, 30), ylim = c(-40, 65), expand = FALSE) +
  theme_void()

```

```{r}

airport_ridge <- function(df) {
  ggplot(df, aes(x = PUNC, y = PHASE, fill = PHASE)) +
    geom_density_ridges(scale = 0.9, alpha = 0.7, rel_min_height = 0.01) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.margin = margin(2, 2, 2, 2)
    )
}

airport_pos <- tibble::tribble(
  ~APT,   ~lon,   ~lat,  ~dx,   ~dy,
  # Brazil
  "SBGR", -46.47, -23.43,  0,    0,
  "SBSP", -46.66, -23.63, -1.5, -0.5,
  "SBRJ", -43.17, -22.81,  0.8,  0,
  "SBGL", -43.24, -22.81, -0.8, -0.5,
  "SBSV", -38.33, -12.91,  0,    0,
  # Europe
  "EDDF",   8.57,  50.05,  0,    0,
  "EGLL",  -0.45,  51.47, -1.2,  0,
  "EHAM",   4.76,  52.31,  0.8,  0.4,
  "LFPG",   2.55,  49.01,  0,   -0.8,
  "LSZH",   8.55,  47.46,  0.8, -0.6
)

```

```{r}
final_plot <- map_bg

for (apt in airport_pos$APT) {

  ridge_grob <- airport_ridge(
    daily_punc |> dplyr::filter(APT == apt)
  ) |> ggplotGrob()

  pos <- airport_pos |> dplyr::filter(APT == apt)

  final_plot <- final_plot +
    cowplot::draw_plot(
      ridge_grob,
      x = scales::rescale(pos$lon + pos$dx, to = c(0, 1), from = c(-80, 30)),
      y = scales::rescale(pos$lat + pos$dy, to = c(0, 1), from = c(-40, 65)),
      width = 0.12,
      height = 0.06
    )
}

final_plot
```

```{r}
map_br <- ggplot() +
  geom_sf(
    data = world |> dplyr::filter(admin == "Brazil"),
    fill = "grey95", colour = "grey50"
  ) +
  coord_sf(
    xlim = c(-75, -35),
    ylim = c(-35,   5),
    expand = FALSE
  ) +
  theme_void()
```

```{r}
airport_ridge <- function(df_airport) {
  ggplot(df_airport, aes(x = PUNC, y = PHASE, fill = PHASE)) +
    geom_density_ridges(
      scale = 0.9,
      alpha = 0.7,
      rel_min_height = 0.01
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.margin = margin(1, 1, 1, 1)
    )
}


pos_br <- tibble::tribble(
  ~APT,   ~lon,   ~lat,  ~dx,   ~dy,
  "SBGR", -46.47, -23.43,  0,    0,
  "SBSP", -46.66, -23.63, -1.3, -0.4,
  "SBRJ", -43.17, -22.81,  0.9,  0,
  "SBGL", -43.24, -22.81, -0.9, -0.5,
  "SBSV", -38.33, -12.91,  0,    0,
  "SBCT", -49.17, -25.53, -0.6, -0.8,
  "SBPA", -51.17, -29.99,  0,   -0.6,
  "SBKP", -47.14, -23.01,  1.0, -0.6,
  "SBCF", -43.97, -19.63,  0.6,  0,
  "SBBR", -47.92, -15.87,  0,    0
)
```

```{r}
library(cowplot)

panel_br <- map_br

for (apt in pos_br$APT) {

  g <- airport_ridge(
    daily_punc |> dplyr::filter(APT == apt)
  ) |> ggplotGrob()

  p <- pos_br |> dplyr::filter(APT == apt)

  panel_br <- panel_br +
    draw_plot(
      g,
      x = scales::rescale(p$lon + p$dx, to = c(0, 1), from = c(-75, -35)),
      y = scales::rescale(p$lat + p$dy, to = c(0, 1), from = c(-35,   5)),
      width  = 0.12,
      height = 0.06
    )
}
```





panel_br <- ggdraw(map_br)

for (apt in pos_br$APT) {
    
    g <- airport_ridge(
        daily_punc |> dplyr::filter(APT == apt)
    ) |> ggplotGrob()
    
    p <- pos_br |> dplyr::filter(APT == apt)
    
    panel_br <- panel_br +
        draw_plot(
            g,
            x = scales::rescale(p$lon + p$dx, to = c(0, 1), from = c(-75, -35)),
            y = scales::rescale(p$lat + p$dy, to = c(0, 1), from = c(-35,   5)),
            width  = 0.12,
            height = 0.06
        )
}

panel_br





library(ggplot2)
library(sf)

p_base <- ggplot() +
  geom_sf(
    data = world |> dplyr::filter(admin == "Brazil"),
    fill = "grey95",
    colour = "grey60"
  ) +
  coord_sf(
    xlim = c(-75, -35),
    ylim = c(-35, 5),
    expand = FALSE
  ) +
  theme_void()

library(ggridges)
library(grid)

ridge_fun <- function(df_airport) {
  ggplot(df_airport, aes(x = PUNC, y = PHASE, fill = PHASE)) +
    geom_density_ridges(
      scale = 0.9,
      alpha = 0.7,
      rel_min_height = 0.01
    ) +
    theme_void() +
    theme(legend.position = "none")
}

g_sbgr <- ridge_fun(daily_punc |> dplyr::filter(APT == "SBGR")) |>
  ggplotGrob()
  
p_br <- p_base +
  annotation_custom(
    grob = g_sbgr,
    xmin = -48.5, xmax = -44.5,   # manually chosen
    ymin = -26.5, ymax = -22.5
  )


p_br <- p_br +
  geom_point(
    data = tibble(lon = -46.47, lat = -23.43),
    aes(lon, lat),
    size = 2
  ) +
  geom_segment(
    aes(
      x = -46.47, y = -23.43,
      xend = -44.5, yend = -24.5
    ),
    linewidth = 0.3,
    colour = "grey40"
  )


layout_br <- tibble::tribble(
  ~APT,   ~xmin,  ~xmax,  ~ymin,  ~ymax,
  "SBGR", -48.5,  -44.5,  -26.5,  -22.5,
  "SBRJ", -45.0,  -41.0,  -24.5,  -20.5,
  "SBSV", -40.5,  -36.5,  -15.0,  -11.0
)

for (i in seq_len(nrow(layout_br))) {
  row <- layout_br[i, ]

  g <- ridge_fun(
    daily_punc |> dplyr::filter(APT == row$APT)
  ) |> ggplotGrob()

  p_br <- p_br +
    annotation_custom(
      grob = g,
      xmin = row$xmin, xmax = row$xmax,
      ymin = row$ymin, ymax = row$ymax
    )
}
 

library(ggtext)

p_br +
  labs(
    title = "**Daily airport punctuality distributions – Brazil (2024)**",
    subtitle = "Arrival and departure OTP15 shown as ridgelines positioned near airport locations",
    caption = "Source: daily airport performance data"
  ) +
  theme(
    plot.title = element_markdown(),
    plot.subtitle = element_markdown(),
    plot.caption = element_text(hjust = 0)
  )


## airport map - augmented


world      <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
brazil_map <- world |> dplyr::filter(admin == "Brazil")
europe_map <- world |> dplyr::filter(continent == "Europe" | admin == "Turkey")

airports_brazil <- data.frame(
  icao    = c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG"),
  name    = c("Brasília","Guarulhos","Congonhas","Campinas","Santos Dumont","Galeão","Belo Horizonte","Salvador","Porto Alegre","Curitiba","Recife","Eduardo Gomes"),
  lon     = c(-47.92,-46.47,-46.66,-47.13,-43.17,-43.35,-43.97,-38.33,-51.17,-49.17,-34.92,-60.05),
  lat     = c(-15.87,-23.43,-23.63,-23.01,-22.91,-22.75,-19.62,-12.91,-29.99,-25.53, -8.13, -3.04),
  lbl_lon = c(-58.0,-57.0,-43.0,-57.0,-36.0,-35.0,-38.0,-45.0,-43.0,-62.0,-40.0,-60.0),
  lbl_lat = c(-13.0,-24.0,-27.5,-19.0,-21.0,-26.0,-16.0,-10.0,-32.0,-29.0, -3.0,  0.0)
)

airports_europe <- data.frame(
  icao    = c("EHAM","LFPG","EGLL","EDDF","EDDM","LEMD","LPPT","LEBL","EGKK","LSZH","LTFM","LGAV"),
  name    = c("Amsterdam","Paris","Heathrow","Frankfurt","Munich","Madrid","Lisbon","Barcelona","Gatwick","Zurich","Istanbul","Athens"),
  lon     = c(  4.76,  2.55, -0.46,  8.57, 11.79, -3.57, -9.13,  2.07, -0.19,  8.55, 28.75, 23.94),
  lat     = c( 52.31, 49.01, 51.48, 50.03, 48.35, 40.47, 38.77, 41.30, 51.15, 47.46, 41.27, 37.94),
  lbl_lon = c(  6.0, -3.0, -8.0, 15.0, 17.0, -2.0, -11.0, 9.0,  -7.0,  14.0,  33.0,  30.0),
  lbl_lat = c( 57.5, 45.0, 53.5, 54.5, 50.5,  38.0,  43.5, 39.0,  49.5,  45.5,  43.5,  37.5)
)

brazil_map = brazil_map |> dplyr::filter(icao %in% c("SBGR", "SBBR"))

map_brazil <- ggplot() +
  geom_sf(data = brazil_map, fill = "gray90", color = "white") +
  coord_sf(xlim = c(-74, -30), ylim = c(-34, 5), expand = FALSE) +
  geom_segment(data = airports_brazil,
               aes(x = lbl_lon, y = lbl_lat, xend = lon, yend = lat),
               color = "gray50", linewidth = 0.3) +
  geom_point(data = airports_brazil,
             aes(x = lon, y = lat), color = "#2E6B3E", size = 4) +
  geom_label(data = airports_brazil,
             aes(x = lbl_lon, y = lbl_lat, label = paste0(icao, "\n", name)),
             size = 5.7, lineheight = 0.8, fill = "white",
             label.padding = unit(0.35, "lines"), label.size = 1) +
  labs(subtitle = "Brazil") +
  theme_void() +
  theme(
    plot.subtitle    = element_text(size = 24, hjust = 0.5 , margin = margin(b = 4)),
    plot.margin      = margin(10, 10, 10, 10),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )

map_europe <- ggplot() +
  geom_sf(data = europe_map, fill = "gray90", color = "white") +
  coord_sf(xlim = c(-15, 37), ylim = c(35, 67), expand = FALSE) +
  geom_segment(data = airports_europe,
               aes(x = lbl_lon, y = lbl_lat, xend = lon, yend = lat),
               color = "gray50", linewidth = 0.3) +
  geom_point(data = airports_europe,
             aes(x = lon, y = lat), color = "#4472C4", size = 4) +
  geom_label(data = airports_europe,
             aes(x = lbl_lon, y = lbl_lat, label = paste0(icao, "\n", name)),
             size = 5.7, lineheight = 0.8, fill = "white",
             label.padding = unit(0.35, "lines"), label.size = 1) +
  labs(subtitle = "Europe") +
  theme_void() +
  theme(
    plot.subtitle    = element_text(size = 24, hjust = 0.5 , margin = margin(b = 4)),
    plot.margin      = margin(10, 10, 10, 10),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )

combined_plot <- map_brazil + map_europe +
  patchwork::plot_layout(widths = c(5, 5)) +
  patchwork::plot_annotation(
    caption = "Charts for both regions not in scale",
    theme   = theme(plot.caption = element_text(size = 20, hjust = 1, color = "#555555"))
  )

combined_plot

ggplot2::ggsave(
  filename = here::here("figures", "scope-airports-2025.png"),
  plot     = combined_plot,
  width    = 16, height = 8, dpi = 100,
  units    = "in", bg = "white"
)


inset_layout_br <- data.frame(
  icao = c("SBGR", "SBSP", "SBRJ", "SBSV", "SBPA"),
  xmin = c(-55, -42, -36, -46, -42),
  xmax = c(-49, -36, -30, -40, -36),
  ymin = c(-26, -31, -26, -15, -34),
  ymax = c(-22, -27, -22, -11, -30)
)

library(ggridges)
library(grid)

ridge_inset <- function(df_airport) {
  ggplot(df_airport, aes(x = PUNC, y = PHASE, fill = PHASE)) +
    geom_density_ridges(
      scale = 0.9,
      alpha = 0.75,
      rel_min_height = 0.01
    ) +
    coord_cartesian(xlim = c(0.3, 0.95)) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.margin = margin(2, 2, 2, 2)
    )
}

```{r}
map_brazil_insets <- map_brazil

for (i in seq_len(nrow(inset_layout_br))) {

  apt  <- inset_layout_br$icao[i]
  box  <- inset_layout_br[i, ]

  g <- ridge_inset(
    daily_punc |> dplyr::filter(APT == apt)
  ) |> ggplotGrob()

  map_brazil_insets <- map_brazil_insets +
    annotation_custom(
      grob = g,
      xmin = box$xmin, xmax = box$xmax,
      ymin = box$ymin, ymax = box$ymax
    )
}

# connector lines
map_brazil_insets <- map_brazil_insets +
  geom_segment(
    data = airports_brazil |> dplyr::filter(icao == "SBGR"),
    aes(
      x = lon, y = lat,
      xend = -49, yend = -24
    ),
    linewidth = 0.4,
    colour = "gray40"
  )
``
```

# helper to create same-sized inset boxes across regions

```{r}
make_inset_box <- function(
  center_lon, center_lat,
  xlim, ylim,
  frac_x = 0.18,   # width as fraction of panel
  frac_y = 0.16    # height as fraction of panel
) {
  width  <- diff(xlim) * frac_x
  height <- diff(ylim) * frac_y

  tibble::tibble(
    xmin = center_lon - width  / 2,
    xmax = center_lon + width  / 2,
    ymin = center_lat - height / 2,
    ymax = center_lat + height / 2
  )
}

xlim_br <- c(-74, -30)
ylim_br <- c(-34, 5)

inset_centers_br <- tibble::tribble(
  ~icao,  ~cx,   ~cy,
  "SBGR", -52,  -24,
  "SBRJ", -33,  -24,
  "SBSV", -43,  -13
  # ---- ADD MORE AIRPORTS HERE ----
)


insets_br <- inset_centers_br |>
  rowwise() |>
  dplyr::bind_cols(
    make_inset_box(
      center_lon = .$cx,
      center_lat = .$cy,
      xlim = xlim_br,
      ylim = ylim_br,
      frac_x = 0.18,
      frac_y = 0.16
    )
  ) |>
  ungroup()

```




ggplot() +
geom_sf(data = brazil_map, fill = "grey90") +
geom_point(data = airports_brazil |> dplyr::filter(icao %in% c("SBGR")),
  aes(x = lon, y = lat), color = "#2E6B3E", size = 4)


```{r}
xlim_eu <- c(-15, 37)
ylim_eu <- c(35, 67)

inset_centers_eu <- tibble::tribble(
  ~icao,  ~cx, ~cy,
  "EDDF",  18, 52,
  "EGLL",  -8, 53,
  "LEMD",   0, 39
)

insets_eu <- inset_centers_eu |>
  rowwise() |>
  dplyr::bind_cols(
    make_inset_box(
      center_lon = .$cx,
      center_lat = .$cy,
      xlim = xlim_eu,
      ylim = ylim_eu,
      frac_x = 0.18,
      frac_y = 0.16
    )
  ) |>
  ungroup()
```

https://github.com/DOI-USGS/vizlab-chart-challenge-23/blob/main/28_trends_msleckman/28_trends_msleckman.Rmd