---
title: "Day 08 - circular"
format: html
---

Circular “route wheel” for international flows
•	What it could show: distribution of international traffic flows to/from Brazil vs Europe by world region, arranged around a circle.
•	Data to tap: OD or entry/exit region mapping; aggregate by destination region and weight by flights/ASK.
•	Discussion hook: link traffic structure to performance (do long-haul shares correlate with punctuality tails?).
•	ggplot2 angle: chord-like effect using ggforce::geom_arc_bar / geom_bezier; more experimental.

```{}

# get is from local file - only works on RQ's comp
path_one_up <- here::here() |> dirname()
path_nm_flt <- here::here(path_one_up, "__DATA","NM-flight-table")

nm_flt_2025 <- "NM-flt-2025.zip"

nm_flt <- read_zip_content(path_nm_flt, nm_flt_2025) |> 
  dplyr::bind_rows()
```

```{r}
# study airports
bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```

```{}
conns <- nm_flt |> 
    dplyr::group_by(ADEP, ADES) |> 
    dplyr::reframe(DEPS = n()) |> 
    dplyr::mutate(
        ADEP_REG = dplyr::case_when(
             ADEP == "AFIL"  ~ "EUR"    # air filed FPLs only regional
            ,grepl(pattern = "^[E|L|BI]", ADEP) ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADEP) ~ "BRA"
            , .default = "INTNL"
            )
        ,ADES_REG = dplyr::case_when(
             grepl(pattern = "^[E|L|BI]", ADES)         ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADES) ~ "BRA"
            , .default = "INTNL"
        )
    )

conns |> group_by(ADEP_REG, ADES_REG) |> reframe(COMBOS = dplyr::n(), DEPS = sum(DEPS))
```

```{}
eur_bra <- conns |> 
  dplyr::filter(ADEP_REG == "EUR", ADES_REG == "BRA") 

eur_bra <- eur_bra |> 
  dplyr::mutate(
      ADEP_STUDY = ifelse(ADEP %in% eur_apts, ADEP, "OTHER_EUR")
    , ADES_STUDY = ifelse(ADES %in% bra_apts, ADES, "OTHER_BRA")
    ) |> 
  dplyr::group_by(ADEP_STUDY, ADES_STUDY) |> 
  dplyr::reframe(DEPS = sum(DEPS))

eur_bra |> readr::write_csv("./data/BRA-EUR-aerodrome-pair-connections.csv")

eur_bra
```

## Playground

Chord diagrams with {circlize} explained in https://r-charts.com/flow/chord-diagram/

Basic data structure

  FROM - TO - VALUE   ~ ADEP - ADES - FLTS
  
we look at this from a bi-directional perspective: ADEP-ADES is an aerodrome pair and not a flight "direction".

```{r}
# load data
eur_bra <- readr::read_csv("./data/BRA-EUR-aerodrome-pair-connections.csv", show_col_types = F)
```

```{r}
library(circlize)

# basic chord diagram
circos.clear()           # clear graphic device
chordDiagram(eur_bra)
```

```{r}
circos.clear() 

grid.col = c(EDDF = "red",LPPT = "green")

chordDiagram(eur_bra, big.gap = 10, grid.col = grid.col, directional = 1)
```

```{r}
circos.clear() 

grid.col = c(EDDF = "red",LPPT = "green")

chordDiagram(eur_bra, big.gap = 10, grid.col = grid.col, directional = 1)
```



```{r}
eur_order <- c(eur_apts, "OTHER_EUR")
bra_order <- c(bra_apts, "OTHER_BRA")

sector_order <- c(eur_order, bra_order)

grid.col <- c(
    setNames(rep("#4472C4", length(eur_order)), eur_order),
    setNames(rep("#2E6B3E", length(bra_order)), bra_order)
    )

circos.clear() 

#grid.col = c(EDDF = "red",LPPT = "green")
arr.col <- data.frame(
   c("LPPT")
  ,c("SBGR")
  ,c("black")
)

chordDiagram(
  eur_bra, big.gap = 10, grid.col = grid.col
  , directional = 1
  #, direction.type = "arrows"
  , direction.type = c("diffHeight", "arrows")
  , link.arr.col = arr.col
  )

```

```{r}
circos.clear() 
chordDiagram(
  eur_bra
  
  , directional = 1
#  , direction.type = c("diffHeight", "arrows")
#  , link.arr.type = "big.arrow", diffHeight = -mm_h(2)
  
#  ,scale = TRUE
  
  ,annotationTrack = "grid"
  ,preAllocateTracks = 1
  )

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.6
    )
  },
  bg.border = NA
)
```


```{r}
circos.clear() 

circos.par(start.degree = 0)

chordDiagram(
  eur_bra
  
  , directional = 1
#  , direction.type = c("diffHeight", "arrows")
#  , link.arr.type = "big.arrow", diffHeight = -mm_h(2)
  
#  ,scale = TRUE
  
  ,annotationTrack = "grid"
  ,preAllocateTracks = 1
  )



circos.track(track.index = 1, panel.fun = function(x, y) {
    xlim = get.cell.meta.data("xlim")
    xplot = get.cell.meta.data("xplot")
    ylim = get.cell.meta.data("ylim")
    sector.name = get.cell.meta.data("sector.index")

    if(abs(xplot[2] - xplot[1]) < 10) {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise",
            niceFacing = TRUE, adj = c(0, 0.5), col = "red")
    } else {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "inside", 
            niceFacing = TRUE, adj = c(0.5, 0), col= "blue")
    }
}, bg.border = NA)
```

cannot get this to work.

```{}

wheel_data <- eur_bra |> 
  dplyr::rename(ADEP_GRP = ADEP_STUDY, ADES_GRP=ADES_STUDY, FLTS = DEPS)

circos.clear()

circos.par(
  start.degree = 90,
  gap.after = c(rep(2, length(eur_order) - 1), 10,
                rep(2, length(bra_order) - 1), 10)
)

chordDiagram(
  x = wheel_data,
  order = sector_order,
  grid.col = c(
    setNames(rep("#4472C4", length(eur_order)), eur_order),
    setNames(rep("#2E6B3E", length(bra_order)), bra_order)
  ),
  transparency = 0.5,
  annotationTrack = "grid",
  preAllocateTracks = 1
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.6
    )
  },
  bg.border = NA
)

```


```{r}
# parameters
circos.clear()
circos.par(start.degree = 90, gap.degree = 4, track.margin = c(-0.1, 0.1), points.overflow.warning = FALSE)

# color palette
mycolor <- viridis::viridis(10, alpha = 1, begin = 0, end = 1, option = "D")
mycolor <- mycolor[sample(1:10)]

# base plot
chordDiagram(
  x = eur_bra, 
 # grid.col = mycolor,
  transparency = 0.25,
  directional = 1,
#  direction.type = c("arrows", "diffHeight")
  diffHeight  = -0.04,
  annotationTrack = "grid", 
  annotationTrackHeight = c(0.05, 0.1)
)


# Add text and axis
circos.trackPlotRegion(
  track.index = 1, 
  bg.border = NA, 
  panel.fun = function(x, y) {
    
    xlim = get.cell.meta.data("xlim")
    sector.index = get.cell.meta.data("sector.index")
    
    # Add names to the sector. 
    circos.text(
      x = mean(xlim), 
      y = 3.2, 
      labels = sector.index, 
      facing = "bending", 
      cex = 0.8
      )
  # # Add graduation on axis
  #   circos.axis(
  #     h = "top", 
  #     major.at = seq(from = 0, to = xlim[2], by = ifelse(test = xlim[2]>10, yes = 2, no = 1)), 
  #     minor.ticks = 1, 
  #     major.tick.percentage = 0.5,
  #     labels.niceFacing = FALSE)
  }
)
```

```{r}
library(igraph)
library(ggraph)
library(tidyverse)

# Example: create a small graph
edges <- data.frame(
  from = c("A", "A", "B", "C", "D", "E"),
  to   = c("B", "C", "D", "E", "F", "A")
)

# Create igraph object
graph <- graph_from_data_frame(edges, directed = TRUE)

# Plot circular graph
ggraph(graph, layout = "circle") +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)),
                 arrow = arrow(length = unit(4, 'mm')),
                 end_shape = "circle",
                 color = "grey50") +
  geom_node_point(size = 5, color = "skyblue") +
  geom_node_text(aes(label = name), repel = TRUE, size = 4) +
  theme_void()
```
# Install required packages if not already installed
if (!requireNamespace("ggchord", quietly = TRUE)) {
  install.packages("remotes")
  remotes::install_github("DangJem/ggchord")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

library(ggplot2)
library(ggchord)

# Example data: relationships between categories
data <- data.frame(
  from = c("A", "A", "B", "B", "C", "D"),
  to   = c("B", "C", "C", "D", "D", "A"),
  value = c(10, 5, 7, 3, 8, 6)
)

# Create chord diagram
ggchord(
  data,
  from = "from",
  to = "to",
  value = "value",
  colour = "from",       # Color arcs by 'from' category
  grid.col = NULL,       # Auto color grid
  transparency = 0.25    # Transparency for links
) +
  theme_void() +
  ggtitle("Chord Diagram with ggplot2 (ggchord)")

