---
title: "Day 08 - circular"
format: html
---

```{r}
library(ggplot2)
library(ggview)
```


Circular “route wheel” for international flows
•	What it could show: distribution of international traffic flows to/from Brazil vs Europe by world region, arranged around a circle.
•	Data to tap: OD or entry/exit region mapping; aggregate by destination region and weight by flights/ASK.
•	Discussion hook: link traffic structure to performance (do long-haul shares correlate with punctuality tails?).
•	ggplot2 angle: chord-like effect using ggforce::geom_arc_bar / geom_bezier; more experimental.

```{}

# get is from local file - only works on RQ's comp
path_one_up <- here::here() |> dirname()
path_nm_flt <- here::here(path_one_up, "__DATA","NM-flight-table")

nm_flt_2025 <- "NM-flt-2025.zip"

nm_flt <- read_zip_content(path_nm_flt, nm_flt_2025) |> 
  dplyr::bind_rows()
```

```{r}
# study airports
bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```

```{}
conns <- nm_flt |> 
    dplyr::group_by(ADEP, ADES) |> 
    dplyr::reframe(DEPS = n()) |> 
    dplyr::mutate(
        ADEP_REG = dplyr::case_when(
             ADEP == "AFIL"  ~ "EUR"    # air filed FPLs only regional
            ,grepl(pattern = "^[E|L|BI]", ADEP) ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADEP) ~ "BRA"
            , .default = "INTNL"
            )
        ,ADES_REG = dplyr::case_when(
             grepl(pattern = "^[E|L|BI]", ADES)         ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADES) ~ "BRA"
            , .default = "INTNL"
        )
    )

conns |> group_by(ADEP_REG, ADES_REG) |> reframe(COMBOS = dplyr::n(), DEPS = sum(DEPS))
```

```{}
eur_bra <- conns |> 
  dplyr::filter(ADEP_REG == "EUR", ADES_REG == "BRA") 

eur_bra <- eur_bra |> 
  dplyr::mutate(
      ADEP_STUDY = ifelse(ADEP %in% eur_apts, ADEP, "OTHER_EUR")
    , ADES_STUDY = ifelse(ADES %in% bra_apts, ADES, "OTHER_BRA")
    ) |> 
  dplyr::group_by(ADEP_STUDY, ADES_STUDY) |> 
  dplyr::reframe(DEPS = sum(DEPS))

eur_bra |> readr::write_csv("./data/BRA-EUR-aerodrome-pair-connections.csv")

eur_bra
```

## Playground

Chord diagrams with {circlize} explained in https://r-charts.com/flow/chord-diagram/

Basic data structure

  FROM - TO - VALUE   ~ ADEP - ADES - FLTS
  
we look at this from a bi-directional perspective: ADEP-ADES is an aerodrome pair and not a flight "direction".

```{r}
# load data
eur_bra <- readr::read_csv("./data/BRA-EUR-aerodrome-pair-connections.csv", show_col_types = F)
```

```{r}
library(circlize)

# basic chord diagram
circos.clear()           # clear graphic device
chordDiagram(eur_bra)
```

```{r}
circos.clear() 

grid.col = c(EDDF = "red",LPPT = "green")

chordDiagram(eur_bra, big.gap = 10, grid.col = grid.col, directional = 1)
```

```{r}
circos.clear() 

grid.col = c(EDDF = "red",LPPT = "green")

chordDiagram(eur_bra, big.gap = 10, grid.col = grid.col, directional = 1)
```



```{r}
eur_order <- c(eur_apts, "OTHER_EUR")
bra_order <- c(bra_apts, "OTHER_BRA")

sector_order <- c(eur_order, bra_order)

grid.col <- c(
    setNames(rep("#4472C4", length(eur_order)), eur_order),
    setNames(rep("#2E6B3E", length(bra_order)), bra_order)
    )

circos.clear() 

#grid.col = c(EDDF = "red",LPPT = "green")
arr.col <- data.frame(
   c("LPPT")
  ,c("SBGR")
  ,c("black")
)

chordDiagram(
  eur_bra, big.gap = 10, grid.col = grid.col
  , directional = 1
  #, direction.type = "arrows"
  , direction.type = c("diffHeight", "arrows")
  , link.arr.col = arr.col
  )

```

```{r}
circos.clear() 
chordDiagram(
  eur_bra
  
  , directional = 1
#  , direction.type = c("diffHeight", "arrows")
#  , link.arr.type = "big.arrow", diffHeight = -mm_h(2)
  
#  ,scale = TRUE
  
  ,annotationTrack = "grid"
  ,preAllocateTracks = 1
  )

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.6
    )
  },
  bg.border = NA
)
```


```{r}
circos.clear() 

circos.par(start.degree = 0)

chordDiagram(
  eur_bra
  
  , directional = 1
#  , direction.type = c("diffHeight", "arrows")
#  , link.arr.type = "big.arrow", diffHeight = -mm_h(2)
  
#  ,scale = TRUE
  
  ,annotationTrack = "grid"
  ,preAllocateTracks = 1
  )



circos.track(track.index = 1, panel.fun = function(x, y) {
    xlim = get.cell.meta.data("xlim")
    xplot = get.cell.meta.data("xplot")
    ylim = get.cell.meta.data("ylim")
    sector.name = get.cell.meta.data("sector.index")

    if(abs(xplot[2] - xplot[1]) < 10) {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise",
            niceFacing = TRUE, adj = c(0, 0.5), col = "red")
    } else {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "inside", 
            niceFacing = TRUE, adj = c(0.5, 0), col= "blue")
    }
}, bg.border = NA)
```

cannot get this to work.

```{}

wheel_data <- eur_bra |> 
  dplyr::rename(ADEP_GRP = ADEP_STUDY, ADES_GRP=ADES_STUDY, FLTS = DEPS)

circos.clear()

circos.par(
  start.degree = 90,
  gap.after = c(rep(2, length(eur_order) - 1), 10,
                rep(2, length(bra_order) - 1), 10)
)

chordDiagram(
  x = wheel_data,
  order = sector_order,
  grid.col = c(
    setNames(rep("#4472C4", length(eur_order)), eur_order),
    setNames(rep("#2E6B3E", length(bra_order)), bra_order)
  ),
  transparency = 0.5,
  annotationTrack = "grid",
  preAllocateTracks = 1
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.6
    )
  },
  bg.border = NA
)

```



```{r}
# parameters
circos.clear()
circos.par(start.degree = 90, gap.degree = 4, track.margin = c(-0.1, 0.1), points.overflow.warning = FALSE)

# color palette
mycolor <- viridis::viridis(10, alpha = 1, begin = 0, end = 1, option = "D")
mycolor <- mycolor[sample(1:10)]

# base plot
chordDiagram(
  x = eur_bra, 
 # grid.col = mycolor,
  transparency = 0.25,
  directional = 1,
#  direction.type = c("arrows", "diffHeight")
  diffHeight  = -0.04,
  annotationTrack = "grid", 
  annotationTrackHeight = c(0.05, 0.1)
)


# Add text and axis
circos.trackPlotRegion(
  track.index = 1, 
  bg.border = NA, 
  panel.fun = function(x, y) {
    
    xlim = get.cell.meta.data("xlim")
    sector.index = get.cell.meta.data("sector.index")
    
    # Add names to the sector. 
    circos.text(
      x = mean(xlim), 
      y = 3.2, 
      labels = sector.index, 
      facing = "bending", 
      cex = 0.8
      )
  # # Add graduation on axis
  #   circos.axis(
  #     h = "top", 
  #     major.at = seq(from = 0, to = xlim[2], by = ifelse(test = xlim[2]>10, yes = 2, no = 1)), 
  #     minor.ticks = 1, 
  #     major.tick.percentage = 0.5,
  #     labels.niceFacing = FALSE)
  }
)
```

```{r}
library(igraph)
library(ggraph)
library(tidyverse)

# Example: create a small graph
edges <- data.frame(
  from = c("A", "A", "B", "C", "D", "E"),
  to   = c("B", "C", "D", "E", "F", "A")
)

# Create igraph object
graph <- graph_from_data_frame(edges, directed = TRUE)

# Plot circular graph
ggraph(graph, layout = "circle") +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)),
                 arrow = arrow(length = unit(4, 'mm')),
                 end_shape = "circle",
                 color = "grey50") +
  geom_node_point(size = 5, color = "skyblue") +
  geom_node_text(aes(label = name), repel = TRUE, size = 4) +
  theme_void()
```

# Install required packages if not already installed
if (!requireNamespace("ggchord", quietly = TRUE)) {
  install.packages("remotes")
  remotes::install_github("DangJem/ggchord")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

library(ggplot2)
library(ggchord)

# Example data: relationships between categories
data <- data.frame(
  from = c("A", "A", "B", "B", "C", "D"),
  to   = c("B", "C", "C", "D", "D", "A"),
  value = c(10, 5, 7, 3, 8, 6)
)

# Create chord diagram
ggchord(
  data,
  from = "from",
  to = "to",
  value = "value",
  colour = "from",       # Color arcs by 'from' category
  grid.col = NULL,       # Auto color grid
  transparency = 0.25    # Transparency for links
) +
  theme_void() +
  ggtitle("Chord Diagram with ggplot2 (ggchord)")


eur_bra_tofm <- eur_bra |> rename(from = ADEP_STUDY, to = ADES_STUDY)

# crreate igraph object
our_graph <- graph_from_data_frame(eur_bra_tofm) 

# plot circular graph
ggraph(our_graph, layout = "circle") +
  geom_edge_link(
    aes( start_cap = label_rect(node1.name)
        ,end_cap = label_rect(node2.name))
  ) +
  geom_node_point(size = 5, color = "skyblue") +
  geom_node_text(aes(label = name), repel = TRUE)

```{r}
mywidth = 7
myheight = 9
mydpi = 300
```


```{r}
nodes <- tibble(
  name = unique(c(eur_bra$ADEP_STUDY, eur_bra$ADES_STUDY))
) |> mutate(
  region = ifelse(grepl("^S|^OTHER_BRA" ,x = name), "BRA", "EUR")
)

edges <- eur_bra |> 
  transmute(
    from = ADEP_STUDY
    , to = ADES_STUDY
    , weight = DEPS
    , highlight = case_when(
         from == "LPPT" ~ "top"
       , to   == "SBGR" ~ "top"
       , .default = "normal"
    ,
    )
    ) |> 
  filter(
    !is.na(from)
    ,!is.na(to)
    ,from %in% nodes$name
    ,to %in% nodes$name
  )

# create graph object
mygraph <- graph_from_data_frame(edges, vertices = nodes, directed = TRUE)

plot_circular_network <- function(this_graph, ...){
  
  gg <- ggraph(this_graph, layout = "circle") +
  geom_edge_arc(
    aes(
       edge_width = weight
      ,edge_alpha = weight
      ,edge_colour = highlight    # set above in edges
    )
 #   , color = "grey40"   # link color now controlled separately
    , strength = 0.4
    , end_cap = circle(3, "mm")
  ) +
  
  scale_edge_color_manual(
    values = c(top = "#4472C4", normal = "grey30")
  ) +
  
  geom_node_point(
    aes(color = region)
    ,size = 8
  ) +
  geom_node_text(
    aes(label = 
           ifelse(name == "OTHER_EUR","OTHER\nEUR",
              ifelse(name == "OTHER_BRA","OTHER\nBRA"
                     , name)        
                  )
         
        ,angle = node_angle(x,y)
        , x = x * 1.1  # Push labels outward horizontally
        , y = y * 1.1  # Push labels outward vertically
        , hjust = x < 0
        ) # end aes
    ,size = 5
    
    
  ) +
  scale_edge_width(range = c(0.3, 3)) +
  scale_edge_alpha(range = c(0.2, 0.8)) +
  
  coord_fixed(
     xlim = c(-1.5, 1.5)
    ,ylim = c(-1.5, 1.5)
  ) +
  
  theme_void() +
  theme(
     legend.position = "top"
    ,legend.direction = "horizontal"
  ) +
  
  canvas(width = mywidth, height = myheight, unit = "in", dpi = mydpi)
  
  return(gg)
}

```

```{r}
library(ggtext)
library(ggview)

our_title <-"Chart-Challenge Day 08 – ***FLIGHT CONNECTIVITY* <br>between the study airports in Brazil and Europe during the year 2025**"
our_sub <-
  "São Paulo/Guarulhos (SBGR) highly connected to the major European study airport, in particular with regular services to Lisbon (LPPT) and Madrid (LEMD). Services from Lisbon (LPPT) connect to most Brazilian in the study. Other European airports (OTHER EUR) service São Paulo (SBGR), Rio de Janeiro/Galeãlo (SBGL), and Campinas (SBKP)."
our_caption <-
  "DECEA Performance Section & EUROCONTROL PRU <br> data available for download on GitHub."


mygraph |> plot_circular_network() +
  labs(
     title    = our_title
    ,subtitle = our_sub
    ,caption = our_caption
  ) +
  theme(
     plot.title    = element_textbox_simple(
       size = rel(1.5)
       ,margin = margin(t = 20, b=20)
     )
    ,plot.title.position = "plot"
    ,plot.subtitle = element_textbox_simple(
      size = rel(1.2)
      ,margin = margin(t = 10, b = 20)
    )
    ,plot.caption  = element_textbox_simple(
      margin = margin(t = 10, b = 10)
    ) 
  )

```
 

ggsave("./figures/day08-circular.png", 
        width  = mywidth
      , height = myheight
      , unit = "in"
      , dpi = 300)
      

```{r}
# Define your color palette
bra_eur_colours <- c(BRA = "#52854C", EUR = "#4E84C4")

plot_circular_network <- function(this_graph, ...){
  
  gg <- ggraph(this_graph, layout = "circle") +
  geom_edge_arc(
    aes(
       edge_width = weight
      ,edge_alpha = weight
      ,edge_colour = highlight
    )
    , strength = 0.4
    , end_cap = circle(3, "mm")
  ) +
  
  scale_edge_color_manual(
    name = NULL, #"Highlighted Routes",  # Add legend title
    values = c(top = "#4472C4", normal = "grey30"),
    labels = c(top = "main LPPT/\nSBGR routes"
               , normal = "minor routes/\nservices")  # Better labels
  ) +
  
  geom_node_point(
    aes(color = region)
    ,size = 8
  ) +
  
  scale_color_manual(  # Add this for node colors
    name = NULL, # "Region",
    values = bra_eur_colours
    ,labels = c(BRA = "Brazil", EUR = "Europe") 
  ) +
  
  geom_node_text(
    aes(label = 
           ifelse(name == "OTHER_EUR","OTHER\nEUR",
              ifelse(name == "OTHER_BRA","OTHER\nBRA"
                     , name)        
                  )
         
        ,angle = node_angle(x,y)
        , x = x * 1.1
        , y = y * 1.1
        , hjust = x < 0
        )
    ,size = 5
  ) +
  
  scale_edge_width(range = c(0.3, 3), guide = "none") +  # Hide width legend
  scale_edge_alpha(range = c(0.2, 0.8), guide = "none") +  # Hide alpha legend
  
  coord_fixed(
     xlim = c(-1.5, 1.5)
    ,ylim = c(-1.5, 1.5)
  ) +
  
  theme_void() +
  theme(
     legend.position = "inside"  # Enable positioning inside plot
    ,legend.position.inside = c(0.15, 0.92)  # Top left for regions
    ,legend.direction = "vertical"
    ,legend.box = "vertical"
  ) +
  
  guides(
    color = guide_legend(
      order = 1, 
      override.aes = list(size = 4),
      theme = theme(legend.position.inside = c(0.05, 0.95))  # Top left
    ),
    edge_colour = guide_legend(
      order = 2,
      theme = theme(legend.position.inside = c(0.9, 0.05))  # Bottom right
    )
  )
  
  return(gg)
}
```

```{r}
library(ggtext)
library(ggview)

our_title <-"Chart-Challenge Day 08 – ***FLIGHT CONNECTIVITY* between the study airports in Brazil and Europe during the year 2025**"
our_sub <-
  "São Paulo/Guarulhos (SBGR) highly connected to the major European study airport, in particular with regular services to Lisbon (LPPT) and Madrid (LEMD). Services from Lisbon (LPPT) connect to most Brazilian airports in the study. Other European airports (OTHER EUR) service São Paulo (SBGR), Rio de Janeiro/Galeãlo (SBGL), and Campinas (SBKP)."
our_caption <-
  "DECEA Performance Section & EUROCONTROL PRU <br> data available for download on GitHub."


mygraph |> plot_circular_network() +
  labs(
     title    = our_title
    ,subtitle = our_sub
    ,caption = our_caption
  ) +
  theme(
     plot.title    = element_textbox_simple(
       size = rel(1.5)
       ,lineheight = 1.1
       ,margin = margin(t = 20, b=20)
     )
    ,plot.title.position = "plot"
    ,plot.subtitle = element_textbox_simple(
      size = rel(1.2)
      ,margin = margin(t = 10, b = 20)
    )
    ,plot.caption  = element_textbox_simple(
      margin = margin(t = 10, b = 10)
    ) 
    ,plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  
  canvas(width = mywidth, height = myheight, unit = "in", dpi = mydpi)

```

    
ggsave("./figures/day08-circular.png", 
        width  = mywidth
      , height = myheight
      , unit = "in"
      , dpi = 300
      , bg = "white")  
      
