---
title: "Day 06 - reporter without borders"
format: html
---

```{r}
library(dplyr)   # data crunching
library(ggplot2) # viz
```


## Data retrieval for "reporters without borders"

Kudos to Peter Baumgartner and his description of how to downlaod and access the reporters-without-borders (reporters sans frontiers) data: https://bookdown.org/pbaumgartner/rwb-book/011-get-rwb-data.html.

DATA DOWNLOADED FOR 2022 - 2025 - CLEAN. SKIP NEXT CHUNKS.

```{r}
# taken from https://bookdown.org/pbaumgartner/rwb-book/011-get-rwb-data.html
url <- "https://rsf.org/sites/default/files/import_classement/"

rsf_year <- function(year) {
    rsf_name <- paste0(url, year, ".csv")
    readr::read_delim(rsf_name,
                      delim = ";", 
                      escape_double = FALSE, 
                      trim_ws = TRUE,
                      show_col_types = FALSE)
}


# per description we work with the 2022-2025 data
our_years <- list(
    #"2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010",
    # 2011 is missing
    #"2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020",
    #"2021", 
  "2022", "2023", "2024", "2025")

#rsf <- our_years |> purrr::map(.f = ~ rsf_year(.x) )
```

ON PETER'S PAGE - CODE 1.7 HIGHLIGHTS THE CLEANED DATA --> get this!

```{r}
#| eval: false
url <- "https://raw.githubusercontent.com/petzi53/rwb-book/master/data/chap011/rwb/rwb1.rds"

download.file(url, destfile = "rwb1.rds", mode = "wb")

rwb <- readRDS("rwb1.rds")
rwb |> readr::write_csv("./data/reporter-whithout-borders-2022-2025.csv")
```

## Exploration

```{r}
# load data
rwb <- readr::read_csv("./data/reporter-whithout-borders-2022-2025.csv", show_col_types = FALSE)

rwb |> glimpse()
```

* 195 countries = difficult to plot per country ==> either subset or groups
* data set has evolution year_n/score/rank vs rank_n_1 and score_n_1 ~ no need to calculate this (thanks Peter)
* zone is a "geographic" or reporter score relevant regional grouping => 8 zones
rwb |> distinct(zone)
# A tibble: 8 × 1
  zone                  
  <chr>                 
1 Asie-Pacifique        
2 UE Balkans            
3 MENA                  
4 Afrique               
5 Amériques             
6 EEAC                  
7 Europe - Asie centrale
8 Maghreb - Moyen-Orient

## Initial idea

* showing change 2024-2025 ~~ difficult for number of countries ~~ volcano chart to show "biggest changes"
* classical barchart showing index for each country  ~~ tiles     

### Volcano Plot

change from 2024 to 2025 already calculated in data

payload

* country_en
* year_n
* score_evolution

```{r}
library(ggrepel)     # deconflict labels

# prep data - volcano --> change per country for 2025 vs 2024
rwb_volcano <- rwb |>
  # ------------------- get payload & trim to 2025 with change
  select(country_en, year_n, score_evolution) |> 
  filter(year_n == 2025) |>
  
  # codify "change" direction
  mutate(
    direction = case_when(
      score_evolution > 0 ~ "Improved",
      score_evolution < 0 ~ "Declined",
      .default = "No change"
        ),
    direction = factor(
      direction
      , levels = c("Improved", "Declined", "No change"))
    )

p <- ggplot(
  rwb_volcano
  , aes(
      x = score_evolution,
      y = abs(score_evolution),     # volcano -> negative values up!
      color = direction
      )
  ) +
    geom_point(alpha = 0.7, size = 3) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
    scale_color_manual(values = c(
        "Improved" = "#1a9850",
        "Declined" = "#d73027",
        "No change" = "gray60"
    )) +
    ggrepel::geom_text_repel(
        data = rwb_volcano |> filter(abs(score_evolution) > 5),
        aes(label = country_en),
        size = 3,
        max.overlaps = 20
    )

p +
    theme_minimal(base_size = 14) +
    labs(
        title = "Volcano Plot of Press Freedom Score Changes (2025)",
        subtitle = "Horizontal = direction of change • Vertical = magnitude of change",
        x = "Score Evolution (2025 vs 2024)",
        y = "Magnitude of Change",
        color = "Direction"
    )

```



### "Tiled" Bar Chart

tile = country <==> fill = score

==> payload:
* year_n      year of observation
* country_en  "object" ~ country of observation
* zone        regional grouping per RWB
* score       assigned score


```{r}
#| fig-height: 6     # increase standard size to make is less squished


rwb_tiles <- rwb |> 
  # subset to payload
  select(year_n, country_en, zone, score) |> 
  # group by zone and arrange for index sequence, set index = row_number
  group_by(year_n, zone) |> 
  arrange(score, .by_group = TRUE) |> 
  mutate(order = row_number())

p <- ggplot(rwb_tiles, aes(x = order, y = zone, fill = score)) +
  geom_tile(color = "white") +                # tile borders
  scale_fill_viridis_c(option = "C") +
  facet_wrap(~ year_n, ncol = 1) +
  theme_minimal() +
  labs(title = "Press Freedom Score Mosaic by Zone")

p
```

## Beautification

```{r}
library(ggview)   # gives us "correct" graph dimensions -> use canvas()
library(ggtext)   # for textbox_simple
library(showtext)

# ---------------- set canvas parameters
our_height <- 8
our_width  <- 8
our_unit   <- "in"
our_dpi    <- 300

#----------------- put self-speaking titles 
our_title <- "<span style = 'font-size:12pt'>*Day 06 - Reporters Without Borders -*</span> <br>**Press Freedom Score by Zone**"
our_subtitle <- "For each zone, countries are ranked in accordance with their annual press freedom score. Data completeness varies across the years for all zones."
our_caption <- "Performance Section DECEA and PRU EUROCONTROL<br>Data: Reporters Without Borders <br>Data preparation: Peter Baumgartner (raw.githubusercontent.com/petzi53/rwb-book/master/data/chap011/rwb/rwb1.rds)"

day06 <- p +
  
  labs(x = NULL, y = NULL,
       title = our_title     # control with theme & textboxsimple
       ,subtitle = our_subtitle
       ,caption = our_caption
       ,fill = "press \nfreedom \nscore"
       ) +
  theme_minimal(base_size = 14) +
  theme(
     plot.title = element_textbox_simple(margin = margin(b = 10))
    ,plot.title.position = "plot"
    ,plot.subtitle = element_textbox_simple()
    ,plot.caption = element_textbox_simple(
       size   = rel(0.7)
      ,margin = margin(t = 10)
    )
    ,plot.caption.position = "plot"
    
    ,legend.position = c(0.98, 1.2)             # near top-right
    ,legend.justification = c("right", "top")
    ,legend.direction = "horizontal",
    ,legend.text  = element_text(size = rel(0.7))
    ,legend.title = element_text(size = rel(0.8)) 
  )

day06 <- day06 +
  canvas(width = our_width, height = our_height, units = our_unit, dpi = our_dpi)

day06
```

save out the final plot

what_day     = "06"     # correct this to the right day! = do not overwrite other files
our_filename = paste0("./figures/day", what_day,".png")  
our_filename

ggsave(our_filename
      , width = our_width
      , height = our_height
      , unit = our_unit
      , dpi = our_dpi)

