---
title: "Day 07 - multi-scale"
format: html
---

## Idea

show distribution per region --> scale ~ variation, scale ~ region

work with a subset of the KPIs (default to traffic and punctuality)

## distribution ~ ridgeline

### data preparation

```{r}
# load data
punc <- fs::dir_ls("data", regexp = "-punc-2024") |> 
  purrr::map(.f = ~ readr::read_csv(.x, show_col_types = FALSE)) |> 
  dplyr::bind_rows()

# label regions
punc <- punc |> mutate(REGION = case_when(
   grepl("^S", APT)   ~ "BRA"
  ,grepl("^E|L", APT) ~ "EUR"
  ,.default = NA
))


# helper functions to aggregate the data
add_dly_early_late_groupings <- function(.punc_df){
  df <- .punc_df |>  
    dplyr::mutate(
      EARLY        = rowSums(across(.cols = `(-INF,-60]`:`(-20,-15]`), na.rm = TRUE)
      ,EARLY_M15M05 = (`(-15,-10]` + `(-10,-5]`)
      ,EARLY_M05M00 =  `(-5,0]`                 
      ,LATE_P00P05  =  `(0,5)`                   
      ,LATE_P05P15  = (`[5,10)` + `[10,15)`)     
      ,LATE         = rowSums(across(.cols = `[15,20)`:`[60,INF)`), na.rm = TRUE) 
      ,WITHIN_M05P05= (`(-5,0]` + `(0,5)`)
      ,WITHIN_M15P15= (`(-15,-10]`+`(-10,-5]`+`(-5,0]`+`(0,5)`+`[5,10)`+`[10,15)`)
    )
}

daily_punc <- punc |> 
  add_dly_early_late_groupings() |> 
  dplyr::select(APT, DATE, PHASE, REGION, EARLY:LATE) |> 
  dplyr::group_by(REGION, APT, DATE, PHASE) |> 
  dplyr::mutate(N_VALID = rowSums(dplyr::across(EARLY:LATE)))  |> 
  dplyr::ungroup()

daily_punc <- daily_punc |> 
  dplyr::mutate(PUNC = rowSums(dplyr::across(EARLY_M15M05:LATE_P05P15)) / N_VALID)
```


df  # one row per airport per time-unit (day/week/year)
# region   : "BRA" / "EUR"
# airport  : airport code (ICAO)
# period   : Date (daily), or week start date, or year (numeric)
# kpi      : numeric value (e.g., arrival punctuality % or delay minutes)
# movements: optional numeric weights (e.g., flights)




### exploratory graph - ridgeline

```{r}
library(ggplot2)
```



```{r}
daily_punc |> # dplyr::filter(PHASE == "ARR") |> 
  ggplot(aes(x = PUNC, colour = PHASE, weight = N_VALID)) +
  geom_density(adjust = 1) +
  facet_wrap(~ APT, scales = "free_y")
```

## using ggridge

```{r}
library(ggridges)

daily_punc |> 
ggplot(aes(x = PUNC, y = APT, fill = PHASE, weight = N_VALID)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  facet_wrap(~ REGION, scale = "free_y")
```  

```{r}
daily_punc |>   
 ggplot(aes(x = PUNC, y = APT, fill = PHASE, weight = N_VALID)) +
  geom_density_ridges(scale = 1) +
  facet_wrap(~ REGION, scale = "free_y")
```

## using ggdist

```{r}
library(ggdist)

daily_punc |> 
ggplot(aes(y = APT, x = PUNC, fill = PHASE)) +
  ggdist::stat_halfeye() +
  facet_wrap(~ REGION, scale = "free_y")
```

```{r}
daily_punc |> 
ggplot(aes(y = APT, x = PUNC, fill = PHASE)) +
  ggdist::stat_halfeye(alpha = 0.6, position = "identity") +
  facet_wrap(~REGION, scale = "free_y")
```