---
title: "Day07 - multi-scale"
format: html
---

This is to download the 2023 data from the github repo.

Note: removed {r} to deactivate the following, but keep the code for reference.

```{}
bra_punc_url <- "https://raw.githubusercontent.com/euctrl-pru/international-BRA-EUR-2025/refs/heads/main/data/BRA-punc.csv"

study_years <- 2023

bra_punc <- readr::read_csv(bra_punc_url, show_col_types = FALSE) |> 
# correct (minimal impact) renamed colums
  dplyr::rename("(-20,-15]" = "(-20,-16]","(-15,-10]" = "(-16,-10]") |> 
  dplyr::filter(lubridate::year(DATE) %in% study_years)

out_name <- paste0("./data/BRA-punc-", study_years, ".csv")
#bra_punc |> readr::write_csv(out_name)

# European data
# ----- note: there is a separate file for LPPT as we changed airports in 2025


#url_2024 <- "https://raw.githubusercontent.com/euctrl-pru/international-BRA-EUR-2025/main/data/PBWG-EUR-PUNC-2024.csv"
url_2023 <- "https://raw.githubusercontent.com/euctrl-pru/international-BRA-EUR-2025/main/data/PBWG-EUR-PUNC-2023.csv"
url_lppt <- "https://raw.githubusercontent.com/euctrl-pru/international-BRA-EUR-2025/main/data/PBWG-EUR-PUNC-LPPT-2019-2024.csv"

eur_punc      <- readr::read_csv(url_2023, show_col_types = FALSE)
eur_punc_lppt <- readr::read_csv(url_lppt, show_col_types = FALSE)

eur_punc <- eur_punc |> dplyr::bind_rows(eur_punc_lppt) |> 
  dplyr::filter(lubridate::year(DATE) == study_years) |> 
  # trim to 10 comparison airports of last year
  dplyr::filter(! ICAO %in% c("LIRF","LTFM","LEPA","EFHK"))

out_name <- paste0("./data/EUR-punc-", study_years, ".csv")
eur_punc |> dplyr::rename(APT = ICAO) |> readr::write_csv(out_name)
```

# Data 

* go to data folder and check for files with PUNC
* loop over the list and read all in

```{r}
# load data
punc <- fs::dir_ls("data", regexp = "-punc-202[3|4]") |> 
  purrr::map(.f = ~ readr::read_csv(.x, show_col_types = FALSE)) |> 
  dplyr::bind_rows() |> 
  #-------- add region label -------------
  dplyr::mutate(REGION = case_when(
   grepl("^S", APT)   ~ "BRA"
  ,grepl("^E|L", APT) ~ "EUR"
  ,.default = NA
))

# check what we have
punc
```

The data has both arrival and departure punctuality in the given slots!

If you wanto to work with the arrival punctuality:

arr_punc <- punc |> dplyr::filter(PHASE == "ARR")

```{r}
# dependent on aggregation level 
punc_annual <- punc |> 
  # set the grouping of aggregation
  # here we have per airport, per phase (if not filtered before), for each YEAR, each REGION
  group_by(APT, PHASE, YEAR = lubridate::year(DATE), REGION) |> 
  reframe(across(.cols = `(-INF,-60]`:`[60,INF)`, .fns = sum))

punc_annual
```

```{r}
# helper function to add the delay groupings in PBWG convention
add_dly_early_late_groupings <- function(.punc_df){
  df <- .punc_df |>  
    dplyr::mutate(
      EARLY        = rowSums(across(.cols = `(-INF,-60]`:`(-20,-15]`), na.rm = TRUE)
      ,EARLY_M15M05 = (`(-15,-10]` + `(-10,-5]`)
      ,EARLY_M05M00 =  `(-5,0]`                 
      ,LATE_P00P05  =  `(0,5)`                   
      ,LATE_P05P15  = (`[5,10)` + `[10,15)`)     
      ,LATE         = rowSums(across(.cols = `[15,20)`:`[60,INF)`), na.rm = TRUE) 
      ,WITHIN_M05P05= (`(-5,0]` + `(0,5)`)
      ,WITHIN_M15P15= (`(-15,-10]`+`(-10,-5]`+`(-5,0]`+`(0,5)`+`[5,10)`+`[10,15)`)
    )
}

# to calculate the N_VALID we sum over all countes
add_N_VALID <- function(.punc_df_with_groupings){
  df <- .punc_df_with_groupings |> 
    dplyr::mutate(N_VALID = rowSums(across(EARLY:LATE)))
  df
}

# ========== example
piep <- punc_annual |> 
  add_dly_early_late_groupings() |> 
  add_N_VALID()

piep |> dplyr::glimpse()
```

```{r}
punctuality_as_percentage <- function(df_with_punc_and_groups){
  df <- df_with_punc_and_groups |> 
    dplyr::mutate(dplyr::across(.cols = EARLY:WITHIN_M15P15, .fns = ~ .x / N_VALID))
  df
}

piep <- punc_annual |> 
  add_dly_early_late_groupings() |> 
  add_N_VALID() |> 
  punctuality_as_percentage()

piep |> dplyr::glimpse()


aptview <- piep |>
     group_by(PHASE , YEAR , REGION , APT) |>
     reframe(CHECK = n()
             , across(`(-INF,-60]`:`[60,INF)`, .fns = ~ sum(.x) )
     ) |>
     add_dly_early_late_groupings2() |>
     mutate(N = rowSums(across(`(-INF,-60]`:`[60,INF)`)) )|>
     mutate(across(.cols = EARLY:WITHIN_M15P15,.fns =  ~ .x / N) ) |> 
     select(PHASE , WITHIN_M15P15 ,YEAR , REGION , APT) |> print()

netview <- piep |>
     group_by(PHASE , YEAR , REGION) |>
     reframe(CHECK = n()
             , across(`(-INF,-60]`:`[60,INF)`, .fns = ~ sum(.x) )
     ) |>
     add_dly_early_late_groupings2() |>
     mutate(N = rowSums(across(`(-INF,-60]`:`[60,INF)`)) )|>
     mutate(across(.cols = EARLY:WITHIN_M15P15,.fns =  ~ .x / N) ) |> 
     select(PHASE , WITHIN_M15P15 ,YEAR , REGION) |> print()


```


### Bene ggplot ###


```{r}

library(tidyverse)
library(patchwork)
library(scales)

# ── Colour palette & shared theme ────────────────────────────────────────────
pal_region  <- c("BRA" = "#009C3B", "EUR" = "#2B4EAE")
pal_year    <- c("2023" = "#888888", "2024" = "#111111")
THRESHOLD   <- 80   # % on-time reference line

theme_day07 <- function(base_size = 11) {
  theme_minimal(base_size = base_size) +
    theme(
      text                = element_text(family = "sans", color = "#1a1a2e"),
      plot.title          = element_text(size = rel(1.05), face = "bold",
                                         margin = margin(b = 4)),
      plot.subtitle       = element_text(size = rel(0.82), color = "#555577",
                                         margin = margin(b = 8)),
      axis.text           = element_text(size = rel(0.82)),
      axis.title          = element_text(size = rel(0.85)),
      panel.grid.minor    = element_blank(),
      panel.grid.major.y  = element_line(color = "#e8e8f0", linewidth = 0.4),
      panel.grid.major.x  = element_line(color = "#e8e8f0", linewidth = 0.4),
      strip.text          = element_text(face = "bold", size = rel(0.88),
                                         color = "#1a1a2e"),
      strip.background    = element_rect(fill = "#f0f0f8", color = NA),
      legend.position     = "bottom",
      legend.title        = element_text(size = rel(0.82), face = "bold"),
      legend.text         = element_text(size = rel(0.80)),
      plot.background     = element_rect(fill = "#fafafa", color = NA),
      panel.background    = element_rect(fill = "#fafafa", color = NA)
    )
}


# ── Level 2: Airport-level summary ───────────────────────────────────────────
set.seed(42)

scopetable_raw <- readr::read_csv(here::here("data", "study_airports_BRA_EUR.csv"))

scopetable <- bind_rows(
  scopetable_raw |>
    transmute(
      REGION       = "BRA",
      AIRPORT_NAME = str_remove(Brazil, "\\s*\\(\\w+\\)$") |> str_trim(),
      AIRPORT_ICAO = str_extract(Brazil, "(?<=\\()\\w+(?=\\))")
    ),
  scopetable_raw |>
    transmute(
      REGION       = "EUR",
      AIRPORT_NAME = str_remove(Europe, "\\s*\\(\\w+\\)$") |> str_trim(),
      AIRPORT_ICAO = str_extract(Europe, "(?<=\\()\\w+(?=\\))")
    )
)


# ── Network aggregates ────────────────────────────────────────────────────────
df_network <- netview |> 
  filter(PHASE == "ARR") |>
  mutate(
    PCT_ONTIME = WITHIN_M15P15 * 100,
    YEAR       = factor(YEAR)
  ) |>
  select(REGION, YEAR, PCT_ONTIME)

theme_day07 <- function(base_size = 11) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title         = element_text(size = rel(1.05), face = "bold",
                                        margin = margin(b = 4)),
      plot.subtitle      = element_text(size = rel(0.82), color = "#555577",
                                        margin = margin(b = 8)),
      axis.text          = element_text(size = rel(0.82)),
      axis.title         = element_text(size = rel(0.85)),
      panel.grid.minor   = element_blank(),
      panel.grid.major.y = element_line(color = "#e8e8f0", linewidth = 0.4),
      panel.grid.major.x = element_line(color = "#e8e8f0", linewidth = 0.4),
      strip.text         = element_text(face = "bold", size = rel(0.88)),
      strip.background   = element_rect(fill = "#f0f0f8", color = NA),
      legend.position    = "bottom",
      plot.background    = element_rect(fill = "#fafafa", color = NA),
      panel.background   = element_rect(fill = "#fafafa", color = NA)
    )
}

# ── P1: Network ───────────────────────────────────────────────────────────────
p1 <- ggplot(df_network,
             aes(x = PCT_ONTIME, y = REGION,
                 color = REGION, group = REGION)) +
  geom_line(color = "#ccccdd", linewidth = 1.8, lineend = "round") +
  geom_point(aes(shape = YEAR), size = 4) +
  ggrepel::geom_text_repel(
    aes(label = sprintf("%.1f%%", PCT_ONTIME)),
    nudge_y    = 0.25,
    size       = 3.2,
    fontface   = "bold",
    direction  = "x",      
    segment.color = NA      
  ) +
  scale_color_manual(values = pal_region, guide = "none") +
  scale_shape_manual(values = c("2023" = 21, "2024" = 19), name = "Year") +
    scale_x_continuous(
    limits = c(50, 80),                             
    labels = scales::label_number(suffix = "%")
  ) +
  scale_y_discrete(labels = c("BRA" = "Brazil", "EUR" = "Europe")) +
  labs(
    title    = "① Network — Arrival Punctuality" ,
    subtitle = "Network-wide % on-time arrivals · 2023 vs 2024",
    x = "% On-Time Arrivals", y = NULL
  ) +
  theme_day07()

# ── P2: Airport slope ─────────────────────────────────────────────────────────

df_airport <- aptview |>
  filter(PHASE == "ARR") |>
  mutate(
    PCT_ONTIME   = WITHIN_M15P15 * 100,
    YEAR         = factor(YEAR),
    AIRPORT_ICAO = APT
  ) |>
  left_join(
    scopetable |> select(AIRPORT_ICAO, AIRPORT_NAME, REGION),
    by = c("AIRPORT_ICAO", "REGION")
  ) |>
  select(REGION, AIRPORT_ICAO, AIRPORT_NAME, YEAR, PCT_ONTIME)


df_airport_wide <- df_airport |>
  select(REGION, AIRPORT_ICAO, AIRPORT_NAME, YEAR, PCT_ONTIME) |>
  pivot_wider(names_from = YEAR, values_from = PCT_ONTIME,
              names_prefix = "Y") |>
  group_by(REGION) |>
  mutate(AIRPORT_FCT = fct_reorder(AIRPORT_ICAO, Y2024)) |>
  ungroup()

df_airport_long <- df_airport_wide |>
  pivot_longer(c(Y2023, Y2024), names_to = "YEAR", values_to = "PCT_ONTIME") |>
  mutate(YEAR = recode(YEAR, Y2023 = "2023", Y2024 = "2024"))

p2 <- ggplot(df_airport_wide, aes(y = AIRPORT_FCT)) +
  geom_segment(aes(x = Y2023, xend = Y2024,
                   yend = AIRPORT_FCT, color = REGION),
               linewidth = 1.2, alpha = 0.45,
               arrow = arrow(length = unit(0.12, "cm"), type = "closed")) +
  geom_point(data = df_airport_long |> filter(YEAR == "2023"),
             aes(x = PCT_ONTIME, color = REGION),
             size = 2.5, shape = 21, fill = "white", stroke = 1.2) +
  geom_point(data = df_airport_long |> filter(YEAR == "2024"),
             aes(x = PCT_ONTIME, color = REGION),
             size = 2.5, shape = 16) +
  facet_wrap(~REGION, scales = "free_y", ncol = 2,
             labeller = labeller(REGION = c("BRA" = "Brazil",
                                            "EUR" = "Europe"))) +
  scale_color_manual(values = pal_region, guide = "none") +
  scale_x_continuous(limits = c(40, 80),
                     labels = scales::label_number(suffix = "%")) +
  labs(
    title    = "② Airport View" ,
    subtitle = "Each airport: ○ 2023 → ● 2024",
    x = "% On-Time Arrivals", y = NULL
  ) +
  theme_day07() +
  theme(axis.text.y = element_text(size = 8))

# ── P3: Distribution ──────────────────────────────────────────────────────────
p3 <- ggplot(df_airport,
             aes(x = PCT_ONTIME, y = REGION,
                 fill = REGION, color = REGION)) +
  geom_boxplot(aes(group = interaction(REGION, YEAR)),
               position = position_dodge(width = 0.55),
               width = 0.35, alpha = 0.25,
               outlier.shape = NA, linewidth = 0.5) +
  geom_jitter(aes(shape = YEAR, group = interaction(REGION, YEAR)),
              position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.55),
              size = 2.2, alpha = 0.8) +
  scale_fill_manual(values  = pal_region, guide = "none") +
  scale_color_manual(values = pal_region, guide = "none") +
  scale_shape_manual(values = c("2023" = 21, "2024" = 19), name = "Year") +
  scale_x_continuous(limits = c(50, 70),
                     labels = scales::label_number(suffix = "%")) +
  scale_y_discrete(labels = c("BRA" = "Brazil", "EUR" = "Europe")) +
  labs(
    title    = "③ Distribution — Spread & Tails Across Airports",
    subtitle = "Box: IQR · Points: individual airports · ○ 2023 ● 2024",
    x = "% On-Time Arrivals (airport level)", y = NULL
  ) +
  theme_day07()

# ── Composição final ──────────────────────────────────────────────────────────
final_plot <- (p1 / p2 / p3) +
  plot_annotation(
    title    = "BRA–EUR Punctuality: A Multiscale View  |  2023–2024",
    subtitle = paste0(
      "Same KPI · Three levels of zoom — Network → Airport → Distribution\n"
    ),
    caption  = "Day 07 · #30DayChartChallenge · DECEA Performance Section & EUROCONTROL PRU",
    theme = theme(
      plot.title    = element_text(size = 15, face = "bold", color = "#1a1a2e"),
      plot.subtitle = element_text(size = 9,  color = "#666688"),
      plot.caption  = element_text(size = 8,  color = "#999999", hjust = 1),
      plot.background = element_rect(fill = "#fafafa", color = NA)
    )
  ) +
  plot_layout(heights = c(1, 2.2, 1.4), guides = "collect") &
  theme(legend.position = "bottom")


ggsave(
  filename = "./figures/day07_multi-scale.png",
  plot     = final_plot,
  width    = 12,
  height   = 13,
  dpi      = 200,
  bg       = "#fafafa"
)

message("✅  Saved: day07_distributions.png")



```