---
title: "Day 01 - part-to-whole"
format: html
---

## Topic / Prompt

part-to-whole

## Ideas

Classic presentation of a part-to-whole would be a grouped barchart.
It would be good, if we can present the scale difference between BRA and EUR.

* shares can be presented as sunburst (hierarchical donut) --> geom_arc or specific sunburst package (tbd)
* bi-directional waffle --> separate for BRA and EUR, then "patched" with patchwork
* mosaic


This is Day 01 - we have no prep. Thus, annotations (e.g. logo) can come with the next submissions.    
What would be cool is to have a background chart showing Brazil and Europe which we could add as background.



## Trials

```{r}
#| message: false
library(tidyverse)

# play data
df <- tibble(
  region = rep(c("Europe", "Brazil"), each = 4),
  group  = rep(c("Study 12", "Study 12", "Others", "Others"), 2),
  dest   = rep(c("Partner", "Elsewhere", "Partner", "Elsewhere"), 2),
  count  = c(1500, 50000, 200, 150000,  # Europe (Huge volume, small % to BR)
             1200, 8000,  100, 15000)   # Brazil (Lower volume, higher % to EUR)
)
```


## Sunburst

```{r}
library(ggforce)

# Data prep for sunburst involves calculating cumulative shares
df_sun <- df |>
  group_by(region) |>
  mutate(share = count / sum(count),
         end = cumsum(share) * 2 * pi,
         start = lag(end, default = 0))

ggplot(df_sun) +
  geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0.6, r = 1, 
                   start = start, end = end, fill = group)
               , stat = "identity") +
  geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 1, r = 1.4, 
                   start = start, end = end, fill = dest)
               , stat = "identity") +
  facet_wrap(~region) +
  coord_fixed() +
  theme_void()
```


## Waffle

```{r}
# install.packages("devtools")
# devtools::install_github("liamgilbey/ggwaffle")
library(ggwaffle)


# Scaled for 1 square = 500 flights
# df_waffle <- df |> 
#   mutate(scaled_count = ceiling(count / 500)) 

# Brazil Waffle
tmp_bra <- df |>
    filter(region == "Brazil") |>
    mutate(
        class = paste0(group, "-", dest),
        scaled_count = round(count / 1000)
    ) |>
    tidyr::uncount(scaled_count)

p_bra <- tmp_bra  |> 
  waffle_iron(aes_d(group = class)) |>
  
    ggplot(aes(x, y, fill = group)) +
    geom_waffle() +
    coord_equal() +
    scale_fill_brewer(palette = "Set2", name = "Class") +
    theme_waffle() + 
  labs(subtitle = "Brazil")



# Europe Waffle
tmp_eur <- df |>
    filter(region == "Europe") |>
    mutate(
        class = paste0(group, "-", dest),
        scaled_count = round(count / 1000)
    ) |>
    tidyr::uncount(scaled_count)

p_eur <- tmp_eur  |> 
  waffle_iron(aes_d(group = class)) |>
  
    ggplot(aes(x, y, fill = group)) +
    geom_waffle() +
    coord_equal() +
    scale_fill_brewer(palette = "Set2", name = "Class") +
    theme_waffle() + 
  labs(subtitle = "Europe")


p_bra
p_eur
```

:::{.callout-note}
Does not look the way I want it ... not sure how to combine.
:::

