---
title: "Day 08 - circular"
format: html
---

Circular “route wheel” for international flows
•	What it could show: distribution of international traffic flows to/from Brazil vs Europe by world region, arranged around a circle.
•	Data to tap: OD or entry/exit region mapping; aggregate by destination region and weight by flights/ASK.
•	Discussion hook: link traffic structure to performance (do long-haul shares correlate with punctuality tails?).
•	ggplot2 angle: chord-like effect using ggforce::geom_arc_bar / geom_bezier; more experimental.

```{r}

# get is from local file - only works on RQ's comp
path_one_up <- here::here() |> dirname()
path_nm_flt <- here::here(path_one_up, "__DATA","NM-flight-table")

nm_flt_2025 <- "NM-flt-2025.zip"

nm_flt <- read_zip_content(path_nm_flt, nm_flt_2025) |> 
  dplyr::bind_rows()
```

```{r}
# study airports
bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```

```{r}
conns <- nm_flt |> 
    dplyr::group_by(ADEP, ADES) |> 
    dplyr::reframe(DEPS = n()) |> 
    dplyr::mutate(
        ADEP_REG = dplyr::case_when(
             ADEP == "AFIL"  ~ "EUR"    # air filed FPLs only regional
            ,grepl(pattern = "^[E|L|BI]", ADEP) ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADEP) ~ "BRA"
            , .default = "INTNL"
            )
        ,ADES_REG = dplyr::case_when(
             grepl(pattern = "^[E|L|BI]", ADES)         ~ "EUR"
            ,grepl(pattern = "^S[B|D|I|J|N|S|W]", ADES) ~ "BRA"
            , .default = "INTNL"
        )
    )

conns |> group_by(ADEP_REG, ADES_REG) |> reframe(COMBOS = dplyr::n(), DEPS = sum(DEPS))
```

```{r}
eur_bra <- conns |> 
  dplyr::filter(ADEP_REG == "EUR", ADES_REG == "BRA") 
eur_bra <- eur_bra |> 
  dplyr::mutate(
      ADEP_STUDY = ifelse(ADEP %in% eur_apts, ADEP, "OTHER_EUR")
    , ADES_STUDY = ifelse(ADES %in% bra_apts, ADES, "OTHER_BRA")
    ) |> 
  dplyr::group_by(ADEP_STUDY, ADES_STUDY) |> 
  dplyr::reframe(DEPS = sum(DEPS))

eur_bra |> readr::write_csv("./data/BRA-EUR-aerodrome-pair-connections.csv")

eur_bra
```

```{r}
eur_order <- c(eur_apts, "OTHER_EUR")
bra_order <- c(bra_apts, "OTHER_BRA")

sector_order <- c(eur_order, bra_order)
```

```{r}
library(circlize)

wheel_data <- eur_bra |> 
  dplyr::rename(ADEP_GRP = ADEP_STUDY, ADES_GRP=ADES_STUDY, FLTS = DEPS)

circos.clear()

circos.par(
  start.degree = 90,
  gap.after = c(rep(2, length(eur_order) - 1), 10,
                rep(2, length(bra_order) - 1), 10)
)

chordDiagram(
  x = wheel_data,
  order = sector_order,
  grid.col = c(
    setNames(rep("#4472C4", length(eur_order)), eur_order),
    setNames(rep("#2E6B3E", length(bra_order)), bra_order)
  ),
  transparency = 0.5,
  annotationTrack = "grid",
  preAllocateTracks = 1
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.6
    )
  },
  bg.border = NA
)

```

