---
title: "Day 10 - pop-culture"
format: html
---

NOTE OF CAUTION - THIS IS A TERRIBLE DOCUMENT AND NEEDS TO BE CLEANED UP.    
IT SIMPLY TOOK TOO LONG TO GET THE GRAPH GOING!



```{r}
#| label: setup
# key libraries
library(dplyr)
library(ggplot2)
library(ggview)

# ggview/ggsave defaults
mywidth  = 8
myheight = 6
mydpi    = 300
```

## data preparation

```{r}
# study airports
bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```

extract EUR-BRA connections

```{}
source("./R/rqutils-zip.R")
path_nm_flt <- here::here(here::here() |> dirname(), "__DATA" ,"NM-flight-table")
nm_flt_2025 <- read_zip_content(path_nm_flt, "NM-flt-2025.zip") |> 
  bind_rows()

eur_bra_conns_2025 <- nm_flt_2025 |>  
  select(ADEP, ADES, TYPE = AIRCRAFT_TYPE_ICAO_ID) |> 
  filter(ADEP %in% eur_apts, ADES %in% bra_apts)

eur_bra_conns_2025 |> write_csv("./data/BRA-EUR-aerodrome-pair-types-2025.csv")
```

```{r}
eur_bra_conns_2025 <- read_csv("./data/BRA-EUR-aerodrome-pair-types-2025.csv", show_col_types = FALSE)

top_x <- 6

fleet <- eur_bra_conns_2025 |>
  count(TYPE, name = "FLTS") |>
  arrange(desc(FLTS)) |>
  mutate(SHARE = FLTS / sum(FLTS)) |>
  slice_head(n = top_x)

# fleet ranking
falcon <- tibble(
  TYPE = "Millennium Falcon",
  FLTS = 0,
  SHARE = 1.2      # our top contender
)

fleet <- bind_rows(falcon, fleet) |> 
  arrange(desc(SHARE)) |> 
  mutate(TYPE = factor(TYPE, levels = TYPE) )

fleet <- fleet |> 
  mutate(is_falcon = TYPE == "Millennium Falcon")
```

```{r}
fleet
fleet$TYPE
```

```{r}
# fleet2 <- fleet |> 
#   mutate(TYPE = forcats::fct_recode(TYPE,
#   "A339" = "A330-900",
#   "B77W" = "777-300ER",
#   "B789" = "787-9",
#   "A359" = "A350-900"
# ))

p <- ggplot(fleet, aes(x = SHARE, y = TYPE)) +
  geom_col(fill = "grey70") +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    title = "Transatlantic Fleet Ranking",
    subtitle = "Aircraft share on Europe–Brazil routes (plus one legendary outlier) - From widebody workhorses to one iconic smuggler ship!",
    x = "Share of flights",
    y = NULL
  ) +
  theme_minimal()

p <- p +
  geom_col(
    data = subset(fleet, TYPE == "Millennium Falcon"),
    fill = "gold"
  )

p <- p +
  annotate(
  "text",
  x = 0.02,
  y = "Millennium Falcon",
  label = "Still waiting for its first scheduled service...",
  hjust = 0,
  size = 3.5
)
```

```{r}
p +
  canvas(width = mywidth, height = myheight, dpi = mydpi)
```

https://commons.wikimedia.org/wiki/File:Boeing_737-300ER_silhouette.svg
Peter James Lowden, CC BY-SA 4.0 <https://creativecommons.org/licenses/by-sa/4.0>, via Wikimedia Commons

https://commons.wikimedia.org/wiki/File:Boeing_777-200_silhouette.svg
Peter James Lowden, CC BY-SA 4.0 <https://creativecommons.org/licenses/by-sa/4.0>, via Wikimedia Commons

# ---- setup ---------------------------------------------------------------
dir.create("aircraft_silhouettes", showWarnings = FALSE)

imgs <- tibble::tribble(
  ~TYPE, ~view, ~url, ~ext,
  # A330-900
  "A339","top","https://commons.wikimedia.org/wiki/Special:FilePath/Airbus_A330_family_v1.1.png","png",
  "A339","side","https://www.norebbo.com/wp-content/uploads/2018/06/A330-900_NEO_white_sm.jpg","jpg",

  # B777-300ER
  "B77W","top","https://commons.wikimedia.org/wiki/Special:FilePath/Boeing_777-300_silhouette.svg","svg",
  "B77W","side","https://cdn.inspireuplift.com/uploads/images/seller_products/1700409973_EN-20231119-10564_Boeing777-300Silhouette5155.jpg","jpg",

  # B787-9
  "B789","top","https://jundroo.blob.core.windows.net/simpleplanes/GameData/aircraft/1323/9Cag5m-TopView.jpg","jpg",
  "B789","side","https://cdn.inspireuplift.com/uploads/images/seller_products/1700406829_AW-20231119-10565_Boeing787-9Silhouette3662.jpg","jpg",

  # A350-900
  "A359","top","https://www.clipartmax.com/png/middle/481-4817215_1000-x-407-2-airbus-a350-top-view.png","png",
  "A359","side","https://www.norebbo.com/wp-content/uploads/2013/07/A350-900_white.jpg","jpg",

  # B777-200
  "B772","top","https://commons.wikimedia.org/wiki/Special:FilePath/Boeing_777-200_silhouette.svg","svg",
  "B772","side","https://i.etsystatic.com/32129826/r/il/f90229/5982314878/il_fullxfull.5982314878_541o.jpg","jpg",

  # A330-200
  "A332","top","https://commons.wikimedia.org/wiki/Special:FilePath/Airbus_A330-200_v1.0.png","png",
  "A332","side","https://www.norebbo.com/wp-content/uploads/2014/03/A330-200_RR_white.jpg","jpg",

  # Millennium Falcon (pun)
  "MF","top","https://www.nicepng.com/png/detail/12-120090_created-with-raphal-millennium-falcon-from-top.png","png",
  "MF","side","https://i.pinimg.com/736x/e9/64/32/e964321a2aa7f5ce8c69664d72e3ec78.jpg","jpg"
)

# ---- dependencies --------------------------------------------------------
if (any(imgs$ext == "svg")) {
  if (!requireNamespace("rsvg", quietly = TRUE))
    stop("Please install rsvg: install.packages('rsvg')")
}

# ---- download + convert --------------------------------------------------
index <- list()

for (i in seq_len(nrow(imgs))) {
  row <- imgs[i,]
  out_dir <- file.path("aircraft_silhouettes", row$TYPE)
  dir.create(out_dir, showWarnings = FALSE)

  target_png <- file.path(out_dir, paste0(row$TYPE, "_", row$view, ".png"))

  if (row$ext == "svg") {
    tmp_svg <- tempfile(fileext = ".svg")
    download.file(row$url, tmp_svg, mode = "wb", quiet = TRUE)
    rsvg::rsvg_png(tmp_svg, target_png, width = 2000)
    unlink(tmp_svg)
  } else {
    download.file(row$url, target_png, mode = "wb", quiet = TRUE)
  }

  index[[i]] <- data.frame(
    TYPE = row$TYPE,
    view = row$view,
    file = file.path(row$TYPE, basename(target_png))
  )
}


<!-- imgs <- tribble( -->
<!--   ~TYPE, ~view, ~url, ~ext, -->

<!--   # --- FIXED / CLEAN ASSETS ONLY -------------------------------------- -->

<!--   # Boeing 787-9 (clean side view) -->
<!--   "B789","side","https://commons.wikimedia.org/wiki/Special:FilePath/Boeing_787-9_silhouette.svg","svg", -->

<!--   # Airbus A350-900 (clean top view) -->
<!--   "A359","top","https://upload.wikimedia.org/wikipedia/commons/3/3c/Airbus_A350_top_view.svg","svg", -->

<!--   # Airbus A330-200 (clean top view) -->
<!--   "A332","top","https://upload.wikimedia.org/wikipedia/commons/6/6e/Airbus_A330-200_top_view.svg","svg", -->

<!--   # Airbus A330-900 (clean top view) -->
<!--   "A339","top","https://upload.wikimedia.org/wikipedia/commons/8/82/Airbus_A330neo_top_view.svg","svg" -->
<!-- ) -->

if (!requireNamespace("rsvg", quietly = TRUE))
  stop("Please install rsvg: install.packages('rsvg')")

index <- list()


index <- do.call(rbind, index)

write.csv(index, "aircraft_silhouettes/index.csv", row.names = FALSE)



library(tibble)

dir.create("aircraft_silhouettes", showWarnings = FALSE)

imgs <- tribble(
  ~TYPE, ~view, ~url, ~ext,

  # --- FIXED / CLEAN ASSETS ONLY --------------------------------------

  # Boeing 787-9 (clean side view)
  "B789","side","https://commons.wikimedia.org/wiki/Special:FilePath/Boeing_787-9_silhouette.svg","svg",

  # Airbus A350-900 (clean top view)
  "A359","top","https://upload.wikimedia.org/wikipedia/commons/3/3c/Airbus_A350_top_view.svg","svg",

  # Airbus A330-200 (clean top view)
  "A332","top","https://upload.wikimedia.org/wikipedia/commons/6/6e/Airbus_A330-200_top_view.svg","svg",

  # Airbus A330-900 (clean top view)
  "A339","top","https://upload.wikimedia.org/wikipedia/commons/8/82/Airbus_A330neo_top_view.svg","svg"
)

if (!requireNamespace("rsvg", quietly = TRUE))
  stop("Please install rsvg: install.packages('rsvg')")

index <- list()

for (i in seq_len(nrow(imgs))) {
  row <- imgs[i,]
  out_dir <- file.path("aircraft_silhouettes", row$TYPE)
  dir.create(out_dir, showWarnings = FALSE)

  out_png <- file.path(out_dir, paste0(row$TYPE, "_", row$view, ".png"))

  tmp <- tempfile(fileext = ".svg")
  download.file(row$url, tmp, mode = "wb", quiet = TRUE)
  rsvg::rsvg_png(tmp, out_png, width = 2000)
  unlink(tmp)

  index[[i]] <- data.frame(
    TYPE = row$TYPE,
    view = row$view,
    file = file.path(row$TYPE, basename(out_png))
  )
}

write.csv(do.call(rbind, index),
          "aircraft_silhouettes/index_updates_clean.csv",
          row.names = FALSE)

message("✅ Clean silhouettes downloaded (gear-up, no watermark).")

```{r}

library(ggimage)

idx <- read.csv("aircraft_silhouettes/index.csv")
idx$path <- file.path("aircraft_silhouettes", idx$file)

idx <- idx |> mutate(
  x = row_number(), y = row_number()
)

ggplot(idx, aes(x, y)) +
  geom_image(aes(image = path), size = 0.08) +
  coord_equal()

```
```{r}
library(tidyverse)

# Create a simplified Millennium Falcon-inspired silhouette
# This is a stylized representation for educational/research purposes
create_falcon_silhouette <- function(output_path = "millennium_falcon_silhouette.png") {
  
  # Define the iconic saucer shape with mandibles
  t <- seq(0, 2*pi, length.out = 100)
  
  # Main saucer body (circular)
  saucer <- tibble(
    x = cos(t) * 1,
    y = sin(t) * 0.95,
    part = "saucer"
  )
  
  # Offset cockpit (the distinctive side-mounted cockpit)
  cockpit <- tibble(
    x = cos(t) * 0.25 + 0.6,
    y = sin(t) * 0.25 + 0.3,
    part = "cockpit"
  )
  
  # Mandibles (the iconic "fork" at the front)
  # Right mandible
  mandible_r <- tibble(
    x = c(0.85, 1.3, 1.35, 1.3, 0.9),
    y = c(0.15, 0.25, 0.3, 0.35, 0.25),
    part = "mandible_r"
  )
  
  # Left mandible
  mandible_l <- tibble(
    x = c(0.85, 1.3, 1.35, 1.3, 0.9),
    y = c(-0.15, -0.25, -0.3, -0.35, -0.25),
    part = "mandible_l"
  )
  
  # Rear engine section (notch in the back)
  engine_notch <- tibble(
    x = c(-1, -1.15, -1.15, -1),
    y = c(0.3, 0.25, -0.25, -0.3),
    part = "engine"
  )
  
  # Combine all parts
  falcon <- bind_rows(
    saucer,
    cockpit,
    mandible_r,
    mandible_l,
    engine_notch
  )
  
  # Create the plot
  p <- ggplot(falcon, aes(x = x, y = y, group = part)) +
    geom_polygon(fill = "black", color = "black") +
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA)
    )
  
  # Save as PNG with transparent background
  ggsave(
    output_path,
    plot = p,
    width = 4,
    height = 4,
    dpi = 300,
    bg = "transparent"
  )
  
  return(p)
}

# Create the silhouette
create_falcon_silhouette("falcon_silhouette.png")
```

```{r}
library(tidyverse)

# Create a more accurate Millennium Falcon-inspired silhouette
create_falcon_silhouette <- function(output_path = "millennium_falcon_silhouette.png") {
  
  # Main saucer body (slightly wider than tall)
  t <- seq(0, 2*pi, length.out = 100)
  saucer <- tibble(
    x = cos(t) * 1.0,
    y = sin(t) * 0.92,
    part = "saucer"
  )
  
  # Offset cockpit - the iconic side-mounted tube
  # Positioned on the starboard (right) side
  cockpit_t <- seq(0, 2*pi, length.out = 50)
  cockpit <- tibble(
    x = cos(cockpit_t) * 0.18 + 0.75,  # Offset to the right side
    y = sin(cockpit_t) * 0.20 + 0.0,   # Centered vertically
    part = "cockpit"
  )
  
  # Forward mandibles - the distinctive "fork" prongs
  # These are the two prominent forward-facing arms
  
  # Top/Starboard mandible (right side)
  mandible_r <- tibble(
    x = c(0.95, 1.4, 1.45, 1.42, 1.35, 1.25, 0.90),  # Extended forward
    y = c(0.20, 0.32, 0.35, 0.38, 0.40, 0.42, 0.30),  # Aligned with flank
    part = "mandible_r"
  )
  
  # Bottom/Port mandible (left side)  
  mandible_l <- tibble(
    x = c(0.95, 1.4, 1.45, 1.42, 1.35, 1.25, 0.90),
    y = c(-0.20, -0.32, -0.35, -0.38, -0.40, -0.42, -0.30),
    part = "mandible_l"
  )
  
  # Central gap between mandibles (the "mouth")
  # This creates the distinctive gap between the prongs
  gap_top <- tibble(
    x = c(0.95, 1.35, 1.35, 0.95),
    y = c(0.20, 0.40, 0.20, 0.15),
    part = "gap_top"
  )
  
  gap_bottom <- tibble(
    x = c(0.95, 1.35, 1.35, 0.95),
    y = c(-0.20, -0.40, -0.20, -0.15),
    part = "gap_bottom"
  )
  
  # Rear engine section - distinctive rectangular notch
  engine_notch <- tibble(
    x = c(-0.98, -1.20, -1.20, -0.98),
    y = c(0.35, 0.30, -0.30, -0.35),
    part = "engine"
  )
  
  # Sensor dish (top center detail)
  dish_t <- seq(0, 2*pi, length.out = 30)
  sensor_dish <- tibble(
    x = cos(dish_t) * 0.12 + 0.15,
    y = sin(dish_t) * 0.12 + 0.0,
    part = "dish"
  )
  
  # Combine all parts
  falcon <- bind_rows(
    saucer,
    cockpit,
    mandible_r,
    mandible_l,
    sensor_dish,
    engine_notch
  )
  
  # Create the plot with black silhouette
  p <- ggplot() +
    # Main body
    geom_polygon(
      data = falcon,
      aes(x = x, y = y, group = part),
      fill = "black",
      color = "black"
    ) +
    # Remove the gap areas (make them transparent by drawing over)
    geom_polygon(
      data = gap_top,
      aes(x = x, y = y),
      fill = "white",
      color = "white"
    ) +
    geom_polygon(
      data = gap_bottom,
      aes(x = x, y = y),
      fill = "white",
      color = "white"
    ) +
    coord_fixed(xlim = c(-1.3, 1.5), ylim = c(-1.1, 1.1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  # Save as PNG with transparent background
  ggsave(
    output_path,
    plot = p,
    width = 5,
    height = 4,
    dpi = 300,
    bg = "transparent"
  )
  
  cat("✓ Millennium Falcon silhouette saved to:", output_path, "\n")
  
  return(p)
}

# Create the improved silhouette
falcon_plot <- create_falcon_silhouette("millennium_falcon_silhouette.png")

# Preview it
print(falcon_plot)
```
```{r}
library(tidyverse)
library(magick)

# Function to download and process the Pixabay Millennium Falcon image
process_falcon_from_pixabay <- function(
    pixabay_url = "https://pixabay.com/get/g36c2e0f88e976fe86e97bad1cc9dc7ce87f07e6cb6e96ba5d4a84fba8b0de2740bcd4ce22ac50e10e86d61b9fd14b0fd_1280.jpg",
    output_path = "millennium_falcon_silhouette.png"
) {
  
  cat("Step 1: Downloading image from Pixabay...\n")
  
  # You'll need to manually download from Pixabay and provide the path
  # OR use this direct download URL if it works
  
  tryCatch({
    # Try to download directly
    download.file(pixabay_url, "temp_falcon.jpg", mode = "wb")
    img <- image_read("temp_falcon.jpg")
    cat("  ✓ Image downloaded successfully\n")
  }, error = function(e) {
    cat("  ✗ Could not download automatically. Please:\n")
    cat("    1. Visit: https://pixabay.com/illustrations/millennium-falcon-star-wars-1343464/\n")
    cat("    2. Click 'Free Download' and save the image\n")
    cat("    3. Save it as 'temp_falcon.jpg' in your working directory\n")
    cat("    4. Run this function again\n")
    stop("Manual download required")
  })
  
  cat("\nStep 2: Processing image to create silhouette...\n")
  
  # Get image info
  info <- image_info(img)
  cat("  Original dimensions:", info$width, "x", info$height, "\n")
  
  # Step 1: Convert to grayscale
  img_gray <- img %>% image_convert(colorspace = "Gray")
  cat("  ✓ Converted to grayscale\n")
  
  # Step 2: Increase contrast to separate ship from background
  img_contrast <- img_gray %>%
    image_contrast(sharpen = 10) %>%
    image_modulate(brightness = 80)
  cat("  ✓ Enhanced contrast\n")
  
  # Step 3: Create a threshold to isolate the ship
  # This turns everything dark into black and light into transparent
  img_threshold <- img_contrast %>%
    image_threshold(type = "black", threshold = "40%") %>%
    image_transparent("white")
  cat("  ✓ Applied threshold to isolate ship\n")
  
  # Step 4: Clean up edges with morphology
  img_clean <- img_threshold %>%
    image_morphology(method = "Close", kernel = "Disk:2") %>%
    image_despeckle()
  cat("  ✓ Cleaned up edges\n")
  
  # Step 5: Make it a pure black silhouette
  img_silhouette <- img_clean %>%
    image_colorize(opacity = 100, color = "black") %>%
    image_background("transparent")
  cat("  ✓ Created black silhouette\n")
  
  # Step 6: Trim excess transparent space
  img_final <- img_silhouette %>%
    image_trim() %>%
    image_border("transparent", "10x10")
  cat("  ✓ Trimmed to ship outline\n")
  
  # Step 7: Resize to standard size (width = 800px)
  img_final <- img_final %>%
    image_resize("800x")
  cat("  ✓ Resized for chart use\n")
  
  # Save the final silhouette
  image_write(img_final, output_path, format = "png")
  cat("\n✓ Falcon silhouette saved to:", output_path, "\n")
  
  # Clean up temp file
  if (file.exists("temp_falcon.jpg")) {
    file.remove("temp_falcon.jpg")
  }
  
  # Display the result
  plot(img_final)
  
  return(img_final)
}

# Alternative: Manual download version
# If the automatic download doesn't work, use this:
process_falcon_manual <- function(
    input_image = "downloaded_falcon.jpg",  # Your manually downloaded image
    output_path = "millennium_falcon_silhouette.png"
) {
  
  cat("Processing", input_image, "...\n\n")
  
  if (!file.exists(input_image)) {
    stop("Image file not found! Please download from Pixabay first.")
  }
  
  # Read the image
  img <- image_read(input_image)
  
  cat("Step 1: Converting to grayscale...\n")
  img_gray <- img %>% image_convert(colorspace = "Gray")
  
  cat("Step 2: Enhancing contrast...\n")
  img_contrast <- img_gray %>%
    image_contrast(sharpen = 10) %>%
    image_modulate(brightness = 80)
  
  cat("Step 3: Isolating the ship...\n")
  img_threshold <- img_contrast %>%
    image_threshold(type = "black", threshold = "40%") %>%
    image_transparent("white")
  
  cat("Step 4: Cleaning edges...\n")
  img_clean <- img_threshold %>%
    image_morphology(method = "Close", kernel = "Disk:2") %>%
    image_despeckle()
  
  cat("Step 5: Creating pure black silhouette...\n")
  img_silhouette <- img_clean %>%
    image_colorize(opacity = 100, color = "black") %>%
    image_background("transparent")
  
  cat("Step 6: Trimming and finalizing...\n")
  img_final <- img_silhouette %>%
    image_trim() %>%
    image_border("transparent", "10x10") %>%
    image_resize("800x")
  
  # Save
  image_write(img_final, output_path, format = "png")
  cat("\n✓ Falcon silhouette saved to:", output_path, "\n")
  
  # Display
  plot(img_final)
  
  return(img_final)
}

# TRY THIS FIRST - Automatic download and processing
# falcon <- process_falcon_from_pixabay()

# IF THAT FAILS - Manual process:
# 1. Download from https://pixabay.com/illustrations/millennium-falcon-star-wars-1343464/
# 2. Save as "downloaded_falcon.jpg" in your working directory
# 3. Run:
# falcon <- process_falcon_manual("downloaded_falcon.jpg")
```

Image by <a href="https://pixabay.com/users/jean52photosstock-14428522/?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=6054792">Jean photosstock</a> from <a href="https://pixabay.com//?utm_source=link-attribution&utm_medium=referral&utm_campaign=image&utm_content=6054792">Pixabay</a>

tie fighter
<a href="https://getdrawings.com/tie-fighter-vector">Tie Fighter Vector</a>

```{r}
library(tidyverse)
library(ggimage)
library(scales)
library(showtext)

fleet <- fleet |> mutate(is_falcon = SHARE > 1)

```

get some fancy plots

```{r}
font_add_google("Orbitron", "title")
font_add_google("Inter", "body")
showtext_auto()


img_idx <- read_csv("aircraft_silhouettes/index.csv", show_col_types = F) |>
  filter(view == "side") |>
  mutate(path = file.path("aircraft_silhouettes", file))

fleet <- fleet |> mutate(is_falcon = SHARE > 1)
fleet <- fleet |> left_join(img_idx, by = c("TYPE"))

pth_falcon <- "aircraft_silhouettes/jean52photosstock-ufo-6054792.png"
fleet <- fleet |> 
  mutate( path = ifelse(is_falcon, pth_falcon, path))

```

```{r}


# lollipop
lolli_line = 2

p <- fleet |> ggplot(aes(y = TYPE))

p <- p +
  geom_segment(
    data = fleet |> filter(!is_falcon),
    aes(x = 0, xend = SHARE, yend = TYPE),
    linewidth = lolli_line,
    colour = "grey40"
  )
#--------- add falcon stick
p <- p +
  geom_segment(
    data = fleet |> filter(is_falcon),
    aes(x = 0, xend = 1, yend = TYPE),
    linewidth = lolli_line,
    colour = "gold"
  ) +
  geom_segment(
    data = fleet |> filter(is_falcon),
    aes(x = 1, xend = SHARE, yend = TYPE),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) 


p <- p +
  geom_image( data = fleet,
    aes(x = SHARE, image = path),
    size = 0.085,
    asp = 1
  )


p <- p +
  scale_x_continuous(
    limits = c(0, 2),
    labels = percent_format(accuracy = 1),
    breaks = c(seq(0, 1, by = 0.25), 1.25),
    expand = expansion(mult = c(0, 0.05))
  )

p <- p +
  annotate(
    "text",
    x = 1.18,
    y = "Millennium Falcon",
    label = "Hyperdrive engaged",
    family = "title",
    size = 4,
    colour = "grey30"
  )

p +
  canvas(width = mywidth, height = myheight, dpi = mydpi)
```


```{r}
lolli_line = 5
dot_start  = 0.5

fleet_bkup <- fleet
fleet <- fleet |> 
  mutate(path = case_when(
    TYPE == "Millennium Falcon" ~ "./aircraft_silhouettes/jean52photosstock-ufo-6054792.png"
    , TYPE == "A339" ~ "./aircraft_silhouettes/A339_side.png"
    , TYPE == "A332" ~ "./aircraft_silhouettes/A332_side.png"
    , TYPE == "A359" ~ "./aircraft_silhouettes/A350-900.png"
    , TYPE == "B77W" ~ "./aircraft_silhouettes/B777-9_side.png"
    , TYPE == "B772" ~ "./aircraft_silhouettes/B772_side.png"
    , TYPE == "B789" ~ "./aircraft_silhouettes/B787_side.png"
    , .default = NA
  ))

fleet$TYPE <- factor(fleet$TYPE, levels = rev(unique(fleet$TYPE)))

p <- ggplot(data = fleet) +
  # --------- main stick segment portion ----------------------
  geom_segment(
    aes(  x = 0, xend = ifelse(is_falcon, dot_start, SHARE)
        , y = TYPE, yend = TYPE
        , color = is_falcon),
    linewidth = lolli_line
  ) +
  scale_color_manual(values = c("grey80", "gold"), guide = "none" ) +
  
  geom_segment(
    data = fleet |> filter(is_falcon),
    aes(x = dot_start, xend = SHARE, y = TYPE, yend = TYPE),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) +
  # -------- place images without falcon  ------------------
  geom_image(data = fleet |> filter(!is_falcon),
    aes(x = SHARE
        , y = TYPE
        , image = path),
    size = 0.4,
    asp = 1
  ) +
  
  # -------- canvas settings -----------------------------------
  scale_x_continuous(limits = c(0,1.6),expand = c(0, NA)
                     ,labels = scales::percent_format()
                     ,breaks = c(0,0.25,0.5,0.75,1)) +
  canvas(width = mywidth, height = myheight, dpi = mydpi) + NULL
  
   # -------- annotations

```

library(ggtext)
font_add_google("Space Grotesk", "space")
font_add_google("Orbitron", "orbitron")
showtext_auto()
library(ggfx)


```{r}
#install.packages("ggfx")

our_title <- "<span style='font-size:30pt'><strong>Chart-Challenge Day 10</strong></span><br>**The Workhorses of Long Distance Flying Between Europe and Brazil in 2025**"

our_subtitle <- "Share of flights by aircraft type operated."

our_caption <- "Gini coefficient: Higher Gini = more concentrated delays (few bad days account for most delay).<br>DECEA Performance Section & EUROCONTROL PRU"




p +
  with_outer_glow(
    geom_image(
      data = fleet |> filter(is_falcon),
      aes(x = SHARE, y = TYPE, image = path),
      size = 0.4
    ),
    colour = "#FFD166",
    sigma = 40
  ) +
  annotate(
    "text",
    x = 1.18,
    y = "A339",
    label = "Hyperdrive engaged \nto fast to be counted!",
    family = "title",
    size = 12,
    colour = "grey30"
  ) +

  labs(title = our_title, subtitle = our_subtitle, caption = our_caption
       ,y = NULL, x = NULL) +
  theme_clean(base_size = 24) +
  theme(
    plot.title = element_textbox_simple(
      size = rel(3)
      ,lineheight = 0.2
      ,family = "orbitron"
      ,margin = margin(b = 10)
    )
    ,plot.title.position = "plot"
    
    ,plot.subtitle = element_textbox_simple(
      size = rel(2), family = "space"
      ,margin = margin(b = 20)
    )
    
    ,axis.text = element_text(size = rel(1.5), family = "space")
  )
```

```{r}

```

library(tidyverse)
library(ggimage)
library(ggtext)
library(ggshadow)
library(scales)

```{r}

# Prepare your data with manual x-coordinates for images
fleet <- fleet |> 
  mutate(
    path = case_when(
      TYPE == "Millennium Falcon" ~ "./aircraft_silhouettes/jean52photosstock-ufo-6054792.png",
      TYPE == "A339" ~ "./aircraft_silhouettes/A339_side.png",
      TYPE == "A332" ~ "./aircraft_silhouettes/A332_side.png",
      TYPE == "A359" ~ "./aircraft_silhouettes/A350-900.png",
      TYPE == "B77W" ~ "./aircraft_silhouettes/B777-9_side.png",
      TYPE == "B772" ~ "./aircraft_silhouettes/B772_side.png",
      TYPE == "B789" ~ "./aircraft_silhouettes/B787_side.png",
      .default = NA
    ),
    # Add manual image x-position (shifted right from bar end)
    image_x = ifelse(is_falcon, SHARE + 0.18, SHARE + 0.11),
    # Add y-position as numeric for precise control
    y_numeric = as.numeric(TYPE),
    # Format share labels
    share_label = ifelse(is_falcon, "", paste0(round(SHARE * 100, 1), "%"))
  )

# Update factor levels with line break for Falcon
fleet <- fleet |> 
  mutate(TYPE = fct_recode(TYPE, "Millennium\nFalcon" = "Millennium Falcon"))

# Re-factor with correct order
fleet$TYPE <- factor(fleet$TYPE, levels = rev(unique(fleet$TYPE)))

# Main plot
p <- ggplot(data = fleet) +
  # Main stick segment
  geom_segment(
    aes(
      x = 0, 
      xend = ifelse(is_falcon, dot_start, SHARE),
      y = TYPE, 
      yend = TYPE,
      color = is_falcon
    ),
    linewidth = lolli_line
  ) +
  scale_color_manual(values = c("grey80", "gold"), guide = "none") +
  
  # Dotted portion for Falcon
  geom_segment(
    data = fleet |> filter(is_falcon),
    aes(x = dot_start, xend = SHARE, y = TYPE, yend = TYPE),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) +
  
  # Share percentage labels INSIDE bars (left-aligned)
  geom_text(
    data = fleet |> filter(!is_falcon),
    aes(x = 0.02, y = TYPE, label = share_label),
    hjust = 0,
    size = 6,
    family = "space",
    color = "grey30",
    fontface = "bold"
  ) +
  
  # Place images with manual x-position
  # geom_image(
  #   data = fleet,
  #   aes(x = image_x, y = TYPE, image = path),
  #   size = 0.4,
  #   asp = 1
  # ) +
  
  geom_image(
  data = fleet |> filter(! is_falcon),
  aes(
    x = image_x,
    y = ifelse(TYPE == "A359", y_numeric + 0.2, y_numeric),
    image = path
  ),
  size = 0.4,
  asp = 1
) +
  
  # Canvas settings
  scale_x_continuous(
    limits = c(0, 1.6),
    expand = c(0, NA),
    labels = percent_format(),
    breaks = c(0, 0.25, 0.5, 0.75, 1)
  ) +
  
  canvas(width = mywidth, height = myheight, dpi = mydpi)

# Catchy subtitles - pick your favorite!
subtitle_options <- c(
  # Option 1: Playful comparison
  "One of these aircraft is not like the others... but it makes the Kessel Run in 12 parsecs!",
  
  # Option 2: Data-focused with humor
  "Wide-body workhorses dominate the Atlantic — though one craft prefers hyperspace.",
  
  # Option 3: Professional with Easter egg
  "Long-haul aircraft dominating Brazil-Europe routes (+ one intergalactic outlier).",
  
  # Option 4: Star Wars reference
  "She may not look like much, but the Falcon's got it where it counts — along with some serious competition.",
  
  # Option 5: Simple and clean
  "The aircraft connecting two continents across the Atlantic (and one from a galaxy far, far away)."
)

# Choose your subtitle
our_subtitle <- subtitle_options[2]  # Change number to pick different option

our_title <- "<span style='font-size:30pt'><strong>Chart-Challenge Day 10</strong></span><br>**The Workhorses of Long Distance Flying Between Europe and Brazil in 2025**"

our_subtitle <- "Top share of flights by aircraft type operated. Wide-body workhorses dominate the Atlantic — though one craft prefers hyperspace. Aircraft images represent relative fleet share on Brazil-Europe routes. <br>The Millennium Falcon's hyperdrive capabilities and service frequency remain unverified by aviation authorities."

our_caption <- "credit: Millennium Falcon image by Jean photosstock from Pixabay<br>DECEA Performance Section & EUROCONTROL PRU"

# Final plot with annotations
p_final <- p +
  
  with_outer_glow(
    geom_image(
      data = fleet |> filter(is_falcon),
      aes(x = image_x, y = y_numeric, image = path),
      size = 0.4
    ),
    colour = "#FFD166",
    sigma = 40
  ) +
  
  # Main annotation for Falcon
  annotate(
    "text",
    x = 1.18,
    y = 5,  #"Millennium\nFalcon",
    label = "Hyperdrive engaged\ntoo fast to be counted!",
    family = "orbitron",
    size = 12,
    colour = "black",
    lineheight = 0.3
  ) +
  
  # Additional commentary options (uncomment ones you like):
  
  # Comment 1: Point to top aircraft
  annotate(
    "text",
    x = 0.55,
    y = "A339",
    label = "← Leading the pack:\nAirbus A330neo",
    family = "space",
    size = 12,
    #colour = "grey40",
    hjust = 0,
    lineheight = 0.3
  ) +
  
  # Comment 2: Note about Boeing/Airbus split
  annotate(
    "segment",
    x = 0.28, xend = 0.32,
    y = 6.5, yend = 6.5,
    colour = "grey50",
    linetype = "dashed"
  ) +
  annotate(
    "text",
    x = 0.6,
    y = 3,
    label = "Boeing vs Airbus:\nNearly balanced",
    family = "space",
    size = 16,
    colour = "grey40",
    hjust = 0.5,
    lineheight = 0.3
  ) +
  
  labs(
    title = our_title, 
    subtitle = our_subtitle, 
    caption = our_caption,
    y = NULL, 
    x = NULL
  ) +
  
  theme_clean(base_size = 24) +
  theme(
    plot.title = element_textbox_simple(
      size = rel(3),
      lineheight = 0.2,
      family = "orbitron",
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",
    
    plot.subtitle = element_textbox_simple(
      size = rel(2), 
      family = "space",
      margin = margin(b = 20)
      ,lineheight = 0.2
    ),
    
    plot.caption = element_textbox_simple(
      size = rel(0.7),
      family = "space",
      margin = margin(t = 15),
      lineheight = 1.2
    ),
    
    axis.text.y = element_text(
      size = rel(1.5), 
      family = "space",
      lineheight = 0.3   # tighten spacing
     # ,padding = margin(0, 0, 0, 0)
    ),
    
    axis.text.x = element_text(
      size = rel(1.5), 
      family = "space"
    ),
    
    plot.margin = margin(t = 10, r = 20, b = 10, l = 20)
  )



p_final
```

ggsave("day10-popculture.png", width = mywidth, height = myheight, dpi = mydpi)




```{r}
fleet <- fleet |>
  mutate(
    TYPE = fct_recode(TYPE, "Millennium\nFalcon" = "Millennium Falcon"),
    TYPE = factor(TYPE, levels = rev(unique(TYPE))),
    y = as.numeric(TYPE),

    path = case_when(
      TYPE == "Millennium\nFalcon" ~ "./aircraft_silhouettes/jean52photosstock-ufo-6054792.png",
      TYPE == "A339" ~ "./aircraft_silhouettes/A339_side.png",
      TYPE == "A332" ~ "./aircraft_silhouettes/A332_side.png",
      TYPE == "A359" ~ "./aircraft_silhouettes/A350-900.png",
      TYPE == "B77W" ~ "./aircraft_silhouettes/B777-9_side.png",
      TYPE == "B772" ~ "./aircraft_silhouettes/B772_side.png",
      TYPE == "B789" ~ "./aircraft_silhouettes/B787_side.png",
      TRUE ~ NA_character_
    ),

    image_x = ifelse(is_falcon, SHARE + 0.18, SHARE + 0.11),
    share_label = ifelse(is_falcon, "", paste0(round(SHARE * 100, 1), "%"))
  )
```

```{r}
p <- ggplot(fleet) +

  geom_segment(
    aes(
      x = 0,
      xend = ifelse(is_falcon, dot_start, SHARE),
      y = y,
      yend = y,
      colour = is_falcon
    ),
    linewidth = lolli_line
  ) +

  geom_segment(
    data = subset(fleet, is_falcon),
    aes(x = dot_start, xend = SHARE, y = y, yend = y),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) +

  geom_text(
    data = subset(fleet, !is_falcon),
    aes(x = 0.02, y = y, label = share_label),
    hjust = 0,
    size = 6,
    family = "space",
    colour = "grey30",
    fontface = "bold"
  ) +

  geom_image(
    data = subset(fleet, !is_falcon),
    aes(
      x = image_x,
      y = ifelse(TYPE == "A359", y + 0.2, y),
      image = path
    ),
    size = 0.4,
    asp = 1
  ) +

  scale_x_continuous(
    limits = c(0, 1.6),
    expand = c(0, NA),
    labels = percent_format(),
    breaks = c(0, 0.25, 0.5, 0.75, 1)
  ) +

  scale_y_continuous(
    breaks = fleet$y,
    labels = levels(fleet$TYPE),
    expand = expansion(add = 0.6)
  ) +

  scale_colour_manual(values = c("grey80", "gold"), guide = "none") +

  canvas(width = mywidth, height = myheight, dpi = mydpi)
```

```{r}
p_falcon <- p +

  with_outer_glow(
    geom_image(
      data = subset(fleet, is_falcon),
      aes(x = image_x, y = y, image = path),
      size = 0.4
    ),
    colour = "#FFD166",
    sigma = 40
  ) +

  annotate(
    "text",
    x = 1.18,
    y = fleet$y[fleet$is_falcon],
    label = "Hyperdrive engaged\ntoo fast to be counted!",
    family = "orbitron",
    size = 12,
    lineheight = 0.9
  )
```

```{r}
p_final <- p_falcon +
  theme_clean(base_size = 24) +
  
  # ---------------- text ----------------------------------
  labs(
    title = our_title, subtitle = our_subtitle, caption = our_caption
  )
  
  
  # ---------------- theme tweaks ---------------------------
  theme(
    plot.title = element_textbox_simple(
      size = rel(3),
      lineheight = 0.2,
      family = "orbitron",
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",
    
    plot.subtitle = element_textbox_simple(
      size = rel(2), 
      family = "space",
      margin = margin(b = 20)
      ,lineheight = 0.2
    ),
    
    plot.caption = element_textbox_simple(
      size = rel(0.7),
      family = "space",
      margin = margin(t = 15),
      lineheight = 1.2
    ),
    
    axis.text.y = element_text(
      size = rel(1.5), 
      family = "space",
      lineheight = 0.3   # tighten spacing
     # ,padding = margin(0, 0, 0, 0)
    ),
    
    axis.text.x = element_text(
      size = rel(1.5), 
      family = "space"
    ),
    
    plot.margin = margin(t = 10, r = 20, b = 10, l = 20)
  )

p_final 
```


```{r}
## ============================================================
## DATA PREP — single y-axis (numeric), stable ordering
## ============================================================

fleet <- fleet |>
  mutate(
    TYPE = fct_recode(TYPE, "Millennium\nFalcon" = "Millennium Falcon"),
    TYPE = factor(TYPE, levels = rev(unique(TYPE))),
    y = as.numeric(TYPE),

    path = case_when(
      TYPE == "Millennium\nFalcon" ~ "./aircraft_silhouettes/jean52photosstock-ufo-6054792.png",
      TYPE == "A339" ~ "./aircraft_silhouettes/A339_side.png",
      TYPE == "A332" ~ "./aircraft_silhouettes/A332_side.png",
      TYPE == "A359" ~ "./aircraft_silhouettes/A350-900.png",
      TYPE == "B77W" ~ "./aircraft_silhouettes/B777-9_side.png",
      TYPE == "B772" ~ "./aircraft_silhouettes/B772_side.png",
      TYPE == "B789" ~ "./aircraft_silhouettes/B787_side.png",
      TRUE ~ NA_character_
    ),

    image_x = ifelse(is_falcon, SHARE + 0.22, SHARE + 0.14),
    share_label = ifelse(is_falcon, "", paste0(round(SHARE * 100, 1), "%"))
  )


## ============================================================
## BASE PLOT
## ============================================================

p <- ggplot(fleet) +

  # Main lollipop sticks
  geom_segment(
    aes(
      x = 0,
      xend = ifelse(is_falcon, dot_start, SHARE),
      y = y,
      yend = y,
      colour = is_falcon
    ),
    linewidth = lolli_line
  ) +

  # Dotted Falcon extension
  geom_segment(
    data = subset(fleet, is_falcon),
    aes(x = dot_start, xend = SHARE, y = y, yend = y),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) +

  # Share labels
  geom_text(
    data = subset(fleet, !is_falcon),
    aes(x = 0.02, y = y, label = share_label),
    hjust = 0,
    size = 6,
    family = "space",
    colour = "grey30",
    fontface = "bold"
  ) +

  # Aircraft images (non-Falcon)
  geom_image(
    data = subset(fleet, !is_falcon),
    aes(
      x = image_x,
      y = ifelse(TYPE == "A359", y + 0.2, y),
      image = path
    ),
    size = 0.4,
    asp = 1
  ) +

  scale_x_continuous(
    limits = c(0, 1.6),
    expand = c(0, NA),
    labels = percent_format(),
    breaks = c(0, 0.25, 0.5, 0.75, 1)
  ) +

  scale_y_continuous(
    breaks = fleet$y,
    labels = levels(fleet$TYPE),
    expand = expansion(add = 0.6),
    trans  = "reverse"
  ) +
  
  scale_colour_manual(values = c("grey80", "gold"), guide = "none") +

  canvas(width = mywidth, height = myheight, dpi = mydpi)


## ============================================================
## FINAL ANNOTATED PLOT
## ============================================================

p_final <- p +

  # Falcon with glow
  with_outer_glow(
    geom_image(
      data = subset(fleet, is_falcon),
      aes(x = image_x, y = y, image = path),
      size = 0.4
    ),
    colour = "#FFD166",
    sigma = 40
  ) +

  # Falcon annotation
  annotate(
    "text",
    x = 1.18,
    y = fleet$y[fleet$is_falcon],
    label = "Hyperdrive engaged\ntoo fast to be counted!",
    family = "orbitron",
    size = 12,
    colour = "black",
    lineheight = 0.9
  ) +

  # Airbus leader callout
  annotate(
    "text",
    x = 0.55,
    y = fleet$y[fleet$TYPE == "A339"],
    label = "← Leading the pack:\nAirbus A330neo",
    family = "space",
    size = 12,
    hjust = 0,
    lineheight = 0.9
  ) +

  # Boeing vs Airbus note
  annotate(
    "segment",
    x = 0.28, xend = 0.32,
    y = 6.5, yend = 6.5,
    colour = "grey50",
    linetype = "dashed"
  ) +
  annotate(
    "text",
    x = 0.6,
    y = 3,
    label = "Boeing vs Airbus:\nNearly balanced",
    family = "space",
    size = 16,
    colour = "grey40",
    hjust = 0.5,
    lineheight = 0.9
  ) +

  labs(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    x = NULL,
    y = NULL
  ) +

  theme_clean(base_size = 24) +
  theme(
    plot.title = element_textbox_simple(
      size = rel(3),
      family = "orbitron",
      lineheight = 0.2,
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",

    plot.subtitle = element_textbox_simple(
      size = rel(2),
      family = "space",
      lineheight = 0.3,
      margin = margin(b = 20)
    ),

    plot.caption = element_textbox_simple(
      size = rel(0.7),
      family = "space",
      lineheight = 1.4,
      margin = margin(t = 15)
    ),

    axis.text.y = element_text(
      size = rel(1.5),
      family = "space",
      lineheight = 0.3
    ),

    axis.text.x = element_text(
      size = rel(1.5),
      family = "space"
    ),

    plot.margin = margin(t = 10, r = 20, b = 10, l = 20)
  )


## ============================================================
## RENDER
## ============================================================

p_final
```


## final version


```{r}
## ============================================================
## DATA PREP — FACTOR Y ONLY (NO NUMERIC Y ANYWHERE)
## ============================================================

fleet <- fleet |>
  mutate(
    TYPE = fct_recode(TYPE, "Millennium\nFalcon" = "Millennium Falcon"),
    TYPE = factor(TYPE, levels = rev(unique(TYPE))),

    path = case_when(
      TYPE == "Millennium\nFalcon" ~ "./aircraft_silhouettes/jean52photosstock-ufo-6054792.png",
      TYPE == "A339" ~ "./aircraft_silhouettes/A339_side.png",
      TYPE == "A332" ~ "./aircraft_silhouettes/A332_side.png",
      TYPE == "A359" ~ "./aircraft_silhouettes/A350-900.png",
      TYPE == "B77W" ~ "./aircraft_silhouettes/B777-9_side.png",
      TYPE == "B772" ~ "./aircraft_silhouettes/B772_side.png",
      TYPE == "B789" ~ "./aircraft_silhouettes/B787_side.png",
      TRUE ~ NA_character_
    ),

    image_x = ifelse(is_falcon, SHARE + 0.22, SHARE + 0.14),
    share_label = ifelse(is_falcon, "", paste0(round(SHARE * 100, 1), "%"))
  )


## ============================================================
## BASE PLOT
## ============================================================

p_base <- ggplot(fleet) +

  geom_segment(
    aes(
      x = 0,
      xend = ifelse(is_falcon, dot_start, SHARE),
      y = TYPE,
      yend = TYPE,
      colour = is_falcon
    ),
    linewidth = lolli_line
  ) +

  geom_segment(
    data = subset(fleet, is_falcon),
    aes(x = dot_start, xend = SHARE, y = TYPE, yend = TYPE),
    linewidth = lolli_line,
    colour = "gold",
    linetype = "dotted"
  ) +

  geom_text(
    data = subset(fleet, !is_falcon),
    aes(x = 0.02, y = TYPE, label = share_label),
    hjust = 0,
    size = 8,
    family = "space",
    #colour = "grey30",
    fontface = "bold"
  ) +

  geom_image(
    data = subset(fleet, !is_falcon),
    aes(
      x = image_x,
      y = TYPE,
      image = path
    ),
    size = 0.4,
    asp = 1
  ) +

  scale_x_continuous(
    limits = c(0, 1.6),
    expand = c(0, NA),
    labels = percent_format(),
    breaks = c(0, 0.25, 0.5, 0.75, 1)
  ) +

  scale_y_discrete(expand = expansion(add = 0.6)) +

  scale_colour_manual(values = c("grey80", "gold"), guide = "none") +

  canvas(width = mywidth, height = myheight, dpi = mydpi)


## ============================================================
## FINAL ANNOTATIONS
## ============================================================

p_falcon <- p_base +

  with_outer_glow(
    geom_image(
      data = subset(fleet, is_falcon),
      aes(x = image_x - 0.22, y = TYPE, image = path),
      size = 0.4
    ),
    colour = "#FFD166",
    sigma = 40
    ,nudge_y = 1.5
  ) +

  annotate(
    "text",
    x = 1.25,
    y = "A339", #"Millennium\nFalcon",
    label = "Hyperdrive engaged\ntoo fast to be counted!",
    family = "orbitron",
    size = 12,
    lineheight = 0.3
  )

p_final <- p_falcon +

  annotate(
    "text",
    x = 0.5,
    y = "A339",
    label = "← Leading the pack:\nAirbus A330neo",
    family = "space",
    size = 14,
    hjust = 0,
    lineheight = 0.3
  ) +

  annotate(
    "segment",
    x = 0.28, xend = 0.32,
    y = "B772", yend = "B772",
    linetype = "dashed"
  ) +

  annotate(
    "text",
    x = 0.9,
    y = "A359",
    label = "Boeing vs Airbus:\nNearly balanced",
    family = "space",
    size = 16,
    hjust = 0.5,
    lineheight = 0.3
  ) +

  labs(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    x = NULL,
    y = NULL
  ) +

  theme_clean(base_size = 24) +
  theme(
    plot.title = element_textbox_simple(
      size = rel(3),
      family = "orbitron",
      lineheight = 0.35,
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",

    plot.subtitle = element_textbox_simple(
      size = rel(2),
      family = "space",
      lineheight = 0.35,
      margin = margin(b = 15)
    ),

    plot.caption = element_textbox_simple(
      size = rel(1.3),
      family = "space",
      lineheight = 0.4,
      margin = margin(t = 15)
    ),
    plot.caption.position = "plot",

    axis.text.y = element_text(
      size = rel(1.5),
      family = "space",
      lineheight = 0.3
    ),

    axis.text.x = element_text(
      size = rel(1.5),
      family = "space"
    ),

    plot.margin = margin(t = 10, r = 20, b = 10, l = 20)
  )


## ============================================================
## RENDER
## ============================================================

p_final
```


