---
title: "Day 19 - Evolution"
format: html
---

## Aim

Today's prompt is interpreted as an aviation recovery arc: how did traffic evolve from the
pre-pandemic baseline through the latest 2024/2025 data?

The chart keeps the method simple. We rebuild a complete European network-traffic series
from the PBWG archive, pair it with the Brazilian daily traffic series from the 2025 project,
index both regions to their own 2019 median, and then show daily points with a 7-day rolling
median line.

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(arrow)
library(ggtext)
library(scales)
library(zoo)
```

## Build Daily Network Files

We run this once to generate the underlying data files for Day 19.

```{}
#| label: build-daily-network-files

dir.create("data", showWarnings = FALSE)

network_dir <- "../PBWG-data/network-traffic"

read_eur_network <- function(path) {
  readr::read_csv(path, show_col_types = FALSE) |>
    transmute(
      DATE = as.Date(DATE),
      REG = "EUR",
      FLTS = as.numeric(FLTS),
      SOURCE = basename(path),
      IS_FIX = str_detect(basename(path), "fix")
    )
}

eur_regular_files <- file.path(
  network_dir,
  paste0("PBWG-EUR-network-traffic-", 2021:2025, ".csv")
)

eur_fix_files <- file.path(
  network_dir,
  c(
    "PBWG-EUR-network-traffic-2019-fix.csv",
    "PBWG-EUR-network-traffic-2020-fix.csv",
    "PBWG-EUR-network-traffic-2024-fix.csv"
  )
)

eur_network_daily <- purrr::map_dfr(c(eur_regular_files, eur_fix_files), read_eur_network) |>
  arrange(DATE, desc(IS_FIX)) |>
  group_by(DATE, REG) |>
  summarise(FLTS = first(FLTS), .groups = "drop") |>
  arrange(DATE) |>
  mutate(
    YEAR = lubridate::year(DATE),
    ROLL_MED_7 = zoo::rollmedian(FLTS, k = 7, fill = NA, align = "center")
  )

bra_network_daily <- arrow::read_parquet("../BRA-EUR-2025/data/ndf_bra_totalbr_daily.parquet") |>
  transmute(
    DATE = as.Date(DATE),
    REG = "BRA",
    FLTS = as.numeric(DLY_FLTS),
    YEAR = lubridate::year(DATE),
    ROLL_MED_7 = zoo::rollmedian(FLTS, k = 7, fill = NA, align = "center")
  ) |>
  arrange(DATE)

readr::write_csv(eur_network_daily, "data/EUR-network-traffic-2019-2025.csv")
readr::write_csv(bra_network_daily, "data/BRA-network-traffic-2019-2025.csv")
```

## Read Challenge Inputs

```{r}
#| label: read-challenge-inputs

network_daily <- bind_rows(
  readr::read_csv("data/EUR-network-traffic-2019-2025.csv", show_col_types = FALSE),
  readr::read_csv("data/BRA-network-traffic-2019-2025.csv", show_col_types = FALSE)
) |>
  mutate(
    DATE = as.Date(DATE),
    REGION = recode(REG, BRA = "Brazil", EUR = "Europe")
  )

baseline_2019 <- network_daily |>
  filter(YEAR == 2019) |>
  summarise(
    BASELINE = median(FLTS, na.rm = TRUE),
    .by = REGION
  )

plot_data <- network_daily |>
  left_join(baseline_2019, by = "REGION") |>
  mutate(
    INDEX_2019 = 100 * FLTS / BASELINE,
    ROLL_INDEX_2019 = 100 * ROLL_MED_7 / BASELINE,
    REGION = factor(REGION, levels = c("Brazil", "Europe"))
  )

recovery_summary <- plot_data |>
  filter(YEAR %in% c(2020, 2025)) |>
  summarise(
    MED_INDEX = median(INDEX_2019, na.rm = TRUE),
    .by = c(REGION, YEAR)
  ) |>
  mutate(MED_INDEX_LAB = percent(MED_INDEX / 100, accuracy = 1))

label_data <- plot_data |>
  filter(!is.na(ROLL_INDEX_2019), REGION %in% c("Brazil", "Europe")) |>
  slice_max(DATE, n = 1, with_ties = FALSE, by = REGION) |>
  select(REGION, DATE, ROLL_INDEX_2019)

label_data2 <- tribble(
  ~REGION, ~DATE, ~ROLL_INDEX_2019
  ,"Europe", ymd("2024-03-01"), 65
  ,"Brazil", ymd("2024-01-01"), 130
)
```

## Final Candidate

```{r}
#| label: final-candidate
#| fig-width: 11
#| fig-height: 7

palette_region <- c(
  Brazil = "#52854C",
  Europe = "#4E84C4"
)


# clean some data irritations
plot_data <- plot_data |> filter(INDEX_2019 > 5)

p_day19 <- ggplot(plot_data, aes(x = DATE, colour = REGION)) +
  
  # main payload plot - timeline ---------------
  # daily indexed traffic
  geom_point(aes(y = INDEX_2019) , alpha = 0.5 , size = 0.6, stroke = 0) +
  # rolling median line 
  geom_line(aes(y = ROLL_INDEX_2019), linewidth = 1.05, na.rm = TRUE) +
  
  scale_colour_manual(values = palette_region, guide = "none") +
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y",
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(
    labels = label_number(suffix = ""),
    breaks = seq(0, 150, 25),
    limits = c(0, 150),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  geom_text(
    data = label_data2,
    aes(y = ROLL_INDEX_2019, label = REGION),
  ##  hjust = 0,
  #  nudge_x = -45,
    size = rel(3),
    fontface = "bold",
    show.legend = FALSE
  )  +
  coord_cartesian(clip = "off")
  
p_day19
```

## Recovery Markers

```{r}
#| label: recovery-markers

recovery_summary
```

## Move to Final Plot

```{r}
our_title <- paste0(
  "<span style='font-size:12pt'><strong>Chart-Challenge Day 19 - Evolution (Timeline)</strong></span><br>",
  "<strong>The Recovery Arc</strong>"
)

our_subtitle <- paste0(
  # ---- hook ------------------------
  "While both regions faced a staggering <b>80-90% traffic collapse</b> in 2020, their recovery 'evolution' follows distinct paths. ",
  # --- context / what does it show -----------------
  "<br>The chart shows the <b>daily traffic</b> indexed to each region's 2019 median (=100). Dots show daily data; the line is a centered <b>7-day rolling median</b>.",
  # --- insights / take-awys -------------------------
  "<br><b>Brazil</b> leveraged a strong domestic market to exceed pre-pandemic levels by late 2022, now trending <b>25% above</b> its 2019 median. ",
  "<br><b>Europe</b> remains shaped by high seasonality showing peaking strong summers versus winter troughs. With 2025, Europe has reached the pre-COVID traffic levels."
)

our_caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Data: Brazil-Europe Comparison Report, available at: github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge"
)
```

```{r}
# our plot dimensions
height = 7
width  = 10
dpi    = 300

base_plot <- ggplot(plot_data, aes(x = DATE, colour = REGION)) +
  
  # COVID period annotations ---------------------------------------
  annotate(
    "rect",
    xmin = as.Date("2020-02-25"),
    xmax = as.Date("2022-03-31"),
    ymin = -Inf,
    ymax = Inf,
    fill = "#d8d2c6",
    alpha = 0.28
  ) +
  annotate(
    "text",
    x = as.Date("2020-03-01"),
    y = 135,
    hjust = 0,
    label = "COVID shock, recovery with \npartial reopening, and \nstabilising return of air traffic \ndemand",
    size = 12/.pt,
    colour = "#6b6259"
  ) + 
  
  # baseline - reference hline --------------------------------------
  geom_hline(yintercept = 100, linewidth = 0.5, colour = "#6b6259") +
  
  # main payload plot - timeline ------------------------------------
  # daily indexed traffic
  geom_point(aes(y = INDEX_2019) , alpha = 0.5 , size = 0.6, stroke = 0)  + 
  # rolling median line 
  geom_line(aes(y = ROLL_INDEX_2019), linewidth = 1.05, na.rm = TRUE) +
  
  scale_colour_manual(values = palette_region, guide = "none") +
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y",
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(
    labels = label_number(suffix = ""),
    breaks = seq(0, 150, 25),
    limits = c(0, 150),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  geom_text(
    data = label_data2,
    aes(y = ROLL_INDEX_2019, label = REGION),
  ##  hjust = 0,
  #  nudge_x = -45,
    size = 12/.pt,
    fontface = "bold",
    show.legend = FALSE
  )  +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 12)
```

base_plot + ggview::canvas(width = width, height = height, dpi = dpi)

```{r}
p_day19_final <- base_plot +
  labs(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    x = NULL,
    y = "Traffic index, 2019 median = 100"
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_textbox_simple(
      size = rel(2),
      lineheight = 1.08,
      margin = margin(b = 8),
      colour = "#181818"
    ),
    plot.subtitle.position = "plot", 
    plot.subtitle = element_textbox_simple(
      size = rel(1),
      lineheight = 1.1,
      margin = margin(b = 12),
      colour = "#333333"
    ),
    plot.caption.position = "plot",
    plot.caption = element_textbox_simple(
      size = rel(0.9),
      colour = "#555555",
      hjust = 1,
      margin = margin(t = 10)
    )
  )
```

visual inspection with final dimensions using ggview:
   
p_day19_final + ggview::canvas(width = width, height = height, dpi = dpi)    


## Save Draft

```{r}
#| label: save-draft

p_day19 <- p_day19_final

ggsave(
  "figures/day19-evolution.png",
  plot = p_day19,
  width = width,
  height = height,
  dpi = dpi,
  bg = "white"
)
```
