---
title: "Day 16 - Simple Causal Graph + Evidence Panels"
format: html
---

## Aim

For today's prompt we keep the causal story deliberately small. The chart does not claim to identify causal effects. Instead, it presents a compact hypothesis - daily demand creates traffic pressure, and traffic pressure is associated with delay - then attaches simple evidence panels to the observable links.

The working language is modest by design: "consistent with", "proxy", and "observational".

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggraph)
library(tidygraph)
library(ggtext)
library(ggrepel)
library(scales)
library(grid)
```

## Data Preparation

```{r}
#| label: data-preparation

bin_midpoints <- tribble(
  ~BIN, ~DLY_MIN,
  "(-INF,-60]", -70,
  "(-60,-55]", -57.5,
  "(-55,-50]", -52.5,
  "(-50,-45]", -47.5,
  "(-45,-40]", -42.5,
  "(-40,-35]", -37.5,
  "(-35,-30]", -32.5,
  "(-30,-25]", -27.5,
  "(-25,-20]", -22.5,
  "(-20,-15]", -17.5,
  "(-15,-10]", -12.5,
  "(-10,-5]", -7.5,
  "(-5,0]", -2.5,
  "(0,5)", 2.5,
  "[5,10)", 7.5,
  "[10,15)", 12.5,
  "[15,20)", 17.5,
  "[20,25)", 22.5,
  "[25,30)", 27.5,
  "[30,35)", 32.5,
  "[35,40)", 37.5,
  "[40,45)", 42.5,
  "[45,50)", 47.5,
  "[50,55)", 52.5,
  "[55,60)", 57.5,
  "[60,INF)", 70
)

airport_names <- c(
  EDDF = "Frankfurt",
  EDDM = "Munich",
  EGLL = "Heathrow",
  EHAM = "Amsterdam",
  LEBL = "Barcelona",
  LEMD = "Madrid",
  LFPG = "Paris CDG",
  LGAV = "Athens",
  LIRF = "Rome Fiumicino",
  LPPT = "Lisbon",
  LSZH = "Zurich",
  LTFM = "Istanbul"
)

traffic_daily <- readr::read_csv(
  "./data/EUR-apt-tfc-2025.csv",
  show_col_types = FALSE
) |>
  mutate(MOVEMENTS = ARRS + DEPS) |>
  group_by(ICAO) |>
  mutate(
    PRESSURE = MOVEMENTS / quantile(MOVEMENTS, 0.95, na.rm = TRUE),
    PRESSURE_PCTL = percent_rank(MOVEMENTS)
  ) |>
  ungroup()

delay_daily <- readr::read_csv(
  "./data/PBWG-EUR-punc-2025.csv",
  show_col_types = FALSE
) |>
  pivot_longer(
    cols = matches("^\\(|^\\["),
    names_to = "BIN",
    values_to = "N_FLTS"
  ) |>
  left_join(bin_midpoints, by = "BIN") |>
  mutate(
    N_DLY_15 = if_else(DLY_MIN >= 15, N_FLTS, 0),
    POS_DLY_MIN = pmax(DLY_MIN, 0) * N_FLTS
  ) |>
  group_by(ICAO, DATE, PHASE) |>
  summarise(
    N_PHASE = sum(N_FLTS, na.rm = TRUE),
    MEAN_DLY = sum(N_FLTS * DLY_MIN, na.rm = TRUE) / N_PHASE,
    SHARE_DLY_15 = sum(N_DLY_15, na.rm = TRUE) / N_PHASE,
    POS_DLY_PER_FLT = sum(POS_DLY_MIN, na.rm = TRUE) / N_PHASE,
    .groups = "drop"
  ) |>
  group_by(ICAO, DATE) |>
  summarise(
    N_FLTS = sum(N_PHASE, na.rm = TRUE),
    MEAN_DLY = sum(MEAN_DLY * N_PHASE, na.rm = TRUE) / N_FLTS,
    SHARE_DLY_15 = sum(SHARE_DLY_15 * N_PHASE, na.rm = TRUE) / N_FLTS,
    POS_DLY_PER_FLT = sum(POS_DLY_PER_FLT * N_PHASE, na.rm = TRUE) / N_FLTS,
    .groups = "drop"
  )

evidence_daily <- traffic_daily |>
  inner_join(delay_daily, by = join_by(ICAO, DATE)) |>
  mutate(
    APT_LAB = recode(ICAO, !!!as.list(airport_names), .default = ICAO),
    PRESSURE_GROUP = case_when(
      PRESSURE_PCTL <= 0.25 ~ "Lower-pressure days",
      PRESSURE_PCTL >= 0.75 ~ "Higher-pressure days",
      TRUE ~ "Middle days"
    ),
    PRESSURE_GROUP = factor(
      PRESSURE_GROUP,
      levels = c("Lower-pressure days", "Higher-pressure days", "Middle days")
    )
  )

delay_tail <- readr::read_csv(
  "./data/PBWG-EUR-punc-2025.csv",
  show_col_types = FALSE
) |>
  pivot_longer(
    cols = matches("^\\(|^\\["),
    names_to = "BIN",
    values_to = "N_FLTS"
  ) |>
  left_join(bin_midpoints, by = "BIN") |>
  inner_join(
    evidence_daily |>
      filter(PRESSURE_GROUP != "Middle days") |>
      select(ICAO, DATE, PRESSURE_GROUP),
    by = join_by(ICAO, DATE)
  ) |>
  mutate(
    BIN_GROUP = case_when(
      DLY_MIN < -15 ~ "Early >15",
      DLY_MIN < 0 ~ "Early 0-15",
      DLY_MIN < 15 ~ "Late 0-15",
      DLY_MIN < 30 ~ "Late 15-30",
      TRUE ~ "Late >30"
    ),
    BIN_GROUP = factor(
      BIN_GROUP,
      levels = c("Early >15", "Early 0-15", "Late 0-15", "Late 15-30", "Late >30")
    )
  ) |>
  group_by(PRESSURE_GROUP, BIN_GROUP) |>
  summarise(N_FLTS = sum(N_FLTS, na.rm = TRUE), .groups = "drop") |>
  group_by(PRESSURE_GROUP) |>
  mutate(SHARE = N_FLTS / sum(N_FLTS)) |>
  ungroup()

pressure_summary <- evidence_daily |>
  group_by(PRESSURE_GROUP) |>
  summarise(
    AVG_SHARE_DLY_15 = mean(SHARE_DLY_15, na.rm = TRUE),
    AVG_POS_DLY = mean(POS_DLY_PER_FLT, na.rm = TRUE),
    .groups = "drop"
  )

delta_dly_15 <- pressure_summary |>
  filter(PRESSURE_GROUP %in% c("Lower-pressure days", "Higher-pressure days")) |>
  arrange(PRESSURE_GROUP) |>
  summarise(delta = diff(AVG_SHARE_DLY_15)) |>
  pull(delta)

outlier_days <- evidence_daily |>
  filter(PRESSURE_PCTL >= 0.85) |>
  group_by(ICAO) |>
  slice_max(SHARE_DLY_15, n = 1, with_ties = FALSE) |>
  ungroup() |>
  slice_max(SHARE_DLY_15, n = 6, with_ties = FALSE) |>
  mutate(
    OUTLIER_LAB = paste0(APT_LAB, "\n", format(DATE, "%d %b"))
  )
```

## Final Candidate

```{r}
#| label: final-candidate
#| fig-width: 10
#| fig-height: 10

palette_edge <- c(
  "Constructed proxy" = "#5a7d8c",
  "Observed association" = "#c8534f",
  "Interpretation" = "#7a7a7a"
)

palette_pressure <- c(
  "Lower-pressure days" = "#5a7d8c",
  "Higher-pressure days" = "#c8534f"
)

graph_nodes <- tibble(
  name = c("demand", "pressure", "delay", "disruption"),
  label = c(
    "Daily demand<br><span style='font-size:8pt'>ARRS + DEPS</span>",
    "Traffic pressure<br><span style='font-size:8pt'>within-airport proxy</span>",
    "Delay outcome<br><span style='font-size:8pt'>share delayed 15+ min</span>",
    "Network disruption<br><span style='font-size:8pt'>plausible, not tested</span>"
  ),
  x = c(0, 1, 2, 3),
  y = c(0, 0, 0, 0),
  fill = c("#f8f4e9", "#edf4f6", "#fbefef", "#f3f3f3")
)

graph_edges <- tibble(
  from = c("demand", "pressure", "delay"),
  to = c("pressure", "delay", "disruption"),
  evidence = c("Constructed proxy", "Observed association", "Interpretation"),
  label = c("proxy", "consistent with", "not shown")
)

graph_edge_labels <- tibble(
  x = c(0.5, 1.5, 2.5),
  y = c(0.16, 0.16, 0.16),
  evidence = c("Constructed proxy", "Observed association", "Interpretation"),
  label = c("proxy", "consistent with", "not shown")
)

causal_graph <- tbl_graph(
  nodes = graph_nodes,
  edges = graph_edges,
  directed = TRUE
)

p_graph <- ggraph(causal_graph, layout = "manual", x = x, y = y) +
  geom_edge_link(
    aes(edge_colour = evidence, edge_linetype = evidence),
    arrow = arrow(length = unit(4.5, "mm"), type = "closed"),
    start_cap = circle(8, "mm"),
    end_cap = circle(9, "mm"),
    linewidth = 1.1
  ) +
  geom_label(
    data = graph_edge_labels,
    aes(x = x, y = y, label = label, colour = evidence),
    inherit.aes = FALSE,
    linewidth = 0,
    label.padding = unit(0.12, "lines"),
    fill = "white",
    size = 3.1,
    show.legend = FALSE
  ) +
  geom_richtext(
    data = graph_nodes,
    aes(x = x, y = y, label = label, fill = fill),
    inherit.aes = FALSE,
    label.color = "#555555",
    label.r = unit(0.10, "lines"),
    label.padding = unit(0.28, "lines"),
    size = 3.15,
    lineheight = 0.92,
    colour = "#222222"
  ) +
  scale_fill_identity() +
  scale_edge_colour_manual(values = palette_edge, name = NULL) +
  scale_edge_linetype_manual(
    values = c(
      "Constructed proxy" = "solid",
      "Observed association" = "solid",
      "Interpretation" = "22"
    ),
    name = NULL
  ) +
  scale_colour_manual(values = palette_edge, name = NULL) +
  coord_cartesian(xlim = c(-0.18, 3.18), ylim = c(-0.35, 0.35), clip = "off") +
  labs(title = "Hypothesised causal chain") +
  theme_void(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", colour = "#222222", margin = margin(b = 4)),
    legend.position = "none",
    plot.margin = margin(0, 18, 0, 18)
  )

p_pressure_delay <- ggplot(
  evidence_daily,
  aes(x = PRESSURE_PCTL, y = SHARE_DLY_15)
) +
  annotate("rect", xmin = 0.75, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#f8dddd", alpha = 0.55) +
  geom_point(aes(size = MOVEMENTS), alpha = 0.18, colour = "#3f5360") +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, colour = "#c8534f", fill = "#f1b6b3", linewidth = 1) +
  geom_label_repel(
    data = outlier_days,
    aes(label = OUTLIER_LAB),
    seed = 16,
    size = 2.7,
    lineheight = 0.88,
    label.size = 0,
    label.padding = unit(0.12, "lines"),
    fill = alpha("#fffdf8", 0.86),
    colour = "#273942",
    box.padding = 0.35,
    point.padding = 0.35,
    min.segment.length = 0,
    segment.color = alpha("#273942", 0.55),
    segment.size = 0.25,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_size_continuous(range = c(0.4, 2.8), guide = "none") +
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, max(evidence_daily$SHARE_DLY_15, na.rm = TRUE) * 1.08)
  ) +
  labs(
    title = "Pressure and delay move together",
    subtitle = "Each dot is one airport-day. Pressure is the daily traffic percentile within that airport's own 2025 distribution.",
    x = "Traffic pressure percentile",
    y = "Flights delayed 15+ min"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", colour = "#222222"),
    plot.subtitle = element_textbox_simple(size = 8.7, lineheight = 1.05, margin = margin(b = 5)),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold", colour = "#333333"),
    plot.margin = margin(4, 8, 4, 8)
  )

p_tail <- delay_tail |>
  filter(PRESSURE_GROUP != "Middle days") |>
  ggplot(aes(x = SHARE, y = BIN_GROUP, fill = PRESSURE_GROUP)) +
  geom_col(
    position = position_dodge2(width = 0.74, preserve = "single"),
    width = 0.64,
    alpha = 0.92
  ) +
  scale_fill_manual(values = palette_pressure, name = NULL) +
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.06))) +
  labs(
    title = "Busy days carry a thicker delay tail",
    subtitle = "Comparing the lower and upper quartile of each airport's daily traffic.",
    x = "Share of flights",
    y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", colour = "#222222"),
    plot.subtitle = element_textbox_simple(size = 8.7, lineheight = 1.05, margin = margin(b = 5)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    axis.title = element_text(face = "bold", colour = "#333333"),
    plot.margin = margin(4, 8, 4, 8)
  )

p_notes <- ggplot() +
  annotate(
    "richtext",
    x = 0,
    y = 1,
    hjust = 0,
    vjust = 1,
    label = paste0(
      "<b>Reading the evidence status</b><br>",
      "<span style='color:#5a7d8c'>Constructed proxy</span>: daily movements are used as a simple pressure measure.<br>",
      "<span style='color:#c8534f'>Observed association</span>: higher pressure days tend to have more 15+ minute delay.<br>",
      "<span style='color:#7a7a7a'>Interpretation</span>: wider disruption is plausible, but not estimated here."
    ),
    label.color = NA,
    fill = NA,
    size = 3.1,
    lineheight = 1.08
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(0, 18, 0, 18))

p_notes_social <- ggplot() +
  annotate(
    "richtext",
    x = 0,
    y = 1,
    hjust = 0,
    vjust = 1,
    label = paste0(
      "<b>Evidence status:</b> ",
      "<span style='color:#5a7d8c'>proxy</span> = daily movements as pressure; ",
      "<span style='color:#c8534f'>observed association</span> = pressure vs delay; ",
      "<span style='color:#7a7a7a'>interpretation</span> = disruption not estimated."
    ),
    label.color = NA,
    fill = NA,
    size = 2.85,
    lineheight = 1.03
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(0, 18, 0, 18))

our_title <- paste0(
  "<span style='font-size:12pt'><strong>Chart-Challenge Day 16 - Simple causal graph + evidence panels</strong></span><br>",
  "<strong>A small causal story, with the evidence kept attached</strong>"
)

our_subtitle <- paste0(
  "The directed graph states a deliberately modest hypothesis: more daily demand can create traffic pressure, ",
  "and higher pressure is <strong>consistent with</strong> more delay. The evidence panels use 2025 European airport-day data only: ",
  "daily arrivals plus departures define a within-airport pressure proxy, while binned punctuality data estimate the share of flights delayed by 15 minutes or more. ",
  "Higher-pressure days show about <strong>", percent(delta_dly_15, accuracy = 0.1),
  "</strong> more flights in the 15+ minute delay group than lower-pressure days. This is observational evidence, not causal identification."
)

our_caption <- "DECEA Performance Section & EUROCONTROL PRU"

p_day16_candidate <- (
  p_graph / p_notes / (p_pressure_delay | p_tail)
) +
  plot_layout(heights = c(0.58, 0.26, 1.55)) +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 18,
        lineheight = 1.08,
        margin = margin(b = 8),
        colour = "#181818"
      ),
      plot.subtitle = element_textbox_simple(
        size = 10.5,
        lineheight = 1.08,
        margin = margin(b = 12),
        colour = "#333333"
      ),
      plot.caption = element_textbox_simple(
        size = 9,
        colour = "#555555",
        hjust = 1,
        margin = margin(t = 10)
      ),
      plot.background = element_rect(fill = "#fffdf8", colour = NA),
      plot.margin = margin(12, 14, 12, 14)
    )
  )

p_day16_candidate
```

## Social Crop

```{r}
#| label: social-crop
#| fig-width: 10
#| fig-height: 8

p_day16_social <- (
  p_graph / p_notes_social / (p_pressure_delay | p_tail)
) +
  plot_layout(heights = c(0.44, 0.24, 1.55)) +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 15,
        lineheight = 1.05,
        margin = margin(b = 6),
        colour = "#181818"
      ),
      plot.subtitle = element_textbox_simple(
        size = 9.2,
        lineheight = 1.05,
        margin = margin(b = 8),
        colour = "#333333"
      ),
      plot.caption = element_textbox_simple(
        size = 8.2,
        colour = "#555555",
        hjust = 1,
        margin = margin(t = 8)
      ),
      plot.background = element_rect(fill = "#fffdf8", colour = NA),
      plot.margin = margin(10, 12, 10, 12)
    )
  )

p_day16_social
```

## Save Draft

```{r}
#| label: save-draft
#| eval: false

ggsave(
  "./figures/day16-causal-evidence.png",
  plot = p_day16_candidate,
  width = 10,
  height = 10,
  dpi = 300,
  units = "in"
)

ggsave(
  "./figures/day16-causal-evidence-social.png",
  plot = p_day16_social,
  width = 10,
  height = 8,
  dpi = 300,
  units = "in"
)
```
