---
title: "Day 12 - Flowing Data"
format: html
---

## Aim

For Day 12 we wanted to turn the 2025 Europe-Brazil traffic network into a readable flow chart that highlights how strongly the market is concentrated around the study airports. The challenge was not just getting `geom_parallel_sets()` to work, but making the result legible enough for a challenge submission with meaningful titles, airport labels, and a polished export.

Packages used in the process:

* `{ggplot2}` and `{ggforce}` for the parallel-sets geometry
* `{dplyr}` and `{tidyr}` for reshaping and aggregation
* `{ggtext}` for title, subtitle, and caption blocks
* `{ggview}` to align the RStudio preview and saved output
* `{ggrepel}` for the outside-label refinement in the final proposal

## Setup And First Test

```{r}
#| label: setup
library(readr)
library(dplyr)
library(tidyr)

# visualisation
library(ggplot2)
library(ggforce)   # parallel sets  --> alluvial style
```

```{r}
flows_long <- read_csv(here::here("data","BRA-EUR-conns-parallel.csv"), show_col_types = FALSE)
```

```{r}
ggplot(flows_long, aes(x = x, id = id, split = y, value = DEPS)) +
geom_parallel_sets(aes(fill = ORIG_GROUP), alpha = 0.6, axis.width = 0.1) +
geom_parallel_sets_axes(axis.width = 0.2, fill = "gray90") +
geom_parallel_sets_labels(angle = 0, size = 3, colour = "black") +
scale_x_discrete(labels = c("Origin", "Destination")) +
scale_fill_manual(values = c("Study" = "blue", "Non-Study" = "lightblue")) +
theme_minimal() +
labs(
# title = "Flight Flows from Europe to Brazil (Study vs Non-Study)",
# subtitle = "Grouped by Origin and Destination Study Status",
y = "Number of Departures", x = element_blank(), fill = element_blank()
) +
theme(legend.position = "inside", legend.position.inside = c(0.1,0.1))
```

## Data Preparation

We then switched from the small CSV prototype to the full 2025 flow dataset and created the study / non-study grouping logic used throughout the rest of the file.

```{r}
#' helper function to assign region/study label based on ICAO or portions thereof
#' 
assign_region <- function(
      icao      # vector of ICAO location indicators
    , prefixes  # combinations of Letters looking for, including full ICAO
    , label = "Study"        # desired label
    , other = NA_character_  # alternative label, default NA
    ) {
    stopifnot(is.character(icao), is.character(prefixes))
    
    pattern <- paste0("^(", paste(prefixes, collapse = "|"), ")")
    
    ifelse(grepl(pattern, icao), label, other)
}

pattern_eur <- c("E","L","BI")
pattern_bra <- c("SB","SD","SI","SJ","SN","SS", "SW")

# study airports
bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```

get NM flights

```{r}
rq <- arrow::read_parquet("./data/EUR-adep-ades-2025.parquet") |> 
  mutate(
    ORIG_REG = assign_region(ADEP, c(pattern_eur,"AFIL"), "EUR")
  , DEST_REG = assign_region(ADES, pattern_bra,"BRA")) |> 
  filter(ORIG_REG == "EUR", DEST_REG == "BRA")

# assign study label
rq <- rq |> 
  mutate(
    ORIG_GROUP = assign_region(ADEP, eur_apts, "study", "non-study")
  , DEST_GROUP = assign_region(ADES, bra_apts, "study", "non-study")
  )
```

## Reshaping Notes

Before moving on to the final design, it was useful to keep a short reminder of how `geom_parallel_sets()` expects its data.

Short explanation (x / id / split)

id
Identifies one flow unit.
→ One ADEP–ADES pair (or aggregated version of it).
→ Must be constant across axes.

x
Identifies the axis position.
→ "Origin" vs "Destination" (factor, ordered).

split (your y)
The category on each axis.
→ At Origin: ORIG_GROUP
→ At Destination: DEST_GROUP

value
Flow thickness.
→ FLTS

Each flow therefore becomes two rows:
one at Origin, one at Destination, sharing the same id.


```{r}
flows_long1 <- rq |>
    mutate(
        id = row_number()
    ) |>
    select(id, FLTS, ORIG_GROUP, DEST_GROUP) |>
    pivot_longer(
        cols = c(ORIG_GROUP, DEST_GROUP),
        names_to = "x",
        values_to = "y"
    ) |>
    mutate(
        x = recode(
            x,
            ORIG_GROUP = "Origin",
            DEST_GROUP = "Destination"
        ),
        x = factor(x, levels = c("Origin", "Destination"))
    )
```

```{r}
ggplot(flows_long1, aes(x = x, id = id, split = y, value = FLTS)) +
  geom_parallel_sets(alpha = 0.6, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.2, fill = "gray90") +
  geom_parallel_sets_labels(angle = 0, size = 3, colour = "black") +
  
  scale_x_discrete(labels = c("Origin", "Destination")) +
  #scale_fill_manual(values = c("Study" = "blue", "Non-Study" = "lightblue")) +
  theme_minimal() +
  labs(
# title = "Flight Flows from Europe to Brazil (Study vs Non-Study)",
# subtitle = "Grouped by Origin and Destination Study Status",
    y = "Number of Departures", x = element_blank(), fill = element_blank()
) +
  theme(legend.position = "inside", legend.position.inside = c(0.1,0.1))
```

## Collapsing The Network

The next step was to collapse the long tail of small airports into grouped buckets while preserving the study-airport structure that matters for the story.

```{r}
threshold <- 50

rq_collapsed <- rq |>
  mutate(
    ORIG_NODE = case_when(
      ORIG_GROUP == "study" ~ ADEP,
      ORIG_GROUP != "study" & FLTS >= threshold ~ ADEP,
      ORIG_GROUP != "study" & ORIG_REG == "EUR" ~ "Other European",
      TRUE ~ "Other Brazilian"
    ),
    DEST_NODE = case_when(
      DEST_GROUP == "study" ~ ADES,
      DEST_GROUP != "study" & FLTS >= threshold ~ ADES,
      DEST_GROUP != "study" & DEST_REG == "BRA" ~ "Other Brazilian",
      TRUE ~ "Other European"
    )
  )

flows_long <- rq_collapsed |>
  group_by(ORIG_NODE, DEST_NODE) |>
  summarise(
    FLTS = sum(FLTS),
    .groups = "drop"
  ) |>
  mutate(id = row_number()) |>
  pivot_longer(
    cols = c(ORIG_NODE, DEST_NODE),
    names_to = "x",
    values_to = "y"
  ) |>
  mutate(
    x = recode(
      x,
      ORIG_NODE = "Origin",
      DEST_NODE = "Destination"
    ),
    x = factor(x, levels = c("Origin", "Destination"))
  )
```

```{r}
ggplot(flows_long, aes(x = x, id = id, split = y, value = FLTS)) +
#geom_parallel_sets(aes(fill = ORIG_NODE), alpha = 0.6, axis.width = 0.1) +
geom_parallel_sets_axes(axis.width = 0.2, fill = "gray90") +
geom_parallel_sets_labels(angle = 0, size = 3, colour = "black") +
scale_x_discrete(labels = c("Origin", "Destination")) +
scale_fill_manual(values = c("Study" = "blue", "Non-Study" = "lightblue")) +
theme_minimal() +
labs(
# title = "Flight Flows from Europe to Brazil (Study vs Non-Study)",
# subtitle = "Grouped by Origin and Destination Study Status",
y = "Number of Departures", x = element_blank(), fill = element_blank()
) +
theme(legend.position = "inside", legend.position.inside = c(0.1,0.1))
```

```{r}
rq_collapsed <- rq_collapsed |>
  mutate(
    FLOW_CLASS = case_when(
      ORIG_GROUP == "study" & DEST_GROUP == "study" ~ "Study–Study",
      ORIG_GROUP == "study" | DEST_GROUP == "study" ~ "Study–Non-Study",
      TRUE                                          ~ "Non-Study–Non-Study"
    )
  )

flows_long <- rq_collapsed |>
  group_by(ORIG_NODE, DEST_NODE) |>
  summarise(
    FLTS = sum(FLTS),
    .groups = "drop"
  ) |>
  mutate(id = row_number()) |>
  transmute(
    id,
    value = FLTS,
    x_origin = 1,
    x_dest   = 2,
    y_origin = ORIG_NODE,
    y_dest   = DEST_NODE
  ) |>
  pivot_longer(
    cols = c(x_origin, x_dest, y_origin, y_dest),
    names_to = c(".value", "axis"),
    names_pattern = "(x|y)_(origin|dest)"
  ) |>
  mutate(
    x = ifelse(axis == "origin", 1, 2)
  ) |>
  select(id, x, y, value)
```

```{r}
threshold <- 60

rq_collapsed <- rq |>
  mutate(
    ORIG_NODE = case_when(
      ORIG_GROUP == "study" ~ ADEP,
      ORIG_GROUP != "study" & FLTS >= threshold ~ ADEP,
      ORIG_GROUP != "study" & ORIG_REG == "EUR" ~ "Other European",
      TRUE ~ "Other Brazilian"
    ),
    DEST_NODE = case_when(
      DEST_GROUP == "study" ~ ADES,
      DEST_GROUP != "study" & FLTS >= threshold ~ ADES,
      DEST_GROUP != "study" & DEST_REG == "BRA" ~ "Other Brazilian",
      TRUE ~ "Other European"
    )
  )

rq_collapsed <- rq_collapsed |>
  mutate(
    FLOW_CLASS = case_when(
      ORIG_GROUP == "study" & DEST_GROUP == "study" ~ "Study–Study",
      ORIG_GROUP == "study" | DEST_GROUP == "study" ~ "Study–Non-Study",
      TRUE                                          ~ "Non-Study–Non-Study"
    )
  )

flows_long <- rq_collapsed |>
  group_by(ORIG_NODE, DEST_NODE, FLOW_CLASS) |>
  summarise(
    FLTS = sum(FLTS),
    .groups = "drop"
  ) |>
  mutate(id = row_number()) |>
  transmute(
    id,
    value = FLTS,
    FLOW_CLASS,
    x_origin = 1,
    x_dest   = 2,
    y_origin = ORIG_NODE,
    y_dest   = DEST_NODE
  ) |>
  pivot_longer(
    cols = c(x_origin, x_dest, y_origin, y_dest),
    names_to = c(".value", "axis"),
    names_pattern = "(x|y)_(origin|dest)"
  ) #|>
#  select(id, x, y, value, FLOW_CLASS)

# add label for fill of axis
flows_long <- flows_long |>
  mutate(
    AXIS_FILL = ifelse(
      y %in% c(bra_apts, eur_apts),
      "study",
      "non-study"
    ),
    FLOW_FILL = recode(
      FLOW_CLASS,
      "Study–Study"         = "blue",
      "Study–Non-Study"     = "lightblue",
      "Non-Study–Non-Study" = "grey70"
    ),
    AXIS_FILL_COL = recode(
      AXIS_FILL,
      "study"     = "yellow",
      "non-study" = "grey70"
    )
  )

```

```{r}
ggplot(
  flows_long,
  aes(
    x     = x,
    id    = id,
    split = y,
    value = value
  )
) +
  # ------------------ the flow(s) ----------------------------
  geom_parallel_sets(
    aes(fill = FLOW_FILL),
    alpha = 0.6,
    axis.width = 0.1
  ) +
  # ------------------- axes "origin" & destination boxes -----
  geom_parallel_sets_axes(
    axis.width = 0.2,
    aes(fill = AXIS_FILL_COL),
    colour = "grey40"
  ) +
  scale_fill_identity() +

  geom_parallel_sets_labels(
    size = 3,
    colour = "black",
    angle = 0
  ) +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = NULL # c("Origin", "Destination")
  ) +
  labs(
    x = NULL,
    y = NULL        # "Number of departures"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

## Styling The Flow And Axis Fills

This section contains the intermediate solution where the flow colors and axis-box colors were handled separately before we moved to the final outside-label proposal.

## Final Version

From here onward the focus shifts to the challenge submission itself: stronger storytelling, readable airport naming, a consistent caption block, and a plot size that behaves the same in the viewer and on export.

```{r}
library(ggtext)
library(ggview)
library(scales)

plot_width  <- 10
plot_height <- 8
plot_dpi    <- 300

col_study <- "#7DB7E8"
col_mixed <- "#B9D9F3"
col_non   <- "grey70"

airport_names <- c(
  EGLL = "Heathrow",
  EGKK = "Gatwick",
  EHAM = "Amsterdam",
  EDDF = "Frankfurt",
  EDDM = "Munich",
  ELLX = "Luxembourg",
  LSZH = "Zurich",
  LFPG = "Paris CDG",
  LEMD = "Madrid",
  LEBL = "Barcelona",
  LPPT = "Lisbon",
  LPPR = "Porto",
  LTFM = "Istanbul",
  LIMC = "Milan",
  LIRF = "Rome",
  LGAV = "Athens",
  SBBE = "Belem",
  SBBR = "Brasilia",
  SBGR = "Guarulhos",
  SBSP = "Congonhas",
  SBKP = "Campinas",
  SBRJ = "Santos Dumont",
  SBGL = "Galeao",
  SBCF = "Belo Horizonte",
  SBFZ = "Fortaleza",
  SBSV = "Salvador",
  SBPA = "Porto Alegre",
  SBCT = "Curitiba",
  SBRF = "Recife",
  SBEG = "Manaus"
)

format_node_label <- function(code, flights) {
  airport_name <- dplyr::recode(code, !!!airport_names, .default = code)
  is_named <- code %in% names(airport_names)

  case_when(
    code == "Other European" ~ "Other European\nairports",
    code == "Other Brazilian" ~ "Other Brazilian\nairports",
    !is_named ~ code,
    TRUE ~ paste0(airport_name, "\n(", code, ")")
  )
}

node_totals <- bind_rows(
  rq_collapsed |>
    count(node = ORIG_NODE, wt = FLTS, name = "node_flights") |>
    mutate(x = 1),
  rq_collapsed |>
    count(node = DEST_NODE, wt = FLTS, name = "node_flights") |>
    mutate(x = 2)
)

flow_summary <- rq_collapsed |>
  count(FLOW_CLASS, wt = FLTS, name = "flights") |>
  mutate(share = flights / sum(flights))

total_flights <- sum(flow_summary$flights)
study_share <- flow_summary |>
  filter(FLOW_CLASS == "Study–Study") |>
  pull(share)

sbgr_share <- node_totals |>
  filter(x == 2, node == "SBGR") |>
  summarise(share = node_flights / total_flights) |>
  pull(share)

flows_final <- rq_collapsed |>
  group_by(ORIG_NODE, DEST_NODE, FLOW_CLASS) |>
  summarise(
    FLTS = sum(FLTS),
    .groups = "drop"
  ) |>
  mutate(
    id = row_number(),
    value = FLTS,
    x_origin = 1,
    x_dest   = 2,
    y_origin = ORIG_NODE,
    y_dest   = DEST_NODE
  ) |>
  pivot_longer(
    cols = c(x_origin, x_dest, y_origin, y_dest),
    names_to = c(".value", "axis"),
    names_pattern = "(x|y)_(origin|dest)"
  ) |>
  left_join(node_totals, by = c("x", "y" = "node")) |>
  mutate(
    AXIS_FILL = ifelse(y %in% c(bra_apts, eur_apts), "study", "non-study"),
    FLOW_FILL = recode(
      FLOW_CLASS,
      "Study–Study"         = col_study,
      "Study–Non-Study"     = col_mixed,
      "Non-Study–Non-Study" = col_non
    ),
    AXIS_FILL_COL = recode(
      AXIS_FILL,
      "study"     = col_study,
      "non-study" = "grey85"
    ),
    y_label = format_node_label(y, node_flights)
  )

axis_levels <- flows_final |>
  distinct(x, y, y_label) |>
  arrange(x, y) |>
  pull(y_label)

flows_final <- flows_final |>
  mutate(y_label = factor(y_label, levels = axis_levels))

our_title <- "<span style='font-size:12pt'><strong>Chart-Challenge Day 12 - Flow</strong></span><br><strong>Europe-Brazil traffic is concentrated in a small study network</strong>"

our_subtitle <- paste0(
  percent(study_share, accuracy = 1), " of ",
  comma(total_flights), " 2025 flights connect study airports at both ends. ",
  "Lisbon, Madrid and Paris CDG feed the largest European volumes, while ",
  "Guarulhos alone absorbs ", percent(sbgr_share, accuracy = 1), " of arrivals. ",
  "Airport names are expanded where box size permits; counts highlight the biggest nodes."
)

our_caption <- "DECEA Performance Section & EUROCONTROL PRU"

eu_hubs <- node_totals |>
  filter(x == 1) |>
  arrange(desc(node_flights)) |>
  slice_head(n = 3) |>
  mutate(
    hub_label = paste0(
      recode(node, !!!airport_names, .default = node),
      " ",
      percent(node_flights / total_flights, accuracy = 1),
      " (",
      comma(node_flights),
      ")"
    )
  ) |>
  pull(hub_label)

br_hubs <- node_totals |>
  filter(x == 2) |>
  arrange(desc(node_flights)) |>
  slice_head(n = 3) |>
  mutate(
    hub_label = paste0(
      recode(node, !!!airport_names, .default = node),
      " ",
      percent(node_flights / total_flights, accuracy = 1),
      " (",
      comma(node_flights),
      ")"
    )
  ) |>
  pull(hub_label)

p_day12_final <- ggplot(
  flows_final,
  aes(
    x = x,
    id = id,
    split = y_label,
    value = value
  )
) +
  geom_parallel_sets(
    aes(fill = FLOW_FILL),
    alpha = 0.7,
    axis.width = 0.12
  ) +
  geom_parallel_sets_axes(
    axis.width = 0.22,
    aes(fill = AXIS_FILL_COL),
    colour = "grey45",
    linewidth = 0.4
  ) +
  scale_fill_identity() +
  geom_parallel_sets_labels(
    size = 3.4,
    colour = "grey10",
    angle = 0,
    lineheight = 0.95
  ) +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("From Europe", "To Brazil"),
    expand = expansion(mult = c(0.08, 0.08))
  ) +
  coord_cartesian(
    xlim = c(0.72, 2.28),
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_textbox_simple(
      margin = margin(b = 10),
      lineheight = 1.15
    ),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      margin = margin(b = 12),
      lineheight = 1.1
    ),
    plot.caption = element_textbox_simple(
      margin = margin(t = 15),
      colour = "grey35"
    ),
    plot.caption.position = "plot",
    axis.text.x = element_text(face = "bold", colour = "grey20"),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(t = 10, r = 50, b = 10, l = 50)
  ) +
  canvas(width = plot_width, height = plot_height, dpi = plot_dpi)

big_nodes <- ggplot_build(p_day12_final)$data[[2]] |>
  as_tibble() |>
  distinct(x, label, value, ymin, ymax) |>
  mutate(
    y = (ymin + ymax) / 2,
    x_pos = ifelse(x < 1.5, 0.79, 2.21),
    hjust = ifelse(x < 1.5, 1, 0),
    flights_label = comma(value)
  ) |>
  filter(value >= 600)

p_day12_final <- p_day12_final +
  geom_text(
    data = big_nodes,
    aes(
      x = x_pos,
      y = y,
      label = flights_label,
      hjust = hjust
    ),
    inherit.aes = FALSE,
    size = 3.2,
    colour = "grey35",
    fontface = "bold"
  )

p_day12_final

ggsave(
  "./xx-figures/day12-flow-final.png",
  plot = p_day12_final,
  width = plot_width,
  height = plot_height,
  dpi = plot_dpi
)
```

## Final Deliverable - Outside Labels Proposal

The final candidate moves airport names outside the blocks, keeps service counts inside the larger nodes, and saves the deliverable to the standard `figures/` folder for the README and challenge post.

```{r}
axis_width_outside <- 0.12

p_day12_outside <- ggplot(
  flows_final,
  aes(
    x = x,
    id = id,
    split = y_label,
    value = value
  )
) +
  geom_parallel_sets(
    aes(fill = FLOW_FILL),
    alpha = 0.72,
    axis.width = axis_width_outside
  ) +
  geom_parallel_sets_axes(
    axis.width = axis_width_outside,
    aes(fill = AXIS_FILL_COL),
    colour = "grey45",
    linewidth = 0.35
  ) +
  scale_fill_identity() +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("From Europe", "To Brazil"),
    expand = expansion(mult = c(0.08, 0.08))
  ) +
  coord_cartesian(
    xlim = c(0.72, 2.28),
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = our_title,
    subtitle = paste0(
      percent(study_share, accuracy = 1), " of ",
      comma(total_flights), " air transport services between Europe and Brazil in 2025 linked study airports at both ends. ",
      "On the European side, ", paste(eu_hubs, collapse = ", "), " dominate the market. ",
      "In Brazil, ", paste(br_hubs, collapse = ", "), " receive the largest shares."
    ),
    caption = our_caption
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_textbox_simple(
      margin = margin(b = 10),
      lineheight = 1.15
    ),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      margin = margin(b = 12),
      lineheight = 1.1
    ),
    plot.caption = element_textbox_simple(
      margin = margin(t = 15),
      colour = "grey35"
    ),
    plot.caption.position = "plot",
    axis.text.x = element_text(face = "bold", colour = "grey20"),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(t = 10, r = 45, b = 10, l = 45)
  )

axis_nodes_outside <- ggplot_build(p_day12_outside)$data[[2]] |>
  as_tibble() |>
  mutate(axis_side = ifelse(x < 1.5, 1, 2)) |>
  distinct(axis_side, label, value, ymin, ymax, fill) |>
  mutate(
    y = (ymin + ymax) / 2,
    label_x = ifelse(axis_side == 1, 0.865, 2.135),
    label_hjust = ifelse(axis_side == 1, 1, 0),
    count_x = axis_side,
    count_label = comma(value),
    count_colour = ifelse(fill == "grey85", "grey15", "white")
  )

inside_counts <- axis_nodes_outside |>
  filter(value >= 650)

outside_labels <- axis_nodes_outside
outside_labels_left <- outside_labels |>
  filter(axis_side == 1)
outside_labels_right <- outside_labels |>
  filter(axis_side == 2)

p_day12_outside <- p_day12_outside +
  ggrepel::geom_text_repel(
    data = outside_labels_left,
    aes(
      x = label_x,
      y = y,
      label = label
    ),
    inherit.aes = FALSE,
    size = 2.75,
    colour = "grey20",
    lineheight = 0.9,
    hjust = 1,
    direction = "y",
    seed = 12,
    force = 1.2,
    box.padding = 0.08,
    point.padding = 0.02,
    min.segment.length = 0,
    segment.colour = "grey70",
    segment.size = 0.25
  ) +
  ggrepel::geom_text_repel(
    data = outside_labels_right,
    aes(
      x = label_x,
      y = y,
      label = label
    ),
    inherit.aes = FALSE,
    size = 2.75,
    colour = "grey20",
    lineheight = 0.9,
    hjust = 0,
    direction = "y",
    seed = 12,
    force = 1.2,
    box.padding = 0.08,
    point.padding = 0.02,
    min.segment.length = 0,
    segment.colour = "grey70",
    segment.size = 0.25
  ) +
  geom_text(
    data = inside_counts,
    aes(
      x = count_x,
      y = y,
      label = count_label,
      colour = count_colour
    ),
    inherit.aes = FALSE,
    size = 3.1,
    fontface = "bold",
    show.legend = FALSE
  ) +
  scale_colour_identity()

p_day12_outside_view <- p_day12_outside +
  canvas(width = plot_width, height = plot_height + 1, dpi = plot_dpi)

p_day12_outside_view

ggsave(
  "./figures/day12-flow.png",
  plot = p_day12_outside,
  width = plot_width,
  height = plot_height + 1,
  dpi = plot_dpi
)
```
