---
title: "Day 20 - Global Change"
format: html
---

## Aim

Today's prompt is interpreted as a small global-change timeline: how did taxi-out time
shift across the Brazil and Europe study airports from 2023 to 2025?

The chart keeps the method deliberately direct. We read the daily taxi-out aggregates,
reduce them to airport-month summaries, then compare the monthly median across airports.
The final version adds a compact message row above the timeline so the submission works
both as a chart and as a quick read.

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(patchwork)
library(glue)
library(ggtext)
```

## Read Taxi-Out Data

The analytic taxi-out files are the daily airport-level performance extracts prepared for
the Brazil-Europe comparison work. Each region contributes twelve study airports, with
one file per year.

```{r}
#| label: read-txxt-data

read_txxt <- function(path, region) {
  readr::read_csv(path, show_col_types = FALSE) |>
    transmute(
      ICAO,
      PHASE,
      DATE = as_date(DATE),
      MVTS_VALID,
      TOT_TXXT,
      region = region,
      avg_txxt = TOT_TXXT / MVTS_VALID
    )
}

bra_files <- list.files("data", pattern = "BRA-txxt", full.names = TRUE)
eur_files <- list.files("data", pattern = "EUR-txxt", full.names = TRUE)

raw_txxt <- bind_rows(
  map(bra_files, read_txxt, region = "Brazil (BRA)"),
  map(eur_files, read_txxt, region = "Europe (EUR)")
)
```

## Build Monthly Series

First we summarize each airport-month, then we summarize across airports. The central line
is therefore the median airport's monthly taxi-out time; the band shows the median of the
airport-level daily interquartile ranges.

```{r}
#| label: monthly-series

monthly <- raw_txxt |>
  filter(PHASE == "DEP", MVTS_VALID > 0) |>
  mutate(month = floor_date(DATE, "month")) |>
  summarise(
    med = median(avg_txxt, na.rm = TRUE),
    q25 = quantile(avg_txxt, 0.25, na.rm = TRUE),
    q75 = quantile(avg_txxt, 0.75, na.rm = TRUE),
    .by = c(ICAO, region, month)
  ) |>
  summarise(
    med_median = median(med, na.rm = TRUE),
    med_q25 = median(q25, na.rm = TRUE),
    med_q75 = median(q75, na.rm = TRUE),
    med_iqr = median(q75 - q25, na.rm = TRUE),
    .by = c(region, month)
  ) |>
  mutate(region = factor(region, levels = c("Brazil (BRA)", "Europe (EUR)"))) |>
  arrange(region, month)

summary_stats <- monthly |>
  mutate(year = year(month)) |>
  filter(year %in% c(2023, 2025)) |>
  summarise(avg = mean(med_median, na.rm = TRUE), .by = c(region, year)) |>
  pivot_wider(names_from = year, values_from = avg) |>
  mutate(
    delta = `2025` - `2023`,
    label = case_when(
      abs(delta) < 0.3 ~ "stable",
      delta > 0 ~ glue("+{round(delta, 1)} min"),
      TRUE ~ glue("{round(delta, 1)} min")
    )
  )

iqr_direction <- monthly |>
  mutate(year = year(month)) |>
  filter(year %in% c(2023, 2025)) |>
  summarise(avg_iqr = mean(med_iqr, na.rm = TRUE), .by = c(region, year)) |>
  pivot_wider(names_from = year, values_from = avg_iqr) |>
  mutate(delta = `2025` - `2023`)

eur_delta <- summary_stats |>
  filter(region == "Europe (EUR)") |>
  pull(label)

bra_delta <- summary_stats |>
  filter(region == "Brazil (BRA)") |>
  pull(label)

iqr_label <- if_else(
  all(abs(iqr_direction$delta) < 0.3),
  "mixed",
  "shifting"
)
```

## Base Plot

This is the core timeline before the submission typography and message panels are added.

```{r}
#| label: base-plot
#| fig-width: 10
#| fig-height: 5

col_bra <- "#0F6E56"
col_eur <- "#185FA5"
palette_region <- c("Brazil (BRA)" = col_bra, "Europe (EUR)" = col_eur)

p_base <- ggplot(monthly, aes(x = month, colour = region, fill = region)) +
  geom_ribbon(
    aes(ymin = med_q25, ymax = med_q75),
    alpha = 0.13,
    colour = NA,
    show.legend = FALSE
  ) +
  geom_line(aes(y = med_median, linetype = region), linewidth = 1.1) +
  geom_vline(
    xintercept = as_date(c("2024-01-01", "2025-01-01")),
    linetype = "dotted",
    colour = "#B4B2A9",
    linewidth = 0.4
  ) +
  annotate("text", x = as_date("2023-07-01"), y = 18.35,
           label = "2023", size = 3.2, colour = "#888780", hjust = 0.5) +
  annotate("text", x = as_date("2024-07-01"), y = 18.35,
           label = "2024", size = 3.2, colour = "#888780", hjust = 0.5) +
  annotate("text", x = as_date("2025-07-01"), y = 18.35,
           label = "2025", size = 3.2, colour = "#888780", hjust = 0.5) +
  scale_colour_manual(values = palette_region) +
  scale_fill_manual(values = palette_region) +
  scale_linetype_manual(values = c("Brazil (BRA)" = "dashed", "Europe (EUR)" = "solid")) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels = "%b '%y",
    expand = expansion(add = 10)
  ) +
  scale_y_continuous(
    labels = \(x) paste0(x, " min"),
    limits = c(10, 19),
    breaks = seq(10, 18, 2),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  guides(
    fill = "none",
    linetype = "none",
    colour = guide_legend(title = NULL)
  ) +
  labs(x = "month", y = "Taxi-out time (min)")

p_base
```

## Move Toward Final Styling

The final styling keeps the timeline quiet, with soft gridlines and an in-panel legend.

```{r}
#| label: styled-main-plot
#| fig-width: 10
#| fig-height: 5

theme_txxt <- function() {
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_line(colour = "#E6E3DA", linewidth = 0.35),
      panel.grid.major = element_line(colour = "#D9D5CB", linewidth = 0.45),
      axis.title.x = element_text(size = 11),
      axis.title.y = element_text(size = 11),
      axis.text = element_text(colour = "#4A4945"),
      legend.text = element_text(size = 10),
      plot.margin = margin(0, 6, 0, 6)
    )
}

p_main <- p_base +
  theme_txxt() +
  theme(
    legend.position = c(0.88, 0.48),
    legend.justification = c(1, 0.5),
    legend.direction = "vertical",
    legend.background = element_rect(fill = scales::alpha("white", 0.75), colour = NA),
    legend.key = element_rect(fill = NA, colour = NA),
    legend.margin = margin(2, 2, 2, 2)
  )

p_main
```

## Key Message Panels

The small panels turn the result into a submission-ready read: Europe edges upward, Brazil
is broadly steady, and the variability signal is not a single clean directional change.

```{r}
#| label: key-message-panels
#| fig-width: 10
#| fig-height: 2

key_tile <- function(value, label, sublabel, accent_col) {
  ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "#F1EFE8", colour = NA) +
    annotate("text", x = 0.07, y = 0.72,
             label = value, size = 6.2, fontface = "bold",
             colour = accent_col, hjust = 0, vjust = 0.5) +
    annotate("text", x = 0.07, y = 0.42,
             label = label, size = 4.2,
             colour = "#444441", hjust = 0, vjust = 0.5) +
    annotate("text", x = 0.07, y = 0.20,
             label = sublabel, size = 3.6,
             colour = "#888780", hjust = 0, vjust = 0.5) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
    theme_void() +
    theme(
      plot.margin = margin(0, 0, 0, 0),
      plot.background = element_rect(fill = NA, colour = NA)
    )
}

tile_eur <- key_tile(eur_delta, "EUR taxi-out change", "2023 avg → 2025 avg", col_eur)
tile_bra <- key_tile(bra_delta, "BRA taxi-out change", "2023 avg → 2025 avg", col_bra)
tile_iqr <- key_tile(iqr_label, "IQR direction", "predictability roughly flat", "#5F5E5A")

p_tiles <- tile_eur + tile_bra + tile_iqr +
  plot_layout(ncol = 3) &
  theme(plot.margin = margin(0, 8, 0, 8))

p_tiles
```

## Final Assembly

```{r}
#| label: final-assembly
#| fig-width: 10
#| fig-height: 7

width <- 10
height <- 7
dpi <- 300
outname <- "figures/day20-global-change.png"

our_title <- paste0(
  "<span style='font-size:12pt;'><b>Chart-Challenge Day 20 - Global Change (Timeline)</b></span><br>",
  "<span style='font-size:18pt;'><b>Taxi-out time: a quiet drift upward in Europe</b></span>"
)

our_subtitle <- paste0(
  "A small but persistent divergence emerges: ",
  "<b>Europe drifts upward, Brazil remains broadly steady</b>. ",
  "The chart shows the median monthly taxi-out times across 12 airports per region. ",
  "The shaded bands show the interquartile range (IQR) of daily airport medians. ",
  "We observe a gentle drift upward in Europe, suggesting steadily increasing constraints ",
  "and pressure on the departure flow."
)

our_caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Data: Brazil-Europe Comparison Report, available at: ",
  "<span style='color:#4A4945;'>github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge</span>"
)

p_final <- p_tiles / p_main +
  plot_layout(heights = c(1.15, 5)) +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 16,
        lineheight = 1.2,
        margin = margin(b = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = element_textbox_simple(
        size = 11,
        colour = "#2F2F2F",
        lineheight = 1.15,
        margin = margin(b = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.caption = element_textbox_simple(
        size = 9,
        colour = "#4F4E49",
        halign = 0,
        lineheight = 1.1,
        margin = margin(t = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.margin = margin(12, 12, 12, 12)
    )
  ) &
  theme(plot.background = element_rect(fill = "white", colour = NA))

p_final
```

## Save And Check

```{r}
#| label: save-final

ggsave(
  filename = outname,
  plot = p_final,
  width = width,
  height = height,
  dpi = dpi,
  bg = "white"
)
```

```{r}
#| label: check-final
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname)
```
