---
title: "Day 27 - animation"
format: html
---

## Aim

For Day 27 we use animation to show the **daily rhythm** of the study-airport networks rather than only their aggregate structure.

The chart compares Brazil and Europe side by side on the **same UTC clock**. That turns time-of-day into the main story: even when the same kind of airports are compared, the network does not wake up at the same UTC hour.

To keep the first implementation feasible, we deliberately make three simplifying choices:

1. `July 2025` only
2. study-airport to study-airport flights only
3. bidirectional airport-pair corridors, ranked and filtered to the busiest links

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(arrow)
library(magick)
library(patchwork)
library(scales)

source("./R/rqutils-zip.R")
```

## Build Cached July Inputs

```{r}
#| label: build-cache
#| message: false
#| warning: false

bra_apts <- c("SBBR", "SBGR", "SBSP", "SBKP", "SBRJ", "SBGL", "SBCF", "SBSV", "SBPA", "SBCT", "SBRF", "SBEG")
eur_apts <- c("EGLL", "EGKK", "EHAM", "EDDF", "EDDM", "LSZH", "LFPG", "LEMD", "LEBL", "LPPT", "LTFM", "LGAV")

cache_file <- "data/day27-animation-cache.rds"
cache_version <- 2L
days_in_july <- 31
keep_top_n <- 12

cache_ok <- FALSE

if (file.exists(cache_file)) {
  old_cache <- readRDS(cache_file)
  cache_ok <- is.list(old_cache) &&
    identical(old_cache$cache_version, cache_version) &&
    identical(old_cache$keep_top_n, keep_top_n)
}

if (!cache_ok) {
  path_nm_flt <- here::here(here::here() |> dirname(), "__DATA", "NM-flight-table")

  eur_july <- read_zip_content(
    path_nm_flt,
    "NM-flt-2025.zip",
    files = "NM-FLT-2025-Q3.parquet",
    .type = "parquet"
  ) |>
    transmute(
      region = "Europe",
      adep = ADEP,
      ades = ADES,
      dep_time_utc = as.POSIXct(AOBT_3, tz = "UTC")
    ) |>
    filter(
      month(dep_time_utc) == 7,
      adep %in% eur_apts,
      ades %in% eur_apts,
      adep != ades
    )

  bra_july <- read_parquet("data-src/totalbr.parquet") |>
    transmute(
      region = "Brazil",
      adep = co_addep,
      ades = co_addes,
      dep_time_utc = force_tz(dh_inicio, tzone = "UTC"),
      year,
      month
    ) |>
    filter(
      year == 2025,
      month == 7,
      adep %in% bra_apts,
      ades %in% bra_apts,
      adep != ades
    ) |>
    select(region, adep, ades, dep_time_utc)

  july_flights <- bind_rows(bra_july, eur_july) |>
    mutate(
      hour_utc = hour(dep_time_utc),
      node_1 = pmin(adep, ades),
      node_2 = pmax(adep, ades),
      corridor = paste(node_1, node_2, sep = " <-> ")
    )

  corridor_totals <- july_flights |>
    count(region, node_1, node_2, corridor, name = "july_flights") |>
    group_by(region) |>
    arrange(desc(july_flights), .by_group = TRUE) |>
    mutate(corridor_rank = row_number()) |>
    ungroup()

  corridor_hourly <- july_flights |>
    count(region, hour_utc, node_1, node_2, corridor, name = "flights_hour") |>
    left_join(corridor_totals, by = c("region", "node_1", "node_2", "corridor")) |>
    filter(corridor_rank <= keep_top_n) |>
    mutate(avg_daily_hour = flights_hour / days_in_july)

  airport_hourly <- july_flights |>
    select(region, hour_utc, adep, ades) |>
    pivot_longer(cols = c(adep, ades), values_to = "icao", names_to = "role") |>
    count(region, hour_utc, icao, name = "movements_hour") |>
    mutate(avg_daily_hour = movements_hour / days_in_july)

  saveRDS(
    list(
      cache_version = cache_version,
      corridor_hourly = corridor_hourly,
      airport_hourly = airport_hourly,
      keep_top_n = keep_top_n,
      days_in_july = days_in_july
    ),
    cache_file
  )
}

anim_cache <- readRDS(cache_file)
corridor_hourly <- anim_cache$corridor_hourly
airport_hourly <- anim_cache$airport_hourly
keep_top_n <- anim_cache$keep_top_n
days_in_july <- anim_cache$days_in_july
```

## Prepare Plot Tables

```{r}
#| label: prepare-plot-tables

pal_region <- c("Brazil" = "#0F6E56", "Europe" = "#185FA5")

corridor_totals_top <- corridor_hourly |>
  distinct(region, corridor, july_flights, corridor_rank) |>
  arrange(region, corridor_rank) |>
  mutate(
    corridor_label = paste0(corridor, "  (", comma(july_flights), " Jul flts)")
  ) |>
  group_by(region) |>
  mutate(
    corridor_key = paste(region, corridor_label, sep = "___"),
    corridor_key = factor(corridor_key, levels = rev(corridor_key))
  ) |>
  ungroup()

heatmap_base <- corridor_hourly |>
  left_join(
    corridor_totals_top |>
      select(region, corridor, corridor_label, corridor_key),
    by = c("region", "corridor")
  ) |>
  group_by(region, corridor_key, corridor_label) |>
  complete(
    hour_utc = 0:23,
    fill = list(
      flights_hour = 0,
      avg_daily_hour = 0
    )
  ) |>
  ungroup() |>
  mutate(
    corridor_label = stringr::str_remove(as.character(corridor_key), "^[^_]+___")
  ) |>
  mutate(
    region = factor(region, levels = c("Brazil", "Europe"))
  )

frame_summary <- heatmap_base |>
  summarise(
    avg_daily_flights = sum(avg_daily_hour),
    active_corridors = sum(avg_daily_hour > 0),
    .by = c(region, hour_utc)
  )

hour_profile <- heatmap_base |>
  summarise(
    avg_daily_flights = sum(avg_daily_hour),
    .by = c(region, hour_utc)
  ) |>
  group_by(region) |>
  mutate(
    hour_share = avg_daily_flights / sum(avg_daily_flights, na.rm = TRUE)
  ) |>
  ungroup()

peak_labels <- hour_profile |>
  group_by(region) |>
  slice_max(avg_daily_flights, n = 1, with_ties = FALSE) |>
  ungroup()
```

## Build Animation

```{r}
#| label: render-animation
#| fig-width: 11
#| fig-height: 7.6

dir.create("figures", showWarnings = FALSE)
frame_dir <- file.path(tempdir(), "day27-animation-frames")
dir.create(frame_dir, showWarnings = FALSE)

plot_network_hour <- function(frame_hour) {
  frame_totals <- frame_summary |>
    filter(hour_utc == frame_hour) |>
    mutate(
      summary_label = paste0(
        "At ",
        sprintf("%02d:00", frame_hour),
        " UTC: ",
        comma(round(avg_daily_flights, 1)),
        " avg daily flights across top corridors"
      )
    )

  p_heat <- ggplot(
    heatmap_base,
    aes(x = hour_utc, y = corridor_key, fill = avg_daily_hour)
  ) +
    geom_tile(width = 0.92, height = 0.88, color = "#F7F4EC", linewidth = 0.15) +
    geom_rect(
      aes(
        xmin = frame_hour - 0.5,
        xmax = frame_hour + 0.5,
        ymin = -Inf,
        ymax = Inf
      ),
      inherit.aes = FALSE,
      fill = NA,
      color = "#A45A00",
      linewidth = 0.7
    ) +
    geom_vline(
      xintercept = frame_hour,
      color = "#A45A00",
      linewidth = 0.55,
      alpha = 0.95
    ) +
    geom_tile(
      data = heatmap_base |>
        filter(hour_utc == frame_hour),
      aes(x = hour_utc, y = corridor_key),
      fill = NA,
      color = "#2F241D",
      linewidth = 0.5
    ) +
    geom_text(
      data = peak_labels,
      aes(
        x = hour_utc,
        y = 1.06,
        label = paste0("Peak: ", sprintf("%02d:00", hour_utc), " UTC")
      ),
      inherit.aes = FALSE,
      hjust = 0.5,
      vjust = 0,
      size = 3.0,
      color = "#3D3A34"
    ) +
    facet_wrap(~ region, ncol = 2, scales = "free_y") +
    scale_fill_gradient(
      low = "#F4EBD9",
      high = "#A33A2B",
      labels = label_number(accuracy = 0.1),
      name = "Avg flights\nper July day"
    ) +
    scale_y_discrete(labels = \(x) stringr::str_remove(x, "^[^_]+___")) +
    scale_x_continuous(
      breaks = c(0, 3, 6, 9, 12, 15, 18, 21, 23),
      labels = \(x) sprintf("%02d", x),
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    coord_cartesian(ylim = c(0, NA), clip = "off") +
    labs(
      title = "Day 27 - Animation: corridor clocks differ on the same UTC day",
      subtitle = paste0(
        "Bidirectional study-airport pairs in July 2025. Highlighted column = ",
        sprintf("%02d:00", frame_hour),
        ":00 UTC. Keeping the top ",
        keep_top_n,
        " airport-pair corridors per region by July traffic."
      ),
      x = "UTC hour",
      y = NULL,
      caption = "DECEA Performance Section and PRU EUROCONTROL; Brazil-Europe comparison study. Corridor = bidirectional study-airport pair. Heat shows average departures per July day."
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.background = element_rect(fill = "#FFFDF8", color = NA),
      panel.background = element_rect(fill = "#FFFDF8", color = NA),
      panel.grid = element_blank(),
      panel.spacing = unit(1.2, "lines"),
      strip.text = element_text(face = "bold", size = 12),
      plot.title = element_text(face = "bold", size = 18, color = "#1F1F1F"),
      plot.subtitle = element_text(size = 11, color = "#3D3A34"),
      plot.caption = element_text(size = 9, color = "#5B5953"),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.text.y = element_text(color = "#3D3A34", size = 9),
      legend.position = "top",
      plot.margin = margin(10, 16, 10, 16)
    )

  p_profile <- ggplot(
    hour_profile,
    aes(x = hour_utc, y = hour_share)
  ) +
    geom_step(color = "#8E8A83", linewidth = 1.1, alpha = 0.9) +
    geom_point(color = "#8E8A83", size = 1.8, alpha = 0.9) +
    geom_vline(
      xintercept = frame_hour,
      color = "#A45A00",
      linewidth = 0.55,
      alpha = 0.95
    ) +
    geom_point(
      data = hour_profile |>
        filter(hour_utc == frame_hour),
      color = "#A45A00",
      size = 2.6
    ) +
    geom_text(
      data = frame_totals,
      aes(
        x = 0.1,
        y = max(hour_profile$hour_share, na.rm = TRUE) * 0.92,
        label = paste0(
          sprintf("%02d:00", frame_hour),
          " UTC: ",
          comma(round(avg_daily_flights, 1)),
          " avg flights"
        )
      ),
      inherit.aes = FALSE,
      hjust = 0,
      vjust = 1,
      size = 3.0,
      color = "#3D3A34"
    ) +
    facet_wrap(~ region, ncol = 2) +
    scale_x_continuous(
      breaks = c(0, 3, 6, 9, 12, 15, 18, 21, 23),
      labels = \(x) sprintf("%02d", x),
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(
      labels = percent_format(accuracy = 1),
      expand = expansion(mult = c(0.05, 0.10))
    ) +
    labs(
      x = "UTC hour",
      y = "% of kept corridor flights"
    ) +
    theme_minimal(base_size = 10) +
    theme(
      panel.background = element_rect(fill = "#FFFDF8", color = NA),
      plot.background = element_rect(fill = "#FFFDF8", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.text = element_blank(),
      axis.text = element_text(color = "#3D3A34"),
      axis.title = element_text(color = "#3D3A34"),
      plot.margin = margin(0, 16, 8, 16)
    )

  p_heat / p_profile + plot_layout(heights = c(8.5, 1.8))
}

frame_hours <- 0:23
frame_paths <- file.path(frame_dir, sprintf("network-heartbeat-%02d.png", frame_hours))

purrr::walk2(
  frame_hours,
  frame_paths,
  \(frame_hour, frame_path) {
    ggsave(
      filename = frame_path,
      plot = plot_network_hour(frame_hour),
      width = 11,
      height = 7.6,
      dpi = 140
    )
  }
)

animation_path <- "figures/day27-animation-network-heartbeat.gif"

network_animation <- image_read(frame_paths) |>
  image_join() |>
  image_animate(fps = 2)

image_write(network_animation, animation_path)

knitr::include_graphics(animation_path)
```

## Debrief

The rough implementation already supports the intended story:

* Brazil and Europe use the same UTC clock
* each row is a bidirectional airport-pair corridor
* the moving UTC cursor guides attention without forcing the geometry to carry the whole message

For a next pass, the main options would be:

* swap `UTC hour` for `arrival hour`
* compare `average July day` against one specific `representative week`
* add local-time annotations or a second synchronized hourly total panel
