---
title: "Day 25 - Space"
format: html
---

## Aim

Today's prompt is interpreted as a ground-capacity question.

In this airport framing:

* `space` means the physical movement area occupied during taxi-out
* `uncertainty` means the extra occupation time beyond the reference taxi-out expectation

The first sketch suggested a density cloud of taxi-out time versus traffic volume. That is a
useful starting point, but the message becomes clearer if the `y` axis isolates the unexpected
part of occupation:

`uncertainty = actual taxi-out minutes - reference taxi-out minutes`

That produces a more self-speaking question:

**When daily departure demand rises, how much longer than planned does each departure occupy
ground space?**

This analytical draft keeps two candidate views:

1. a raw occupancy field, showing the cloud of airport-days
2. a binned uncertainty frontier, showing how median and upper-tail uncertainty move with demand

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(patchwork)
library(scales)
library(ggtext)
```

## Read Taxi-Out Data

```{r}
#| label: read-txxt-data

read_txxt <- function(path, region) {
  readr::read_csv(path, show_col_types = FALSE) |>
    transmute(
      ICAO,
      PHASE,
      DATE = as_date(DATE),
      MVTS_VALID,
      TOT_TXXT,
      TOT_REF,
      TOT_ADD_TIME,
      region = region
    )
}

bra_files <- list.files("data", pattern = "PBWG-BRA-txxt-analytic-2025", full.names = TRUE)
eur_files <- list.files("data", pattern = "PBWG-EUR-txxt-analytic-2025", full.names = TRUE)

txxt_raw <- bind_rows(
  map(bra_files, read_txxt, region = "Brazil"),
  map(eur_files, read_txxt, region = "Europe")
)
```

## Build Analytical Inputs

We work with departure flows only. The core quantity is the average unexpected taxi-out time per
departure on each airport-day.

```{r}
#| label: build-inputs

txxt_daily <- txxt_raw |>
  filter(PHASE == "DEP", MVTS_VALID > 0) |>
  mutate(
    actual_txxt = TOT_TXXT / MVTS_VALID,
    ref_txxt = TOT_REF / MVTS_VALID,
    uncertainty_min = actual_txxt - ref_txxt,
    extra_space_hours = TOT_ADD_TIME / 60,
    space_demand = MVTS_VALID
  ) |>
  mutate(
    region = factor(region, levels = c("Brazil", "Europe"))
  )

movement_breaks <- seq(0, 900, by = 50)

uncertainty_frontier <- txxt_daily |>
  mutate(
    demand_band = cut(
      space_demand,
      breaks = movement_breaks,
      include.lowest = TRUE,
      right = FALSE
    )
  ) |>
  filter(!is.na(demand_band)) |>
  group_by(region, demand_band) |>
  summarise(
    demand_mid = mean(space_demand, na.rm = TRUE),
    n_days = n(),
    median_unc = median(uncertainty_min, na.rm = TRUE),
    p75_unc = quantile(uncertainty_min, 0.75, na.rm = TRUE),
    p90_unc = quantile(uncertainty_min, 0.90, na.rm = TRUE),
    median_extra_hours = median(extra_space_hours, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(n_days >= 20)

region_summary <- txxt_daily |>
  summarise(
    median_unc = median(uncertainty_min, na.rm = TRUE),
    p90_unc = quantile(uncertainty_min, 0.90, na.rm = TRUE),
    .by = region
  )
```

## Candidate 1: Raw Occupancy Field

This preserves the original idea of a cloud, but makes the vertical axis the uncertainty gap
rather than the full taxi-out time.

```{r}
#| label: candidate-raw-field
#| fig-width: 10
#| fig-height: 5.5

pal_region <- c("Brazil" = "#0F6E56", "Europe" = "#185FA5")

p_field <- ggplot(txxt_daily, aes(x = space_demand, y = uncertainty_min)) +
  geom_bin2d(bins = 28) +
  geom_hline(yintercept = 0, linewidth = 0.45, linetype = "22", colour = "#9A9488") +
  facet_wrap(~ region, ncol = 2) +
  scale_fill_gradient(
    low = "#E8E2D3",
    high = "#8D5A2B",
    labels = label_number()
  ) +
  scale_x_continuous(
    breaks = seq(0, 900, 100),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    labels = \(x) paste0(round(x, 1), " min"),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  labs(
    title = "Candidate 1: where uncertain occupation clusters",
    subtitle = "Each tile is an airport-day in 2025 departures only; y shows extra taxi-out minutes above the reference expectation",
    x = "Departure movements per day",
    y = "Unexpected occupation time per departure",
    fill = "Airport-days"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "#DDD7C9", linewidth = 0.4),
    strip.text = element_text(face = "bold"),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14, colour = "#1C1B18"),
    plot.margin = margin(8, 8, 8, 8)
  )

p_field
```

## Candidate 2: Binned Uncertainty Frontier

This version makes the story more explicit. The line tracks the median uncertainty at each demand
level; the ribbon shows the upper tail where space occupation becomes notably less predictable.

```{r}
#| label: candidate-frontier
#| fig-width: 10
#| fig-height: 5.8

p_frontier <- ggplot(
  uncertainty_frontier,
  aes(x = demand_mid, y = median_unc, colour = region, fill = region)
) +
  geom_ribbon(
    aes(ymin = median_unc, ymax = p90_unc),
    alpha = 0.18,
    colour = NA,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 1.2, show.legend = FALSE) +
  geom_point(aes(size = n_days), alpha = 0.95, show.legend = FALSE) +
  geom_text(
    data = uncertainty_frontier |>
      group_by(region) |>
      slice_max(demand_mid, n = 1, with_ties = FALSE) |>
      ungroup(),
    aes(label = region),
    hjust = -0.15,
    fontface = "bold",
    size = 3.7,
    show.legend = FALSE
  ) +
  geom_hline(yintercept = 0, linewidth = 0.45, linetype = "22", colour = "#9A9488") +
  scale_colour_manual(values = pal_region) +
  scale_fill_manual(values = pal_region) +
  scale_size_continuous(range = c(2, 5)) +
  scale_x_continuous(
    breaks = seq(0, 900, 100),
    limits = c(0, 860),
    expand = expansion(mult = c(0.01, 0.08))
  ) +
  scale_y_continuous(
    labels = \(x) paste0(round(x, 1), " min"),
    expand = expansion(mult = c(0.03, 0.08))
  ) +
  facet_wrap(~ region, ncol = 2) +
  labs(
    title = "Candidate 2: more departures usually mean more uncertain use of ground space",
    subtitle = "Median line plus 90th-percentile ribbon of extra taxi-out minutes above reference, grouped into demand bands",
    x = "Departure movements per day",
    y = "Unexpected occupation time per departure"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "#DDD7C9", linewidth = 0.4),
    strip.text = element_text(face = "bold"),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14, colour = "#1C1B18"),
    plot.margin = margin(8, 12, 8, 8)
  )

p_frontier
```

## Chosen Direction

For the publication pass, candidate 2 is the better chart.

Why it works better:

* it encodes the prompt directly: demand on the `x` axis, uncertainty on the `y` axis
* it separates the central tendency from the disruption tail
* it reads as a capacity story rather than as a generic point cloud
* it still lets us compare Brazil and Europe without overloading the page

The most promising headline is:

**Ground space becomes less predictable as departure demand rises, and Europe's upper tail is
consistently thicker.**

## Publication Metrics

For the publication version we summarize three signals:

* the typical uncertainty per departure in each region
* the thicker European disruption tail
* the system-wide amount of extra occupied space on the busiest demand bands

```{r}
#| label: publication-metrics

tile_stats <- region_summary |>
  mutate(
    median_label = paste0(round(median_unc, 1), " min"),
    p90_label = paste0(round(p90_unc, 1), " min")
  )

busiest_band <- uncertainty_frontier |>
  group_by(region) |>
  slice_max(demand_mid, n = 1, with_ties = FALSE) |>
  ungroup() |>
  mutate(
    extra_hr_label = paste0(round(median_extra_hours), " hr"),
    demand_label = paste0(round(demand_mid), " dep/day")
  )

bra_med_label <- tile_stats |>
  filter(region == "Brazil") |>
  pull(median_label)

eur_med_label <- tile_stats |>
  filter(region == "Europe") |>
  pull(median_label)

eur_p90_label <- tile_stats |>
  filter(region == "Europe") |>
  pull(p90_label)

eur_busy_label <- busiest_band |>
  filter(region == "Europe") |>
  pull(extra_hr_label)

eur_busy_demand <- busiest_band |>
  filter(region == "Europe") |>
  pull(demand_label)
```

## Final Plot

```{r}
#| label: final-plot
#| fig-width: 10
#| fig-height: 5.8

p_main <- ggplot(
  uncertainty_frontier,
  aes(x = demand_mid, y = median_unc, colour = region, fill = region)
) +
  geom_ribbon(
    aes(ymin = median_unc, ymax = p90_unc),
    alpha = 0.18,
    colour = NA,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 1.25, show.legend = FALSE) +
  geom_point(aes(size = n_days), alpha = 0.98, show.legend = FALSE) +
  geom_text(
    data = uncertainty_frontier |>
      group_by(region) |>
      slice_max(demand_mid, n = 1, with_ties = FALSE) |>
      ungroup(),
    aes(x = demand_mid - 28, label = region),
    hjust = 0,
    fontface = "bold",
    size = 3.9,
    show.legend = FALSE
  ) +
  geom_hline(yintercept = 0, linewidth = 0.45, linetype = "22", colour = "#9A9488") +
  annotate(
    "label",
    x = 530,
    y = 6.8,
    label = "shaded band = median to 90th percentile\nof unexpected occupation time",
    size = 3.2,
    label.size = 0,
    fill = alpha("white", 0.88),
    colour = "#4D4B47"
  ) +
  scale_colour_manual(values = pal_region) +
  scale_fill_manual(values = pal_region) +
  scale_size_continuous(range = c(2.3, 5.3)) +
  scale_x_continuous(
    breaks = seq(0, 900, 100),
    limits = c(0, 860),
    expand = expansion(mult = c(0.01, 0.08))
  ) +
  scale_y_continuous(
    labels = \(x) paste0(round(x, 1), " min"),
    limits = c(0, 8.2),
    breaks = seq(0, 8, 2),
    expand = expansion(mult = c(0.01, 0.03))
  ) +
  facet_wrap(~ region, ncol = 2) +
  labs(
    x = "Departure movements per day",
    y = "Unexpected occupation time per departure"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "#D9D5CB", linewidth = 0.45),
    strip.text = element_text(face = "bold", size = 12),
    axis.text = element_text(colour = "#45433E"),
    axis.title = element_text(colour = "#262522"),
    plot.margin = margin(0, 12, 0, 8)
  )

p_main
```

## Message Row

```{r}
#| label: message-row
#| fig-width: 10
#| fig-height: 2

key_tile <- function(value, label, sublabel, accent_col) {
  ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "#F1EFE8", colour = NA) +
    annotate("text", x = 0.07, y = 0.72,
             label = value, size = 6.2, fontface = "bold",
             colour = accent_col, hjust = 0, vjust = 0.5) +
    annotate("text", x = 0.07, y = 0.42,
             label = label, size = 4.0,
             colour = "#444441", hjust = 0, vjust = 0.5) +
    annotate("text", x = 0.07, y = 0.20,
             label = sublabel, size = 3.45,
             colour = "#888780", hjust = 0, vjust = 0.5) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
    theme_void() +
    theme(
      plot.margin = margin(0, 0, 0, 0),
      plot.background = element_rect(fill = NA, colour = NA)
    )
}

tile_bra <- key_tile(
  bra_med_label,
  "Brazil median uncertainty",
  "extra taxi-out minutes per departure in 2025",
  pal_region[["Brazil"]]
)

tile_eur <- key_tile(
  eur_med_label,
  "Europe median uncertainty",
  "higher baseline unpredictability",
  pal_region[["Europe"]]
)

tile_tail <- key_tile(
  eur_busy_label,
  "Busiest European demand band",
  "median extra occupied ground-time",
  "#7A5C2E"
)

p_tiles <- tile_bra + tile_eur + tile_tail +
  plot_layout(ncol = 3) &
  theme(plot.margin = margin(0, 8, 0, 8))

p_tiles
```

## Final Assembly

```{r}
#| label: final-assembly
#| fig-width: 10
#| fig-height: 7

width <- 10
height <- 7
dpi <- 300
outname_final <- "figures/day25-space-uncertainty.png"

our_title <- paste0(
  "<span style='font-size:12pt;'><b>Chart-Challenge Day 25 - Space</b></span><br>",
  "<span style='font-size:18pt;'><b>Busy departure banks make ground-space occupation less predictable</b></span>"
)

our_subtitle <- paste0(
  "Here, <b>space</b> means the runway and apron system occupied during taxi-out, and <b>uncertainty</b> means ",
  "the extra taxi-out time above the reference expectation. ",
  "Across the 2025 study airports, unexpected occupation usually rises with departure demand. ",
  "<b>Europe starts from a higher baseline and carries the thicker upper tail</b>, implying more days when the ground system stays occupied longer than planned."
)

our_caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Measure shown: daily departure taxi-out uncertainty = actual minus reference taxi-out minutes per departure. ",
  "Point size reflects the number of airport-days in each demand band."
)

p_final <- p_tiles / p_main +
  plot_layout(heights = c(1.15, 5)) +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 16,
        lineheight = 1.2,
        margin = margin(b = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = element_textbox_simple(
        size = 11,
        colour = "#2F2F2F",
        lineheight = 1.15,
        margin = margin(b = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.caption = element_textbox_simple(
        size = 9,
        colour = "#4F4E49",
        halign = 0,
        lineheight = 1.1,
        margin = margin(t = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.margin = margin(12, 12, 12, 12)
    )
  ) &
  theme(plot.background = element_rect(fill = "white", colour = NA))

p_final
```

## Save Rough Figure

```{r}
#| label: save-rough-figure

outname <- "figures/day25-space-uncertainty-rough.png"

ggsave(
  filename = outname,
  plot = p_frontier,
  width = 10,
  height = 5.8,
  dpi = 300,
  bg = "white"
)
```

```{r}
#| label: show-rough-figure
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname)
```

## Save And Check Final

```{r}
#| label: save-final

ggsave(
  filename = outname_final,
  plot = p_final,
  width = width,
  height = height,
  dpi = dpi,
  bg = "white"
)
```

```{r}
#| label: check-final
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname_final)
```
