---
title: "Day 26 - Trend"
format: html
---

## Aim

Today's prompt is read very directly: show a meaningful operational trend in the 12-airport
Brazil-Europe study set.

The uncertainties rubric offers a useful hook, but it does not need to dominate the chart. We use
it simply to define the metric:

`uncertainty = actual taxi-out minutes - reference taxi-out minutes`

That gives us a clear trend question:

**From 2023 to 2025, did taxi-out predictability improve or deteriorate across the study
airports?**

This rough draft tests two chart types:

1. a regional time trend with airport texture
2. an airport-level trend comparison from 2023 to 2025

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
```

## Read Taxi-Out Data

```{r}
#| label: read-txxt

read_txxt <- function(path, region) {
  readr::read_csv(path, show_col_types = FALSE) |>
    transmute(
      ICAO,
      PHASE,
      DATE = as_date(DATE),
      MVTS_VALID,
      TOT_TXXT,
      TOT_REF,
      region = region
    )
}

bra_files <- list.files("data", pattern = "PBWG-BRA-txxt-analytic-202[3-5]", full.names = TRUE)
eur_files <- list.files("data", pattern = "PBWG-EUR-txxt-analytic-202[3-5]", full.names = TRUE)

txxt_raw <- bind_rows(
  map(bra_files, read_txxt, region = "Brazil"),
  map(eur_files, read_txxt, region = "Europe")
)
```

## Build Inputs

```{r}
#| label: build-inputs

txxt_daily <- txxt_raw |>
  filter(PHASE == "DEP", MVTS_VALID > 0) |>
  mutate(
    uncertainty_min = (TOT_TXXT - TOT_REF) / MVTS_VALID,
    month = floor_date(DATE, "month")
  ) |>
  mutate(region = factor(region, levels = c("Brazil", "Europe")))

airport_names <- tribble(
  ~ICAO, ~AIRPORT_NAME,
  "SBBR", "Brasilia",
  "SBCF", "Belo Horizonte",
  "SBCT", "Curitiba",
  "SBEG", "Manaus",
  "SBGL", "Rio Galeao",
  "SBGR", "Sao Paulo GRU",
  "SBKP", "Campinas",
  "SBPA", "Porto Alegre",
  "SBRF", "Recife",
  "SBRJ", "Rio Santos Dumont",
  "SBSP", "Sao Paulo Congonhas",
  "SBSV", "Salvador",
  "EDDF", "Frankfurt",
  "EDDM", "Munich",
  "EGKK", "London Gatwick",
  "EGLL", "London Heathrow",
  "EHAM", "Amsterdam",
  "LEBL", "Barcelona",
  "LEMD", "Madrid",
  "LFPG", "Paris CDG",
  "LGAV", "Athens",
  "LPPT", "Lisbon",
  "LSZH", "Zurich",
  "LTFM", "Istanbul"
)

monthly_airport <- txxt_daily |>
  summarise(
    uncertainty_min = median(uncertainty_min, na.rm = TRUE),
    .by = c(region, ICAO, month)
  ) |>
  arrange(region, ICAO, month)

monthly_region <- monthly_airport |>
  summarise(
    med_unc = median(uncertainty_min, na.rm = TRUE),
    q25_unc = quantile(uncertainty_min, 0.25, na.rm = TRUE),
    q75_unc = quantile(uncertainty_min, 0.75, na.rm = TRUE),
    .by = c(region, month)
  ) |>
  arrange(region, month)

airport_trend <- monthly_airport |>
  mutate(year = year(month)) |>
  filter(year %in% c(2023, 2025)) |>
  summarise(
    uncertainty_min = mean(uncertainty_min, na.rm = TRUE),
    .by = c(region, ICAO, year)
  ) |>
  mutate(year_lab = factor(year, levels = c(2023, 2025))) |>
  select(-year) |>
  pivot_wider(names_from = year_lab, values_from = uncertainty_min) |>
  rename(unc_2023 = `2023`, unc_2025 = `2025`) |>
  mutate(
    delta = unc_2025 - unc_2023
  ) |>
  group_by(region) |>
  arrange(delta, .by_group = TRUE) |>
  mutate(
    airport_ord = factor(ICAO, levels = ICAO)
  ) |>
  ungroup() |>
  left_join(airport_names, by = "ICAO") |>
  mutate(
    airport_label = paste0(AIRPORT_NAME, "\n", ICAO)
  )

trend_summary <- airport_trend |>
  summarise(
    n_up = sum(delta > 0, na.rm = TRUE),
    med_delta = median(delta, na.rm = TRUE),
    .by = region
  )
```

## Candidate 1: Time Trend With Airport Texture

This version treats trend literally as a monthly path. The faint lines preserve airport-level
texture while the darker regional line shows the median airport.

```{r}
#| label: candidate-1
#| fig-width: 10
#| fig-height: 5.5

pal_region <- c("Brazil" = "#0F6E56", "Europe" = "#185FA5")

p_time <- ggplot(monthly_airport, aes(x = month, y = uncertainty_min, group = ICAO)) +
  geom_line(colour = "#A9AAA6", alpha = 0.22, linewidth = 0.45) +
  geom_ribbon(
    data = monthly_region,
    aes(x = month, ymin = q25_unc, ymax = q75_unc, fill = region, group = region),
    inherit.aes = FALSE,
    alpha = 0.16,
    colour = NA
  ) +
  geom_line(
    data = monthly_region,
    aes(x = month, y = med_unc, colour = region, group = region),
    inherit.aes = FALSE,
    linewidth = 1.2
  ) +
  facet_wrap(~ region, ncol = 2) +
  scale_colour_manual(values = pal_region, guide = "none") +
  scale_fill_manual(values = pal_region, guide = "none") +
  scale_x_date(
    date_breaks = "6 months",
    date_labels = "%b\n%Y"
  ) +
  scale_y_continuous(
    labels = \(x) paste0(round(x, 1), " min")
  ) +
  labs(
    title = "Candidate 1: the regional trend is down in Brazil and up in Europe",
    subtitle = "Monthly median uncertainty by airport; dark line shows the regional median and ribbon the airport IQR",
    x = NULL,
    y = "Unexpected taxi-out time per departure"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14)
  )

p_time
```

## Candidate 2: Airport-Level Trend Scorecard

This version makes trend more legible at airport level. Each segment shows the direction and
magnitude of change between the 2023 and 2025 airport averages.

```{r}
#| label: candidate-2
#| fig-width: 10
#| fig-height: 6.5

airport_long <- airport_trend |>
  pivot_longer(
    cols = c(unc_2023, unc_2025),
    names_to = "year_lab",
    values_to = "uncertainty_min"
  ) |>
  mutate(
    year_lab = factor(
      year_lab,
      levels = c("unc_2023", "unc_2025"),
      labels = c("2023 avg", "2025 avg")
    )
  )

p_score <- ggplot(airport_long, aes(x = uncertainty_min, y = airport_ord, group = ICAO)) +
  geom_segment(
    data = airport_trend,
    aes(x = unc_2023, xend = unc_2025, y = airport_ord, yend = airport_ord, colour = region),
    linewidth = 1.1,
    alpha = 0.75,
    inherit.aes = FALSE
  ) +
  geom_point(
    aes(shape = year_lab, fill = region, colour = region),
    size = 3.1,
    stroke = 0.9
  ) +
  geom_vline(xintercept = 0, linewidth = 0.4, linetype = "22", colour = "#A8A39A") +
  facet_wrap(~ region, scales = "free_y", ncol = 2) +
  scale_colour_manual(values = pal_region, guide = "none") +
  scale_fill_manual(values = pal_region, guide = "none") +
  scale_shape_manual(values = c(21, 24)) +
  scale_x_continuous(
    labels = \(x) paste0(round(x, 1), " min")
  ) +
  labs(
    title = "Candidate 2: almost every European airport trends up, while most Brazilian airports trend down",
    subtitle = "Average monthly additional taxi-out time in 2023 versus 2025",
    x = "Average additional taxi-out time per departure",
    y = NULL,
    shape = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "top"
  )

p_score
```

## Chosen Direction

Candidate 2 is the more promising rough direction.

Why it fits better:

* the prompt is about `trend`, and the airport segments show trend directly
* the contrast between regions is unusually clean
* the uncertainty rubric stays present but does not need heavy explanation
* the airport-level structure adds more value than another regional time series alone

Current one-sentence takeaway:

**From 2023 to 2025, taxi-out uncertainty trends upward at nearly every European study airport,
while most Brazilian study airports trend downward.**

## Publication Metrics

```{r}
#| label: publication-metrics

summary_tiles <- airport_trend |>
  summarise(
    n_up = sum(delta > 0, na.rm = TRUE),
    n_down = sum(delta < 0, na.rm = TRUE),
    med_delta = median(delta, na.rm = TRUE),
    mean_delta = mean(delta, na.rm = TRUE),
    .by = region
  ) |>
  mutate(
    count_up_label = paste0(n_up, " of 12"),
    count_down_label = paste0(n_down, " of 12"),
    delta_label = paste0(
      if_else(mean_delta > 0, "+", ""),
      number(mean_delta, accuracy = 0.1),
      " min"
    )
  )

top_changes <- airport_trend |>
  group_by(region) |>
  slice_max(abs(delta), n = 2, with_ties = FALSE) |>
  ungroup() |>
  mutate(
    label = paste0(
      ICAO, " ",
      if_else(delta > 0, "+", ""),
      number(delta, accuracy = 0.1),
      " min"
    ),
    x_lab = unc_2025 + if_else(delta > 0, 0.18, -0.18),
    hjust_lab = if_else(delta > 0, 0, 1)
  )

bra_count_label <- summary_tiles |>
  filter(region == "Brazil") |>
  pull(count_down_label)

eur_count_label <- summary_tiles |>
  filter(region == "Europe") |>
  pull(count_up_label)

bra_delta_label <- summary_tiles |>
  filter(region == "Brazil") |>
  pull(delta_label)

eur_delta_label <- summary_tiles |>
  filter(region == "Europe") |>
  pull(delta_label)
```

## Final Plot

```{r}
#| label: final-plot
#| fig-width: 10
#| fig-height: 6.8

airport_long_final <- airport_trend |>
  pivot_longer(
    cols = c(unc_2023, unc_2025),
    names_to = "year_lab",
    values_to = "uncertainty_min"
  ) |>
  mutate(
    year_lab = factor(
      year_lab,
      levels = c("unc_2023", "unc_2025"),
      labels = c("2023 avg", "2025 avg")
    ),
    direction = case_when(
      delta > 0 ~ "Up",
      delta < 0 ~ "Down",
      TRUE ~ "Flat"
    )
  )

dir_cols <- c(
  Up = "#C24A3A",
  Down = "#138A74",
  Flat = "#6E6B64"
)

p_final_main <- ggplot(airport_long_final, aes(x = uncertainty_min, y = airport_ord, group = ICAO)) +
  geom_segment(
    data = airport_trend |>
      mutate(
        direction = case_when(
          delta > 0 ~ "Up",
          delta < 0 ~ "Down",
          TRUE ~ "Flat"
        )
      ),
    aes(
      x = unc_2023,
      xend = unc_2025,
      y = airport_ord,
      yend = airport_ord,
      colour = direction
    ),
    linewidth = 1.4,
    alpha = 0.85,
    inherit.aes = FALSE,
    arrow = grid::arrow(length = unit(0.12, "inches"), type = "closed")
  ) +
  geom_point(
    aes(shape = year_lab),
    size = 3.2,
    stroke = 1,
    colour = "#1F1F1C",
    fill = "white"
  ) +
  geom_text(
    data = filter(top_changes, ICAO != "SBRJ"),
    aes(
      x = x_lab,
      y = airport_ord,
      label = label,
      hjust = hjust_lab
    ),
    size = 3.2,
    fontface = "bold",
    colour = "#3D3B37",
    inherit.aes = FALSE
  ) +
  geom_text(
    data = filter(top_changes, ICAO == "SBRJ"),
    aes(
      x = unc_2023 + 0.15,
      y = airport_ord,
      label = label
    ),
    vjust = 1.8,
    hjust = 0,
    size = 3.2,
    fontface = "bold",
    colour = "#3D3B37",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ region, scales = "free_y", ncol = 2) +
  scale_colour_manual(values = dir_cols, guide = "none") +
  scale_shape_manual(values = c(21, 24)) +
  scale_x_continuous(
    labels = \(x) paste0(number(x, accuracy = 1), " min"),
    limits = c(1.15, 5.5),
    breaks = 1:5,
    expand = expansion(mult = c(0.02, 0.08))
  ) +
  scale_y_discrete(labels = setNames(airport_trend$airport_label, airport_trend$airport_ord)) +
  labs(
    x = "Average additional taxi-out time per departure",
    y = NULL,
    shape = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "#E2DED3", linewidth = 0.45),
    panel.grid.major.x = element_line(colour = "#D7D2C6", linewidth = 0.45),
    strip.text = element_text(face = "bold", size = 12),
    axis.text = element_text(colour = "#45433E"),
    legend.position = "top",
    plot.margin = margin(0, 10, 0, 10)
  )

p_final_main
```

## Message Row

```{r}
#| label: message-row
#| fig-width: 10
#| fig-height: 2.35

key_tile <- function(value, label, sublabel, accent_col) {
  ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1, fill = "#F1EFE8", colour = NA) +
    annotate(
      "text", x = 0.07, y = 0.78, label = value,
      size = 5.8, fontface = "bold", colour = accent_col, hjust = 0, vjust = 0.5
    ) +
    annotate(
      "text", x = 0.07, y = 0.46, label = label,
      size = 3.8, colour = "#444441", hjust = 0, vjust = 0.5, lineheight = 0.95
    ) +
    annotate(
      "text", x = 0.07, y = 0.13, label = sublabel,
      size = 3.0, colour = "#888780", hjust = 0, vjust = 0.5, lineheight = 0.95
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
    theme_void() +
    theme(
      plot.margin = margin(0, 0, 0, 0),
      plot.background = element_rect(fill = NA, colour = NA)
    )
}

tile_bra <- key_tile(
  bra_count_label,
  "Brazil airports trend down",
  paste0("mean shift ", bra_delta_label, "\nfrom 2023 to 2025"),
  "#138A74"
)

tile_eur <- key_tile(
  eur_count_label,
  "Europe airports trend up",
  paste0("mean shift ", eur_delta_label, "\nfrom 2023 to 2025"),
  "#C24A3A"
)

tile_note <- key_tile(
  "Arrow = trend",
  "2023 circle to\n2025 triangle",
  "rightward means less\npredictable by 2025",
  "#6E6B64"
)

p_tiles <- tile_bra + tile_eur + tile_note +
  plot_layout(ncol = 3) &
  theme(plot.margin = margin(0, 8, 0, 8))

p_tiles
```

## Final Assembly

```{r}
#| label: final-assembly
#| fig-width: 10
#| fig-height: 8.3

width <- 10
height <- 8.3
dpi <- 300
outname_final <- "figures/day26-trend.png"

our_title <- paste0(
  "<span style='font-size:12pt;'><b>Chart-Challenge Day 26 - Trend</b></span><br>",
  "<span style='font-size:18pt;'><b>Additional taxi-out time trends in opposite directions across the two study regions</b></span>"
)

our_subtitle <- paste0(
  "Each airport is shown twice: its <b>2023 average</b> and its <b>2025 average</b> additional taxi-out time per departure, ",
  "where additional means actual taxi-out time minus reference taxi-out time. ",
  "<b>Most Brazilian airports moved left</b>, indicating more predictable taxi-out by 2025, while ",
  "<b>almost every European airport moved right</b>, indicating the opposite trend."
)

our_caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Study set: 12 airports in Brazil and 12 airports in Europe. ",
  "Metric: average additional taxi-out time. ",
  "<span style='color:#4A4945;'>github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge</span>"
)

p_final <- p_tiles / p_final_main +
  plot_layout(heights = c(1.45, 5.6)) +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = ggtext::element_textbox_simple(
        size = 16,
        lineheight = 1.2,
        margin = margin(b = 18),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = ggtext::element_textbox_simple(
        size = 11,
        colour = "#2F2F2F",
        lineheight = 1.15,
        margin = margin(b = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.caption = ggtext::element_textbox_simple(
        size = 9,
        colour = "#4F4E49",
        halign = 0,
        lineheight = 1.1,
        margin = margin(t = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.margin = margin(12, 12, 12, 12)
    )
  ) &
  theme(plot.background = element_rect(fill = "white", colour = NA))

p_final
```

## Save Rough Figure

```{r}
#| label: save-rough

outname <- "figures/day26-trend-rough.png"

ggsave(
  filename = outname,
  plot = p_score,
  width = 10,
  height = 6.5,
  dpi = 300,
  bg = "white"
)
```

```{r}
#| label: show-rough
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname)
```

## Save And Check Final

```{r}
#| label: save-final

ggsave(
  filename = outname_final,
  plot = p_final,
  width = width,
  height = height,
  dpi = dpi,
  bg = "white"
)
```

```{r}
#| label: check-final
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname_final)
```
