---
title: "Day 28 - modelling"
format: html
---

30 Day Chart Challenge — Day: "Modelling"

idea: Additional Taxi-Out Time: The Model Beneath the KPI
data expected: DATE, AIRPORT, ACTUAL_TXOT, REF_TXOT, ADD_TXOT
                covering 2023, 2024, 2025
                12 airports: 6 Brazilian (SB*), 6 European

```{r}
# -----------------------------------------------------------------------------
# 0. Libraries & Fonts
# -----------------------------------------------------------------------------
library(tidyverse)
library(ggdist)
library(fitdistrplus)
library(ggtext)
library(patchwork)
library(showtext)

font_add_google("DM Sans", "dm_sans")
font_add_google("DM Mono", "dm_mono")
showtext_auto()
showtext_opts(dpi = 300)

# -----------------------------------------------------------------------------
# 1. Load & prep REAL data
# -----------------------------------------------------------------------------

df <- fs::dir_ls("./data", regexp = "-txxt") |> 
  purrr::map(.f = ~ readr::read_csv(.x, show_col_types = FALSE)) |>
  bind_rows() |> 
  mutate(
    #DATE = as.Date(DATE),
    AIRPORT = ICAO,
    ACTUAL_TXOT = TOT_TXXT / MVTS_VALID,
    REF_TXOT    = TOT_REF  / MVTS_VALID,
    ADD_TXOT    = TOT_ADD_TIME / MVTS_VALID
  ) |>
  filter(PHASE == "DEP") |>
  mutate(
    REGION = case_when(
      grepl("^SB", AIRPORT) ~ "Brazil",
      TRUE ~ "Europe"
    )
  )

# -----------------------------------------------------------------------------
# 2. Airport metadata 
# -----------------------------------------------------------------------------

airport_meta <- tibble(
  AIRPORT = c(
    "SBBR","SBCF","SBCT","SBEG","SBGL","SBGR",
    "SBKP","SBPA","SBRF","SBRJ","SBSP","SBSV",
    "EDDF","EDDM","EGKK","EGLL","EHAM","LEBL",
    "LEMD","LFPG","LGAV","LPPT","LSZH","LTFM"
  ),
  REGION = c(
    rep("Brazil", 12),
    rep("Europe", 12)
  ),
  CITY = c(
    "Brasília",
    "Belo Horizonte (Confins)",
    "Curitiba",
    "Manaus",
    "Rio de Janeiro (Galeão)",
    "São Paulo (Guarulhos)",
    "Campinas (Viracopos)",
    "Porto Alegre",
    "Recife",
    "Rio de Janeiro (Santos Dumont)",
    "São Paulo (Congonhas)",
    "Salvador",
    "Frankfurt",
    "Munich",
    "London (Gatwick)",
    "London (Heathrow)",
    "Amsterdam",
    "Barcelona",
    "Madrid",
    "Paris (CDG)",
    "Athens",
    "Lisbon",
    "Zurich",
    "Istanbul"
  )
)

df <- df |>
  left_join(airport_meta, by = c("AIRPORT", "REGION"))

# -----------------------------------------------------------------------------
# 3. Gamma fitting on ADD_TXOT
# -----------------------------------------------------------------------------
fit_gamma <- function(x) {
  x_pos <- x[x > 0.01]
  if (length(x_pos) < 30) return(NULL)
  tryCatch(
    fitdist(x_pos, "gamma", method = "mle"),
    error = function(e) NULL
  )
}

gamma_fits <- df |>
  group_by(AIRPORT, REGION, CITY) |>
  summarise(
    fit = list(fit_gamma(ADD_TXOT)),
    x_min = quantile(ADD_TXOT[ADD_TXOT > 0], 0.01, na.rm = TRUE),
    x_max = quantile(ADD_TXOT[ADD_TXOT > 0], 0.995, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(!map_lgl(fit, is.null)) |>
  mutate(
    shape = map_dbl(fit, ~ .x$estimate["shape"]),
    rate  = map_dbl(fit, ~ .x$estimate["rate"]),
    mean_add = shape / rate
  )

gamma_curves <- gamma_fits |>
  rowwise() |>
  reframe(
    AIRPORT = AIRPORT,
    REGION  = REGION,
    CITY    = CITY,
    shape   = shape,
    rate    = rate,
    mean_add = mean_add,
    x = seq(x_min, x_max, length.out = 300),
    y = dgamma(x, shape = shape, rate = rate)
  )

# -----------------------------------------------------------------------------
# 4. Palette & theme
# -----------------------------------------------------------------------------
pal_brazil <- "#C25B2A"
pal_europe <- "#2A6696"
pal_bg     <- "#F7F4EF"

theme_anp <- function() {
  theme_minimal(base_family = "dm_sans", base_size = 9) +
    theme(
      plot.background  = element_rect(fill = pal_bg, color = NA),
      panel.background = element_rect(fill = pal_bg, color = NA),
      panel.grid.major.x = element_line(color = "#DDD9D3", linewidth = 0.3),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.text.x = element_text(size = 7, color = "#6B6760", family = "dm_mono"),
      axis.text.y = element_blank(),
      axis.ticks  = element_blank(),
      strip.text  = element_text(size = 8, face = "bold", hjust = 0),
      strip.background = element_blank(),
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_markdown(size = 9),
      plot.caption = element_text(size = 7, family = "dm_mono"),
      legend.position = "none"
    )
}

region_fill  <- c(Brazil = pal_brazil, Europe = pal_europe)
region_color <- c(Brazil = pal_brazil, Europe = pal_europe)

# -----------------------------------------------------------------------------
# 5. Density data
# -----------------------------------------------------------------------------
dens_data <- df |>
  dplyr::select(AIRPORT, REGION, CITY, ACTUAL_TXOT, REF_TXOT) |>
  pivot_longer(
    cols = c(ACTUAL_TXOT, REF_TXOT),
    names_to = "metric",
    values_to = "value"
  ) |>
  filter(value > 0) |>
  group_by(metric) |>
  filter(value < quantile(value, 0.995, na.rm = TRUE)) |>
  ungroup()

mean_labels <- gamma_fits |>
  mutate(label = paste0("γ̄=", round(mean_add, 1)))

# -----------------------------------------------------------------------------
# 6. Plot
# -----------------------------------------------------------------------------
p_main <- ggplot() +

  geom_density(
    data = filter(dens_data, metric == "ACTUAL_TXOT"),
    aes(x = value, fill = REGION, color = REGION),
    alpha = 0.18, linewidth = 0.3
  ) +

  geom_density(
    data = filter(dens_data, metric == "REF_TXOT"),
    aes(x = value, color = REGION),
    fill = NA, linewidth = 0.6, linetype = "22"
  ) +

  geom_line(
    data = gamma_curves,
    aes(x = x, y = y, color = REGION),
    linewidth = 1.0
  ) +

  geom_vline(
    data = gamma_fits,
    aes(xintercept = mean_add, color = REGION),
    linewidth = 0.4, linetype = "dotted"
  ) +

  geom_text(
    data = mean_labels,
    aes(x = mean_add, y = Inf, label = label, color = REGION),
    hjust = 0, vjust = 1.5, size = 2.2, family = "dm_mono"
  ) +

  scale_fill_manual(values = region_fill) +
  scale_color_manual(values = region_color) +

  scale_x_continuous(
    name = "Minutes",
    limits = c(0, quantile(df$ACTUAL_TXOT, 0.99, na.rm = TRUE))
  ) +

  facet_wrap(~ AIRPORT) +

  labs(
    title = "The model beneath the KPI",
    subtitle = "Taxi-out time distributions (DEP) with fitted gamma on additional time",
    caption = "Source: DECEA / EUROCONTROL · 30 Day Chart Challenge",
    x = "Minutes", y = NULL
  ) +

  theme_anp()

# -----------------------------------------------------------------------------
# 7. Export
# -----------------------------------------------------------------------------
ggsave(
  "chart_challenge_day_modelling.png",
  plot = p_main,
  width = 14,
  height = 7,
  dpi = 300,
  bg = pal_bg
)

message("✓ Saved: chart_challenge_day_modelling.png")

```



Trying to get this into a more challenge type of form


```{r}
# Ensure ordering: Brazil first, then Europe
df <- df |>
  mutate(
    REGION = factor(REGION, levels = c("Brazil", "Europe"))
  )

gamma_curves <- gamma_curves |>
  mutate(REGION = factor(REGION, levels = c("Brazil", "Europe")))

gamma_fits <- gamma_fits |>
  mutate(REGION = factor(REGION, levels = c("Brazil", "Europe")))

```

```{r}
p_main <- ggplot() +

  # Actual density
  geom_density(
    data = filter(dens_data, metric == "ACTUAL_TXOT"),
    aes(x = value, fill = REGION, color = REGION),
    alpha = 0.18, linewidth = 0.3
  ) +

  # Reference density
  geom_density(
    data = filter(dens_data, metric == "REF_TXOT"),
    aes(x = value, color = REGION),
    fill = NA, linewidth = 0.6, linetype = "22"
  ) +

  # Gamma model (ADD_TXOT)
  geom_line(
    data = gamma_curves,
    aes(x = x, y = y, color = REGION),
    linewidth = 1.0
  ) +

  # Mean line
  geom_vline(
    data = gamma_fits,
    aes(xintercept = mean_add, color = REGION),
    linewidth = 0.4, linetype = "dotted"
  ) +

  # Mean label
  geom_text(
    data = mean_labels,
    aes(x = mean_add, y = Inf, label = label, color = REGION),
    hjust = 0, vjust = 1.5, size = 2.2, family = "dm_mono"
  ) +

  scale_fill_manual(values = region_fill) +
  scale_color_manual(values = region_color) +

  scale_x_continuous(
    name = "Minutes",
    limits = c(0, quantile(df$ACTUAL_TXOT, 0.99, na.rm = TRUE))
  ) +
  
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +

  # ✅ KEY CHANGE: facet by REGION (columns) + AIRPORT (rows)
  facet_wrap(
  ~ REGION + AIRPORT,
  ncol = 2,
  scales = "fixed"
) +

  labs(
    title = "The model beneath the KPI",

    subtitle = paste0(
      "Each row shows one airport; <b>Brazil (left)</b> vs <b>Europe (right)</b>.<br>",
      "Filled curves: <b>actual taxi-out</b> · dashed: <b>reference</b> · solid line: ",
      "<b>gamma model of additional time</b>.<br>",
      "<b>Key takeaway:</b> while average additional time is similar (~2–3 min), ",
      "some airports show heavier tails — indicating less predictable surface operations."
    ),

    caption = "Source: DECEA / EUROCONTROL · 30 Day Chart Challenge",
    x = "Minutes", y = NULL
  ) +

  theme_anp() +
  theme(
    strip.text.y.left = element_text(angle = 0, hjust = 0),
    strip.placement = "outside"
  ) +
  theme(
  axis.text.y = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major.y = element_blank()
)

```

```{r}
# -----------------------------------------------------------------------------
# FINAL PLOT (fixed naming issue)
# -----------------------------------------------------------------------------

# Create combined facet variable (NOT called PANEL!)
df <- df |>
  mutate(FACET_ID = paste(REGION, AIRPORT, sep = " - "))

gamma_curves <- gamma_curves |>
  mutate(FACET_ID = paste(REGION, AIRPORT, sep = " - "))

gamma_fits <- gamma_fits |>
  mutate(FACET_ID = paste(REGION, AIRPORT, sep = " - "))

mean_labels <- mean_labels |>
  mutate(FACET_ID = paste(REGION, AIRPORT, sep = " - "))

# Ensure ordering: Brazil first, then Europe
facet_order <- df |>
  distinct(REGION, AIRPORT, FACET_ID) |>
  arrange(REGION, AIRPORT) |>
  pull(FACET_ID)

df$FACET_ID <- factor(df$FACET_ID, levels = facet_order)
gamma_curves$FACET_ID <- factor(gamma_curves$FACET_ID, levels = facet_order)
gamma_fits$FACET_ID <- factor(gamma_fits$FACET_ID, levels = facet_order)
mean_labels$FACET_ID <- factor(mean_labels$FACET_ID, levels = facet_order)

# Plot
p_main <- ggplot() +

  geom_density(
    data = filter(dens_data, metric == "ACTUAL_TXOT"),
    aes(x = value, fill = REGION, color = REGION),
    alpha = 0.18, linewidth = 0.3
  ) +

  geom_density(
    data = filter(dens_data, metric == "REF_TXOT"),
    aes(x = value, color = REGION),
    fill = NA, linewidth = 0.6, linetype = "22"
  ) +

  geom_line(
    data = gamma_curves,
    aes(x = x, y = y, color = REGION),
    linewidth = 1.0
  ) +

  geom_vline(
    data = gamma_fits,
    aes(xintercept = mean_add, color = REGION),
    linewidth = 0.4, linetype = "dotted"
  ) +

  geom_text(
    data = mean_labels,
    aes(x = mean_add, y = Inf, label = label, color = REGION),
    hjust = 0, vjust = 1.5, size = 2.2, family = "dm_mono"
  ) +

  scale_fill_manual(values = region_fill) +
  scale_color_manual(values = region_color) +

  scale_x_continuous(
    name = "Minutes",
    limits = c(0, quantile(df$ACTUAL_TXOT, 0.99, na.rm = TRUE))
  ) +

  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +

  facet_wrap(
    ~ FACET_ID,
    ncol = 2,
    scales = "fixed"
  ) +

  labs(
    title = "The model beneath the KPI",
    subtitle = paste0(
      "<b>Brazil (left)</b> vs <b>Europe (right)</b> across airports.<br>",
      "Similar averages, but <b>heavier tails at some airports</b> reveal less predictable operations."
    ),
    caption = "Source: DECEA / EUROCONTROL · 30 Day Chart Challenge",
    x = "Minutes", y = NULL
  ) +

  theme_anp() +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.text = element_text(size = 7, face = "bold")
  )


p_main
```



# Export
ggsave(
  "chart_challenge_day_modelling.png",
  plot = p_main,
  width = 10,
  height = 14,
  dpi = 300,
  bg = pal_bg
)

message("✓ Saved: chart_challenge_day_modelling.png")




START ALL OVER AGAIN

```{r}
library(tidyverse)
library(fitdistrplus)

# -----------------------------------------------------------------------------
# 1. Load & prepare
# -----------------------------------------------------------------------------

df <- fs::dir_ls("./data", regexp = "-txxt") |> 
  purrr::map(.f = ~ readr::read_csv(.x, show_col_types = FALSE)) |>
  bind_rows() |> 
  mutate(
    #DATE = as.Date(DATE),
    AIRPORT = ICAO,
    ACTUAL_TXOT = TOT_TXXT / MVTS_VALID,
    REF_TXOT    = TOT_REF  / MVTS_VALID,
    ADD_TXOT    = TOT_ADD_TIME / MVTS_VALID
  ) |>
  filter(PHASE == "DEP") |>
  mutate(
    REGION = case_when(
      grepl("^SB", AIRPORT) ~ "Brazil",
      TRUE ~ "Europe"
    )
  )

# ✅ Check: we MUST have 24 combinations (12 airports × 2 regions total)
df |>
  distinct(AIRPORT, REGION) |>
  arrange(REGION, AIRPORT) |>
  print(n = 30)

# -----------------------------------------------------------------------------
# 2. Gamma fit (NO dropping airports)
# -----------------------------------------------------------------------------
fit_gamma <- function(x) {
  x_pos <- x[x > 0.05]
  x_pos <- x_pos[!is.na(x_pos)]   # ✅ remove NA explicitly
  
  if (length(x_pos) < 30) {
    return(list(shape = NA_real_, rate = NA_real_))
  }
  
  if (sd(x_pos, na.rm = TRUE) == 0) {
    return(list(shape = NA_real_, rate = NA_real_))
  }
  
  fit <- tryCatch(
    suppressWarnings(fitdist(x_pos, "gamma", method = "mle")),
    error = function(e) NULL
  )
  
  if (is.null(fit)) {
    return(list(shape = NA_real_, rate = NA_real_))
  }
  
  list(
    shape = unname(fit$estimate["shape"]),
    rate  = unname(fit$estimate["rate"])
  )
}


gamma_fits <- df |>
  group_by(AIRPORT, REGION) |>
  summarise(
    fit = list(fit_gamma(ADD_TXOT)),
    .groups = "drop"
  ) |>
  mutate(
    shape = map_dbl(fit, "shape"),
    rate  = map_dbl(fit, "rate"),
    mean_add = shape / rate
  )

# ✅ Check: NO airports dropped
gamma_fits |>
  print(n = 30)

```

```{r}

airport_order <- df |>
  distinct(AIRPORT, REGION) |>
  arrange(REGION, AIRPORT) |>
  pull(AIRPORT)

df <- df |>
  mutate(AIRPORT = factor(AIRPORT, levels = airport_order))


p <- ggplot(df, aes(x = ACTUAL_TXOT)) +
  geom_density() +
  facet_wrap(
    ~ AIRPORT,
    ncol = 6
  )
p
```

```{r}
# -----------------------------------------------------------------------------
# 1. Ordering: Brazil first, then Europe
# -----------------------------------------------------------------------------
airport_order <- df |>
  distinct(AIRPORT, REGION) |>
  arrange(REGION, AIRPORT) |>
  pull(AIRPORT)

df <- df |>
  mutate(AIRPORT = factor(AIRPORT, levels = airport_order))

gamma_curves <- gamma_fits |>
  filter(!is.na(shape), !is.na(rate)) |>
  rowwise() |>
  reframe(
    AIRPORT = AIRPORT,
    REGION  = REGION,
    x = seq(0, 10, length.out = 200),
    y = dgamma(x, shape = shape, rate = rate)
  ) |>
  mutate(AIRPORT = factor(AIRPORT, levels = airport_order))

# -----------------------------------------------------------------------------
# 2. Colours (minimal but clear)
# -----------------------------------------------------------------------------
region_cols <- c(Brazil = "#C25B2A", Europe = "#2A6696")

# -----------------------------------------------------------------------------
# 3. Plot (clean small multiples, no empty panels)
# -----------------------------------------------------------------------------
p <- ggplot() +

  # actual density
  geom_density(
    data = df,
    aes(x = ACTUAL_TXOT, fill = REGION, color = REGION),
    alpha = 0.2
  ) +

  # reference density
  geom_density(
    data = df,
    aes(x = REF_TXOT, color = REGION),
    linetype = "dashed",
    linewidth = 0.6
  ) +

  # gamma overlay (ADD_TXOT)
  geom_line(
    data = gamma_curves,
    aes(x = x, y = y, color = REGION),
    linewidth = 0.9
  ) +

  scale_color_manual(values = region_cols) +
  scale_fill_manual(values = region_cols) +

  facet_wrap(
    ~ AIRPORT,
    ncol = 6
  ) +

  labs(
    title = "Taxi-out performance distributions",
    subtitle = "Brazil (left/top) vs Europe (right/bottom) · actual vs reference vs modelled additional time",
    x = "Minutes",
    y = NULL
  )

```



```{r}
library(patchwork)

# -----------------------------------------------------------------------------
# 1. Gamma curves (shared)
# -----------------------------------------------------------------------------
gamma_curves <- gamma_fits |>
  filter(!is.na(shape), !is.na(rate)) |>
  rowwise() |>
  reframe(
    AIRPORT = AIRPORT,
    REGION  = REGION,
    x = seq(0, 10, length.out = 200),
    y = dgamma(x, shape = shape, rate = rate)
  )

# -----------------------------------------------------------------------------
# 2. Colours
# -----------------------------------------------------------------------------
region_cols <- c(Brazil = "#C25B2A", Europe = "#2A6696")

# -----------------------------------------------------------------------------
# 3. Brazil plot
# -----------------------------------------------------------------------------
p_bra <- ggplot() +
  geom_density(
    data = df |> filter(REGION == "Brazil"),
    aes(x = ACTUAL_TXOT, fill = REGION, color = REGION),
    alpha = 0.2
  ) +
  geom_density(
    data = df |> filter(REGION == "Brazil"),
    aes(x = REF_TXOT, color = REGION),
    linetype = "dashed"
  ) +
  geom_line(
    data = gamma_curves |> filter(REGION == "Brazil"),
    aes(x = x, y = y, color = REGION)
  ) +
  scale_color_manual(values = region_cols) +
  scale_fill_manual(values = region_cols) +
  facet_wrap(~ AIRPORT, ncol = 3) +
  labs(title = "Brazil") +
  theme(legend.position = "none")

# -----------------------------------------------------------------------------
# 4. Europe plot
# -----------------------------------------------------------------------------
p_eur <- ggplot() +
  geom_density(
    data = df |> filter(REGION == "Europe"),
    aes(x = ACTUAL_TXOT, fill = REGION, color = REGION),
    alpha = 0.2
  ) +
  geom_density(
    data = df |> filter(REGION == "Europe"),
    aes(x = REF_TXOT, color = REGION),
    linetype = "dashed"
  ) +
  geom_line(
    data = gamma_curves |> filter(REGION == "Europe"),
    aes(x = x, y = y, color = REGION)
  ) +
  scale_color_manual(values = region_cols) +
  scale_fill_manual(values = region_cols) +
  facet_wrap(~ AIRPORT, ncol = 3) +
  labs(title = "Europe") +
  theme(legend.position = "none")



p_bra <- p_bra +
  geom_vline(
  data = gamma_fits |> filter(REGION == "Brazil"),
  aes(xintercept = mean_add),
  linetype = "dotted",
  linewidth = 0.4
)

p_eur <- p_eur +
  geom_vline(
  data = gamma_fits |> filter(REGION == "Europe"),
  aes(xintercept = mean_add),
  linetype = "dotted",
  linewidth = 0.4
)


# -----------------------------------------------------------------------------
# 5. Combine (THIS is the key)
# -----------------------------------------------------------------------------
p_main <- (p_bra + p_eur) +
  plot_layout(widths = c(1, 1))

p_main
```

```{r}
title <- paste0(
  "<span style='font-size:9pt;'><b>Chart-Challenge Day 28 - Modelling</b></span><br>",
  "<span style='font-size:18pt;'><b>The hidden shape of taxi-out delay</b></span>"
)

subtitle <- paste0(
      "Brazil (left) and Europe (right) show similar averages, ",
      "but different distribution tails — revealing where predictability breaks down"
    )

caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Data file available in this on github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge."
)


p <- p_main + 
  
  scale_x_continuous(
      limits = c(0, quantile(df$ACTUAL_TXOT, 0.99, na.rm = TRUE))
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  
  plot_annotation(
    title    = title,
    subtitle = subtitle,
    caption  = caption
  ) &
  
  theme_minimal(base_size = 9) &
  theme(
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.title.y       = element_blank(),
    axis.title.x       = element_text(size = 12/.pt) ,
    axis.text.y        = element_blank(),
    axis.ticks         = element_blank(),
    strip.text         = element_text(face = "bold", size = rel(1)),
    plot.title = element_textbox(
        size = 14,
        lineheight = 1.2,
        margin = margin(b = 5),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = element_textbox_simple(
        size = 11,
        colour = "#2E2C28",
        lineheight = 1,
        margin = margin(b = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA,
         linetype = 1
      ),
    plot.caption = element_textbox_simple(
        size = 8.8,
        colour = "#4F4B45",
        halign = 0,
        lineheight = 1.08,
        margin = margin(t = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
    
    panel.spacing      = unit(0.6, "lines"),
    
    plot.margin = margin(10, 10, 10, 10)
  ) 
```

check visual parameters
```{r}
width <- 8
height <- 7
dpi <- 300

p + ggview::canvas(width = width, height = height, dpi = dpi)
```



## Save And Check

```{r}
#| label: save-final

outname <- "figures/day28-modelling-txxt.png"



ggsave(
  filename = outname,
  plot = p,
  # width = width,
  # height = height,
  dpi = dpi,
  bg = "white"
)
```