---
title: "Day 24 - South China Morning Post"
format: html
---

## Aim

Today's prompt points to the South China Morning Post rather than to an obvious machine-readable
dataset. For this day we therefore build a small public-facts dataset from COMAC product pages,
COMAC news releases, and one recent SCMP article on the C919 delivery slowdown.

The chart focuses on the two China-made passenger jets that are easiest to compare from those
public sources:

* `C909`, the regional jet formerly marketed as `ARJ21`
* `C919`, the larger narrowbody intended to challenge Airbus and Boeing in the trunk-route market

The final layout combines three ideas:

* the aircraft design envelope, using seating and range
* the current scale gap, using cumulative deliveries and passengers carried
* the current SCMP signal, namely the slow C919 delivery pace in the first quarter of 2026

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(ggtext)
library(patchwork)
library(scales)
```

## Read Challenge Inputs

```{r}
#| label: read-challenge-inputs

facts_raw <- readr::read_csv(
  "data/comac-c909-c919-public-facts-2024-2026.csv",
  show_col_types = FALSE
)

facts <- facts_raw |>
  mutate(
    observation_date = as.Date(observation_date),
    source_date = as.Date(source_date),
    value_num = readr::parse_number(as.character(value)),
    value_min_num = readr::parse_number(as.character(value_min)),
    value_max_num = readr::parse_number(as.character(value_max))
  )
```

## Build Plot Inputs

```{r}
#| label: build-plot-inputs

aircraft_palette <- c(
  C909 = "#138A74",
  C919 = "#C24A3A"
)

aircraft_fill <- c(
  C909 = alpha("#138A74", 0.18),
  C919 = alpha("#C24A3A", 0.18)
)

spec_data <- facts |>
  filter(fact_category == "specification") |>
  select(aircraft, fact_name, value_min_num, value_max_num) |>
  pivot_wider(
    names_from = fact_name,
    values_from = c(value_min_num, value_max_num),
    names_sep = "_"
  ) |>
  transmute(
    aircraft,
    seats_min = value_min_num_seating_capacity,
    seats_max = value_max_num_seating_capacity,
    range_min = value_min_num_range,
    range_max = value_max_num_range,
    aircraft_lab = recode(
      aircraft,
      C909 = "C909\nregional jet",
      C919 = "C919\nnarrowbody"
    )
  )

latest_ops <- facts |>
  filter(fact_name %in% c("cumulative_deliveries", "cumulative_passengers")) |>
  arrange(aircraft, fact_name, observation_date) |>
  group_by(aircraft, fact_name) |>
  slice_tail(n = 1) |>
  ungroup() |>
  mutate(
    value = value_num,
    metric = recode(
      fact_name,
      cumulative_deliveries = "Delivered aircraft",
      cumulative_passengers = "Passengers carried"
    ),
    metric = factor(metric, levels = c("Delivered aircraft", "Passengers carried"))
  )

c919_signal <- facts |>
  filter(aircraft == "C919", fact_name %in% c("q1_deliveries_2026", "cumulative_deliveries")) |>
  arrange(observation_date) |>
  group_by(fact_name) |>
  slice_tail(n = 1) |>
  ungroup() |>
  mutate(
    value = value_num,
    label = case_when(
      fact_name == "q1_deliveries_2026" ~ paste0(comma(value), " deliveries in Q1 2026"),
      fact_name == "cumulative_deliveries" ~ paste0(comma(value), " total delivered by Apr 20, 2026")
    )
  )

caption_note <- paste0(
  "Data assembled manually on April 24, 2026 from public COMAC product/news pages and SCMP reporting. ",
  "One row per published fact; source URLs are included in ",
  "data/comac-c909-c919-public-facts-2024-2026.csv."
)
```

## Design Envelope

```{r}
#| label: design-envelope
#| fig-width: 8
#| fig-height: 5

p_specs <- ggplot(spec_data) +
  geom_rect(
    aes(
      xmin = range_min,
      xmax = range_max,
      ymin = seats_min,
      ymax = seats_max,
      fill = aircraft,
      colour = aircraft
    ),
    linewidth = 0.9
  ) +
  geom_text(
    aes(
      x = (range_min + range_max) / 2,
      y = (seats_min + seats_max) / 2,
      label = aircraft_lab,
      colour = aircraft
    ),
    lineheight = 0.95,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(values = aircraft_fill, guide = "none") +
  scale_colour_manual(values = aircraft_palette, guide = "none") +
  scale_x_continuous(
    labels = label_number(big.mark = ","),
    breaks = seq(2000, 6000, 1000),
    limits = c(2000, 5800),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    breaks = seq(50, 200, 25),
    limits = c(70, 200),
    expand = expansion(mult = c(0.02, 0.04))
  ) +
  labs(
    title = "Two different jobs",
    x = "Range (km)",
    y = "Typical seating"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "#DDD7C9", linewidth = 0.4),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14, colour = "#1C1B18"),
    plot.margin = margin(6, 8, 6, 6)
  )

p_specs
```

## Current Scale Gap

```{r}
#| label: current-scale-gap
#| fig-width: 7
#| fig-height: 5

ops_labels <- latest_ops |>
  mutate(
    label = case_when(
      metric == "Delivered aircraft" ~ comma(value),
      TRUE ~ label_number(accuracy = 0.1, scale_cut = cut_short_scale())(value)
    ),
    label_x = case_when(
      metric == "Delivered aircraft" ~ value + 3,
      TRUE ~ value + 3e5
    )
  )

p_ops <- ggplot(latest_ops, aes(x = value, y = fct_rev(aircraft), colour = aircraft)) +
  geom_segment(
    aes(x = 0, xend = value, yend = fct_rev(aircraft)),
    linewidth = 1.2,
    alpha = 0.55,
    show.legend = FALSE
  ) +
  geom_point(size = 4.5, show.legend = FALSE) +
  geom_text(
    data = ops_labels,
    aes(x = label_x, label = label),
    hjust = 0,
    size = 3.5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  scale_colour_manual(values = aircraft_palette, guide = "none") +
  scale_x_continuous(
    labels = label_number(scale_cut = cut_short_scale()),
    expand = expansion(mult = c(0, 0.18))
  ) +
  facet_wrap(~ metric, scales = "free_x", ncol = 1) +
  labs(
    title = "The scale gap is still large",
    x = NULL,
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = "#E2DDD1", linewidth = 0.4),
    strip.text = element_text(face = "bold", size = 10),
    axis.text = element_text(colour = "#45433E"),
    plot.title = element_text(face = "bold", size = 14, colour = "#1C1B18"),
    plot.margin = margin(6, 24, 6, 8)
  )

p_ops
```

## SCMP Signal

```{r}
#| label: scmp-signal
#| fig-width: 7
#| fig-height: 1.8

signal_top <- c919_signal |>
  filter(fact_name == "q1_deliveries_2026") |>
  pull(label)

signal_bottom <- c919_signal |>
  filter(fact_name == "cumulative_deliveries") |>
  pull(label)

p_signal <- ggplot() +
  annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1, fill = "#F2ECE1", colour = NA) +
  annotate("text", x = 0.05, y = 0.72,
           label = "SCMP current signal",
           hjust = 0, vjust = 0.5,
           size = 4.2, fontface = "bold", colour = "#1C1B18") +
  annotate("text", x = 0.05, y = 0.43,
           label = signal_top,
           hjust = 0, vjust = 0.5,
           size = 4.7, fontface = "bold", colour = aircraft_palette[["C919"]]) +
  annotate("text", x = 0.05, y = 0.18,
           label = signal_bottom,
           hjust = 0, vjust = 0.5,
           size = 3.7, colour = "#4A4740") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(0, 8, 0, 8),
    plot.background = element_rect(fill = NA, colour = NA)
  )

p_signal
```

## Principal First Shot

```{r}
#| label: first-shot-assembly
#| fig-width: 11
#| fig-height: 7

width <- 11
height <- 7
dpi <- 300
outname <- "figures/day24-south-china-morning-post.png"

p_right <- p_signal / p_ops +
  plot_layout(heights = c(1.1, 4))

p_first_shot <- p_specs | p_right +
  plot_layout(widths = c(1.08, 0.92)) +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA)
  )

p_first_shot
```

## Beautification

The first shot already establishes the payload. The final step below only upgrades the
submission framing: a clearer challenge title, a compact explanatory subtitle, and the
usual caption note including public-source assembly and the GitHub file location.

```{r}
#| label: beautify-final
#| fig-width: 11
#| fig-height: 7

our_title <- paste0(
  "<span style='font-size:12pt;'><b>Chart-Challenge Day 24 - South China Morning Post</b></span><br>",
  "<span style='font-size:18pt;'><b>COMAC's C909 and C919 reveal two different stages of industrial maturity</b></span>"
)

our_subtitle <- paste0(
  "Using a small public-facts dataset assembled from public COMAC pages and recent SCMP reporting, ",
  "the chart compares the two aircraft across their <b>design envelope</b> and their current public ",
  "<b>operational scale</b>. The aircraft do not target the same market segment, yet the available counts ",
  "still reveal a clear gap: the <b>C909</b> has accumulated much broader operating exposure, ",
  "while the <b>C919</b> remains in an earlier and slower production ramp-up."
)

our_caption <- paste0(
  "DECEA Performance Section and EUROCONTROL PRU.",
  "<br>Prompt source: South China Morning Post. Public facts assembled manually from COMAC product pages, COMAC news releases, and SCMP reporting dated April 20, 2026.",
  "<br>Data file available in this repository: ",
  "<span style='color:#4A4945;'>data/comac-c909-c919-public-facts-2024-2026.csv</span> on github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge.",
  "<br>", caption_note
)

p_final <- p_first_shot +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 16,
        lineheight = 1.22,
        margin = margin(b = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = element_textbox_simple(
        size = 11,
        colour = "#2E2C28",
        lineheight = 1.12,
        margin = margin(b = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.caption = element_textbox_simple(
        size = 8.8,
        colour = "#4F4B45",
        halign = 0,
        lineheight = 1.08,
        margin = margin(t = 12),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.margin = margin(12, 12, 12, 12)
    )
  )

p_final
```

## Save And Check

```{r}
#| label: save-final

ggsave(
  filename = outname,
  plot = p_final,
  width = width,
  height = height,
  dpi = dpi,
  bg = "white"
)
```

```{r}
#| label: check-final
#| echo: false
#| out-width: "100%"

knitr::include_graphics(outname)
```
