---
title: "Day 30 - Global Health Data Exchange"
format: html
---

## Aim

The last prompt is **Global Health Data Exchange**. We use it as a human-factors lens:
GHDx/GBD makes sleep disorders a measurable public-health burden, and the Brazil-Europe
operation tells us when long-haul arrival pressure reaches the destination clock.

The chart deliberately keeps both directions. Brazil-to-Europe exposes the classic early
European morning arrival wave. Europe-to-Brazil shows the reciprocal flow and makes the
asymmetry visible.

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(ggridges)
library(ggtext)
library(scales)
library(patchwork)
```

## Build Inputs

```{r}
#| label: build-inputs

bra_apts <- c(
  "SBBR", "SBCF", "SBCT", "SBEG", "SBGL", "SBGR",
  "SBKP", "SBPA", "SBRF", "SBRJ", "SBSP", "SBSV"
)

eur_apts <- c(
  "EDDF", "EDDM", "EGKK", "EGLL", "EHAM", "LEBL",
  "LEMD", "LFPG", "LGAV", "LPPT", "LSZH", "LTFM"
)

airport_meta <- tribble(
  ~icao, ~city, ~tz,
  "SBBR", "Brasilia", "America/Sao_Paulo",
  "SBCF", "Belo Horizonte", "America/Sao_Paulo",
  "SBCT", "Curitiba", "America/Sao_Paulo",
  "SBEG", "Manaus", "America/Manaus",
  "SBGL", "Rio Galeao", "America/Sao_Paulo",
  "SBGR", "Sao Paulo GRU", "America/Sao_Paulo",
  "SBKP", "Campinas", "America/Sao_Paulo",
  "SBPA", "Porto Alegre", "America/Sao_Paulo",
  "SBRF", "Recife", "America/Recife",
  "SBRJ", "Rio Santos Dumont", "America/Sao_Paulo",
  "SBSP", "Sao Paulo Congonhas", "America/Sao_Paulo",
  "SBSV", "Salvador", "America/Bahia",
  "EDDF", "Frankfurt", "Europe/Berlin",
  "EDDM", "Munich", "Europe/Berlin",
  "EGKK", "London Gatwick", "Europe/London",
  "EGLL", "London Heathrow", "Europe/London",
  "EHAM", "Amsterdam", "Europe/Amsterdam",
  "LEBL", "Barcelona", "Europe/Madrid",
  "LEMD", "Madrid", "Europe/Madrid",
  "LFPG", "Paris CDG", "Europe/Paris",
  "LGAV", "Athens", "Europe/Athens",
  "LPPT", "Lisbon", "Europe/Lisbon",
  "LSZH", "Zurich", "Europe/Zurich",
  "LTFM", "Istanbul", "Europe/Istanbul"
)

parse_vra_time <- function(x) {
  dmy_hm(x, tz = "UTC", quiet = TRUE)
}

local_hour <- function(time_utc, tzone) {
  map2_dbl(
    time_utc,
    tzone,
    \(time, zone) {
      if (is.na(time) || is.na(zone)) return(NA_real_)
      local <- with_tz(time, zone)
      hour(local) + minute(local) / 60 + second(local) / 3600
    }
  )
}

vra_raw <- readr::read_delim(
  "data-src/VRA_ano2025.csv",
  delim = ";",
  show_col_types = FALSE,
  locale = locale(encoding = "UTF-8")
)

flight_arrivals <- vra_raw |>
  filter(
    (sg_origem %in% bra_apts & sg_destino %in% eur_apts) |
      (sg_origem %in% eur_apts & sg_destino %in% bra_apts)
  ) |>
  mutate(
    direction = if_else(
      sg_origem %in% bra_apts,
      "Brazil -> Europe",
      "Europe -> Brazil"
    ),
    dep_utc = coalesce(
      parse_vra_time(dt_partida_real),
      parse_vra_time(dt_partida_prevista)
    ),
    arr_utc = coalesce(
      parse_vra_time(dt_chegada_real),
      parse_vra_time(dt_chegada_prevista)
    )
  ) |>
  left_join(
    airport_meta |> select(origin = icao, origin_city = city),
    by = c("sg_origem" = "origin")
  ) |>
  left_join(
    airport_meta |> select(dest = icao, dest_city = city, dest_tz = tz),
    by = c("sg_destino" = "dest")
  ) |>
  filter(!is.na(arr_utc), !is.na(dest_tz)) |>
  mutate(
    arr_hour_local = local_hour(arr_utc, dest_tz),
    corridor = paste0(sg_origem, " -> ", sg_destino),
    corridor_city = paste0(origin_city, " -> ", dest_city)
  ) |>
  filter(!is.na(arr_hour_local))

top_n <- 9

top_corridors <- flight_arrivals |>
  count(
    direction,
    corridor,
    sg_origem,
    sg_destino,
    origin_city,
    dest_city,
    name = "flights"
  ) |>
  group_by(direction) |>
  slice_max(flights, n = top_n, with_ties = FALSE) |>
  arrange(direction, flights) |>
  mutate(
    corridor_label = paste0(
      origin_city, " (", sg_origem, ")",
      "\n",
      dest_city, " (", sg_destino, ") · ", comma(flights)
    ),
    corridor_key = paste(direction, corridor_label, sep = "___")
  ) |>
  ungroup()

ridge_data <- flight_arrivals |>
  inner_join(
    top_corridors |> select(direction, corridor, corridor_key, corridor_label, flights),
    by = c("direction", "corridor")
  ) |>
  tidyr::crossing(hour_shift = c(-24, 0, 24)) |>
  mutate(
    arr_hour_plot = arr_hour_local + hour_shift,
    direction = factor(direction, levels = c("Brazil -> Europe", "Europe -> Brazil")),
    corridor_key = factor(corridor_key, levels = top_corridors$corridor_key)
  )

direction_summary <- flight_arrivals |>
  mutate(
    in_wocl = arr_hour_local >= 2 & arr_hour_local < 6
  ) |>
  summarise(
    flights = n(),
    wocl_share = mean(in_wocl),
    peak_hour = floor(arr_hour_local[which.max(tabulate(floor(arr_hour_local) + 1, nbins = 24))]),
    .by = direction
  ) |>
  mutate(
    direction = factor(direction, levels = c("Brazil -> Europe", "Europe -> Brazil")),
    label = paste0(
      comma(flights), " flights · ",
      percent(wocl_share, accuracy = 1),
      " in WOCL"
    )
  )
```

## Basic Plot

```{r}
#| label: basic-chart
#| fig-width: 13
#| fig-height: 8

pal_bg <- "#F6F3EC"
pal_ink <- "#202124"
pal_grid <- "#D8D2C5"
pal_wocl <- "#C43B32"
pal_peak <- "#0B6E69"
pal_relief <- "#16865C"

theme_day30 <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.background = element_rect(fill = pal_bg, colour = NA),
      panel.background = element_rect(fill = pal_bg, colour = NA),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_line(colour = pal_grid, linewidth = 0.3),
      axis.ticks = element_blank(),
      axis.text.x = element_text(colour = "#5C5750", size = 8),
      axis.text.y = element_text(colour = pal_ink, size = 8, hjust = 1),
      strip.text = element_text(face = "bold", size = 11, colour = pal_ink, hjust = 0),
      strip.background = element_blank(),
      plot.title = element_markdown(face = "bold", size = 21, colour = pal_ink),
      plot.subtitle = element_markdown(size = 10.5, colour = "#514D47", lineheight = 1.15),
      plot.caption = element_text(size = 7.2, colour = "#6B665E", hjust = 0, lineheight = 1.1),
      legend.position = "bottom",
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 8),
      panel.spacing.x = unit(2.5, "lines"),
      plot.margin = margin(18, 22, 18, 22)
    )
}

p_basic <- ggplot(
  ridge_data,
  aes(
    x = arr_hour_plot,
    y = corridor_key,
    fill = after_stat(x),
    height = after_stat(density)
  )
) +
  annotate(
    "rect",
    xmin = 2, xmax = 6, ymin = -Inf, ymax = Inf,
    fill = pal_wocl, alpha = 0.13
  ) +
  geom_density_ridges_gradient(
    scale = 2.65,
    rel_min_height = 0.006,
    bandwidth = 0.48,
    colour = "white",
    linewidth = 0.28,
    alpha = 0.96
  ) +
  geom_vline(xintercept = c(2, 6), colour = pal_wocl, linewidth = 0.35, linetype = "22") +
  geom_label(
    data = direction_summary,
    aes(x = 23.7, y = Inf, label = label),
    inherit.aes = FALSE,
    hjust = 1,
    vjust = 1.15,
    label.size = 0,
    fill = "#FFF8EA",
    colour = pal_ink,
    size = 3,
    label.padding = unit(0.14, "lines")
  ) +
  facet_wrap(~ direction, ncol = 2, scales = "free_y") +
  scale_x_continuous(
    limits = c(0, 24),
    breaks = seq(0, 24, by = 3),
    labels = \(x) if_else(x == 24, "24", str_pad(x, 2, pad = "0"))
  ) +
  scale_y_discrete(labels = \(x) str_remove(x, "^[^_]+___")) +
  scale_fill_gradientn(
    colours = c("#17213A", "#305F8F", "#2E8C93", "#F0C64E", "#C76035", "#5C1B2A"),
    limits = c(0, 24),
    oob = squish,
    breaks = c(0, 6, 12, 18, 24),
    labels = c("00", "06", "12", "18", "24"),
    name = "Destination local arrival hour"
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "The red-eye exchange",
    subtitle = "Busiest 2025 Brazil-Europe study-airport corridors, by direction.<br>Arrival times are shown on the <b>destination-local clock</b>; the <span style='color:#C43B32'>02:00-06:00</span> band marks the Window of Circadian Low used in fatigue-risk work.",
    x = "Destination local time of arrival",
    y = NULL,
    caption = paste(
      "Operational data: ANAC VRA 2025, filtered to BRA-EUR study airports; actual arrival time used when available, scheduled arrival otherwise.",
      "Health-data frame: GHDx / IHME Global Burden of Disease Results Tool, where sleep disorders are available as prevalence/YLD/DALY estimates by country and year.",
      sep = "\n"
    )
  ) +
  theme_day30()

p_basic
```

## Beautify

This keeps the basic ridgeline intact, then moves the explanatory text to the plot boundary,
wraps it with `element_textbox_simple()`, and places the colour scale above the panels to leave
more space for the two clocks.

```{r}
#| label: beautify-chart
#| fig-width: 13
#| fig-height: 8.4

our_title <- paste0(
  "<span style='font-size:10pt;color:#827D73;'>DAY 30 · GLOBAL HEALTH DATA EXCHANGE</span><br>",
  "The <span style='color:#C43B32'><b>red-eye</b></span> exchange ",
  "<span style='color:#C43B32;font-size:22pt;'>◉</span>"
)

our_subtitle <- paste0(
  "Long-haul Brazil-Europe traffic is not just a network story; it <b>lands on human clocks</b>. ",
  "The Global Health Data Exchange (GHDx) makes <b>sleep disorders</b> visible as a population health burden, ",
  "while data from the Brazil-Europe comparison study show when these corridors arrive at the destination. ",
  "Both directions concentrate roughly <b>three in ten arrivals inside the Window of Circadian Low (WOCL)</b>, ",
  "but the network shape differs: on the Brazilian side Sao Paulo GRU (SBGR) dominates the top Europe links, ",
  "while the European side is more distributed, with Lisbon (LPPT) especially prominent."
)

our_caption <- paste0(
  "Operational data: Brazil-Europe comparison study, 2025 study-airport long-haul arrivals; actual arrival time used when available, scheduled arrival otherwise.",
  "<br>Health-data frame: GHDx / IHME Global Burden of Disease Results Tool, where sleep disorders are available as prevalence/YLD/DALY estimates by country and year.",
  "<br><span style='color:#827D73;'>github.com/euctrl-pru/BRA-EUR-2026-30DayChallenge</span>"
)

relief_annotation <- tibble(
  direction = factor("Brazil -> Europe", levels = c("Brazil -> Europe", "Europe -> Brazil")),
  label_x = 13.55,
  label_y = 2.55,
  label = "less red-eye\npressure"
)

theme_day30_final <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.background = element_rect(fill = pal_bg, colour = NA),
      panel.background = element_rect(fill = pal_bg, colour = NA),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_line(colour = pal_grid, linewidth = 0.32),
      axis.ticks = element_blank(),
      axis.text.x = element_text(colour = "#5C5750", size = 8),
      axis.text.y = element_text(colour = pal_ink, size = 8.2, hjust = 1, lineheight = 0.94),
      axis.title.x = element_text(size = 12, margin = margin(t = 8)),
      strip.text = element_text(face = "bold", size = 12, colour = pal_ink, hjust = 0),
      strip.background = element_blank(),
      legend.position = "top",
      legend.justification = "right",
      legend.direction = "horizontal",
      legend.title = element_text(size = 8.5, colour = "#4F4A43"),
      legend.text = element_text(size = 8, colour = "#4F4A43"),
      legend.background = element_rect(fill = pal_bg, colour = NA),
      legend.margin = margin(0, 0, 6, 0),
      panel.spacing.x = unit(2.4, "lines"),
      plot.margin = margin(4, 8, 4, 8)
    )
}

p_final_main <- ggplot(
  ridge_data,
  aes(
    x = arr_hour_plot,
    y = corridor_key,
    fill = after_stat(x),
    height = after_stat(density)
  )
) +
  annotate(
    "rect",
    xmin = 2, xmax = 6, ymin = -Inf, ymax = Inf,
    fill = pal_wocl, alpha = 0.13
  ) +
  geom_density_ridges_gradient(
    scale = 2.5,
    rel_min_height = 0.006,
    bandwidth = 0.48,
    colour = "white",
    linewidth = 0.28,
    alpha = 0.96
  ) +
  geom_vline(xintercept = c(2, 6), colour = pal_wocl, linewidth = 0.35, linetype = "22") +
  geom_label(
    data = direction_summary,
    aes(x = 23.7, y = Inf, label = label),
    inherit.aes = FALSE,
    hjust = 1,
    vjust = 1.18,
    label.size = 0,
    fill = "#FFF8EA",
    colour = pal_ink,
    size = 3,
    label.padding = unit(0.14, "lines")
  ) +
  geom_text(
    data = relief_annotation,
    aes(x = label_x, y = label_y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    lineheight = 0.9,
    colour = pal_relief,
    fontface = "bold",
    size = 3.1
  ) +
  facet_wrap(~ direction, ncol = 2, scales = "free_y") +
  scale_x_continuous(
    limits = c(0, 24),
    breaks = seq(0, 24, by = 3),
    labels = \(x) if_else(x == 24, "24", str_pad(x, 2, pad = "0"))
  ) +
  scale_y_discrete(labels = \(x) str_remove(x, "^[^_]+___")) +
  scale_fill_gradientn(
    colours = c("#17213A", "#305F8F", "#2E8C93", "#F0C64E", "#C76035", "#5C1B2A"),
    limits = c(0, 24),
    oob = squish,
    breaks = c(0, 6, 12, 18, 24),
    labels = c("00", "06", "12", "18", "24"),
    name = "Destination local arrival hour",
    guide = guide_colourbar(
      title.position = "left",
      title.vjust = 0.8,
      barwidth = unit(3.6, "cm"),
      barheight = unit(0.28, "cm"),
      ticks.colour = "white"
    )
  ) +
  coord_cartesian(clip = "off") +
  labs(
    x = "Destination local time of arrival",
    y = NULL
  ) +
  theme_day30_final()

p_final <- p_final_main +
  plot_annotation(
    title = our_title,
    subtitle = our_subtitle,
    caption = our_caption,
    theme = theme(
      plot.title = element_textbox_simple(
        size = 21,
        lineheight = 1.08,
        margin = margin(b = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.subtitle = element_textbox_simple(
        size = 10.5,
        colour = "#514D47",
        lineheight = 1.13,
        margin = margin(b = 8),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.caption = element_textbox_simple(
        size = 7.5,
        colour = "#6B665E",
        halign = 0,
        lineheight = 1.08,
        margin = margin(t = 10),
        padding = margin(0, 0, 0, 0),
        fill = NA,
        box.color = NA
      ),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.margin = margin(14, 16, 12, 16)
    )
  ) &
  theme(plot.background = element_rect(fill = pal_bg, colour = NA))

p_final
```

## Save Figures

```{r}
#| label: save-figures
#| eval: true
#| echo: false
#| message: false
#| warning: false

ggsave(
  "figures/day30-global-health-data-exchange-rough.png",
  plot = p_basic,
  width = 13,
  height = 8,
  dpi = 320,
  bg = pal_bg
)

ggsave(
  "figures/day30-global-health-data-exchange.png",
  plot = p_final,
  width = 13,
  height = 8.4,
  dpi = 320,
  bg = pal_bg
)
```
