---
title: "Day 14 - Trade Balance"
format: html
---

NOTE - HANDLE THIS FILE WITH CARE - THE CODE DOES NOT RUN WITHOUT "TOUCHING" IT.
WE NEED TO CLEAN THIS UP.


```{r}
#| label: setup
#| message: false

library(dplyr)

# helper for data prep
source(here::here("R", "rqutils-zip.R"))
source(here::here("R", "assign-region.R"))
source(here::here("R", "turnaround.R"))

# some defaults
pattern_eur <- c("E","L","BI")
pattern_bra <- c("SB","SD","SI","SJ","SN","SS", "SW")

bra_apts <- c("SBBR","SBGR","SBSP","SBKP","SBRJ","SBGL","SBCF","SBSV","SBPA","SBCT","SBRF","SBEG")
eur_apts <- c("EGLL","EGKK","EHAM","EDDF","EDDM","LSZH","LFPG","LEMD","LEBL","LPPT","LTFM","LGAV")
```


“Trade balance” of delay: who exports/absorbs delay? 

What it could show: compare departure delay generated at origin vs arrival delay observed at destination (by airport). 

Data to tap: flight-level dep/arr delay by airport; compute net ‘delay balance’. 

Discussion hook: which airports are net generators vs net absorbers; does this differ BRA vs EUR? 

Design notes: needs careful definition; may require filtering to comparable flights. 

ggplot2 angle: diverging bar/dot plot with reorder; good discussion on metric definition. 

## data preparation

run this to ingest data for today and then store the turnaournd files for the analysis.

### Brazil

The full flight data is stored / collected in totalbr - where we currently hold a working version in data-src.
Let's get this and extract the payload needed for this study.

```{}
# first call throws error - not sure why - ran it by hand.
bra <- fs::dir_ls("data-src", regexp = "totalbr") |> arrow::read_parquet()
bra
```

We emulate the turnaround data for Brazil as the data set does not allow for an aircraft by aircraft centered join of the data.    
Based on the punctuality data we assume "transfer of delay" on an hourly basis.

```{r}
source(here::here("R","turnaround-bra.R"))

tat_list_bra <- 
```



### Europe

```{r}
path_apdf <- here::here(here::here() |> dirname(),"__DATA", "APDF")

fns_apdf <- check_zip_content(path_apdf, "apdf-2025.zip") |> 
  mutate(ICAO = str_extract(Name, "[A-Z]{4}")) 
```

test for one

```{r}
df <- read_zip_content(path_apdf, "apdf-2025.zip", fns_apdf[1])

# turnaround calculation
extract_payload_eur <- function(apdf, this_pattern = pattern_eur){
  payload <- apdf |> 
    select(  FLTID = AP_C_FLTID
           , PHASE = SRC_PHASE
           , ADEP = ADEP_ICAO, ADES = ADES_ICAO
           , REG = AP_C_REG
           , BLOCK_TIME = BLOCK_TIME_UTC
           , SCHED_TIME = SCHED_TIME_UTC
           , MVT_TIME = MVT_TIME_UTC) |> 
    mutate(REGION_FLAG = paste(
        assign_region(ADEP, pattern_eur, "EUR", "INTL")
      , assign_region(ADES, pattern_eur, "EUR", "INTL")
      , sep = "-")) 
  
  return(payload)
}

```

```{r}  
# runf for one:  df <- df |> extract_payload_eur()|> match_turnarounds() |> tibble()

# loop over all entries
tat_list_eur <- fns_apdf |> select(Name, ICAO) |>
  
    pmap(\(Name, ICAO, ...){ #-----------------loop over all airports
      read_zip_content(path_apdf, "apdf-2025.zip", Name) |> 
        extract_payload_eur() |> 
        match_turnarounds() |> 
        tibble() |> 
        mutate(ICAO = ICAO, .before = 1)
      } 
      ) |> 
    set_names(fns_apdf$ICAO)
```



## getting to the plot

```{r}
# Step 1 — bind all airport turnaround tables
all_tat <- bind_rows(list(
  EDDF = df #, #tat_eddf,
  #EHAM = tat_eham,
  # ...
), .id = "AIRPORT")

all_tat <- bind_rows(tat_list_eur, .id = "AIRPORT")

# Step 2 — compute delay metrics
tat_delays <- all_tat |>
  mutate(
    ARR_DELAY     = as.numeric(difftime(ARR_BLOCK_TIME, ARR_SCHED_TIME, units = "mins")),
    DEP_DELAY     = as.numeric(difftime(DEP_BLOCK_TIME, DEP_SCHED_TIME, units = "mins")),
    DELAY_TRANSFER = DEP_DELAY - ARR_DELAY,   # +ve = amplified, -ve = absorbed
    TAT_BUFFER    = TAT_SCHED - TAT_ACTUAL    # +ve = turn faster than planned
  ) |>
  filter(
    TAT_REGION_FLAG == "EUR-EUR",             # or parameterise
    TAT_ACTUAL > 10,
    !is.na(ARR_SCHED_TIME), !is.na(DEP_SCHED_TIME)
  )

# Step 3 — airport-level summary
airport_balance <- tat_delays |>
  group_by(AIRPORT, TAT_REGION_FLAG) |>
  summarise(
    n                    = n(),
    mean_arr_delay       = mean(ARR_DELAY,      na.rm = TRUE),
    mean_dep_delay       = mean(DEP_DELAY,      na.rm = TRUE),
    mean_delay_transfer  = mean(DELAY_TRANSFER, na.rm = TRUE),
    mean_tat_buffer      = mean(TAT_BUFFER,     na.rm = TRUE),
    pct_absorbed         = mean(DELAY_TRANSFER < 0, na.rm = TRUE) * 100,
    .groups = "drop"
  ) |>
  mutate(REGION = if_else(AIRPORT %in% eur_apts, "EUR", "BRA"))
```

```{r}
# Pipeline
#   1. Bind all airport tibbles, adding ICAO column
#   2. Compute delay metrics at flight level
#   3. Summarise at airport level
#   4. Plot A — diverging dot plot  (net delay transfer per airport)
#   5. Plot B — arrival vs departure delay scatter (absorber / generator)
#   6. Plot C — TAT buffer vs delay transfer (flight level, faceted)
# =============================================================================

library(tidyverse)

# ── Palette & theme -----------------------------------------------------------
clr_eur    <- "#1F78B4"
clr_bra    <- "#E66101"
clr_mid    <- "grey70"
clr_pos    <- "#D73027"   # amplified / bad
clr_neg    <- "#1A9850"   # absorbed / good

theme_delay <- function() {
  theme_minimal(base_size = 11) +
    theme(
      plot.title       = element_text(face = "bold", size = 13),
      plot.subtitle    = element_text(colour = "grey40", size = 10),
      strip.text       = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom"
    )
}

# =============================================================================
# 0.  INPUT
# =============================================================================
# Provide a named list of turnaround tibbles.
# Names become the ICAO column.  Example:
#
#   tat_list <- list(
#     EDDF = tat_eddf,
#     EHAM = tat_eham,
#     EGLL = tat_egll,
#     ...
#     SBGR = tat_sbgr,
#     SBSP = tat_sbsp
#   )
#
# EUR airports: define the ICAO codes so the plots can colour by region.
eur_airports <- c(
  "EDDF", "EHAM", "EGLL", "LFPG", "LEMD", "LIRF",
  "LSZH", "EBBR", "LOWW", "LTFM", "EPWA", "ENGM"
)
bra_airports <- c(
  "SBGR", "SBSP", "SBRJ", "SBGL", "SBPA", "SBSV",
  "SBRF", "SBCT", "SBCF", "SBFZ", "SBBE", "SBMN"
)

# TAT_REGION_FLAG levels to include. Adjust to filter before plotting.
# Set to NULL to include all levels present in the data.
region_flags_keep <- NULL   # e.g. c("EUR-EUR") to restrict

# Minimum TAT to exclude artefacts (minutes)
tat_min <- 10

# =============================================================================
# 1.  BIND & COMPUTE FLIGHT-LEVEL METRICS
# =============================================================================

prepare_delay_data <- function(tat_list,
                               eur_airports,
                               bra_airports,
                               region_flags_keep = NULL,
                               tat_min = 10) {

  all_tat <- bind_rows(
    lapply(names(tat_list), function(icao) {
      tat_list[[icao]] |> mutate(ICAO = icao)
    })
  )

  flight_level <- all_tat |>
    filter(
      !is.na(ARR_SCHED_TIME),
      !is.na(DEP_SCHED_TIME),
      TAT_ACTUAL > tat_min
    ) |>
    mutate(
      ARR_DELAY      = as.numeric(difftime(ARR_BLOCK_TIME, ARR_SCHED_TIME, units = "mins")),
      DEP_DELAY      = as.numeric(difftime(DEP_BLOCK_TIME, DEP_SCHED_TIME, units = "mins")),
      DELAY_TRANSFER = DEP_DELAY - ARR_DELAY,   # +ve amplified, -ve absorbed
      TAT_BUFFER     = TAT_SCHED - TAT_ACTUAL,  # +ve faster than planned
      GEO_REGION     = case_when(
        ICAO %in% eur_airports ~ "EUR",
        ICAO %in% bra_airports ~ "BRA",
        TRUE                   ~ "OTHER"
      )
    )

  if (!is.null(region_flags_keep)) {
    flight_level <- flight_level |>
      filter(TAT_REGION_FLAG %in% region_flags_keep)
  }

  flight_level
}

# =============================================================================
# 2.  AIRPORT-LEVEL SUMMARY
# =============================================================================

summarise_airports <- function(flight_level) {
  flight_level |>
    group_by(ICAO, GEO_REGION, TAT_REGION_FLAG) |>
    summarise(
      n                   = n(),
      mean_arr_delay      = mean(ARR_DELAY,      na.rm = TRUE),
      mean_dep_delay      = mean(DEP_DELAY,      na.rm = TRUE),
      mean_delay_transfer = mean(DELAY_TRANSFER, na.rm = TRUE),
      mean_tat_buffer     = mean(TAT_BUFFER,     na.rm = TRUE),
      pct_absorbed        = mean(DELAY_TRANSFER < 0, na.rm = TRUE) * 100,
      .groups = "drop"
    )
}

# =============================================================================
# 3.  PLOT A — Diverging dot plot: net delay transfer per airport
# =============================================================================
# One panel per TAT_REGION_FLAG level present in the data.
# Airports sorted by mean_delay_transfer within each facet.
# Colour encodes sign: red = amplifier, green = absorber.
# Shape encodes geo region: EUR vs BRA.

plot_diverging_transfer <- function(airport_summary) {

  # Sort airports by transfer within each flag level
  plot_data <- airport_summary |>
    group_by(TAT_REGION_FLAG) |>
    mutate(ICAO = fct_reorder(ICAO, mean_delay_transfer)) |>
    ungroup()

  n_flags <- n_distinct(plot_data$TAT_REGION_FLAG)

  ggplot(plot_data,
         aes(x = mean_delay_transfer, y = ICAO, colour = mean_delay_transfer > 0)) +
    geom_vline(xintercept = 0, colour = "grey50", linewidth = 0.4, linetype = "dashed") +
    geom_segment(aes(x = 0, xend = mean_delay_transfer, yend = ICAO),
                 colour = "grey80", linewidth = 0.4) +
    geom_point(aes(shape = GEO_REGION), size = 3.5) +
    scale_colour_manual(
      values = c("TRUE" = clr_pos, "FALSE" = clr_neg),
      labels = c("TRUE" = "Amplifier (+)", "FALSE" = "Absorber (−)"),
      name   = "Role"
    ) +
    scale_shape_manual(
      values = c("EUR" = 16, "BRA" = 17, "OTHER" = 15),
      name   = "Region"
    ) +
    facet_wrap(~ TAT_REGION_FLAG, scales = "free_y", ncol = n_flags) +
    labs(
      title    = "Plot A — Net Delay Transfer by Airport",
      subtitle = "Mean (departure delay − arrival delay) per turnaround · +ve = delay amplified · −ve = delay absorbed",
      x        = "Mean delay transfer (minutes)",
      y        = NULL
    ) +
    theme_delay()
}

# =============================================================================
# 4.  PLOT B — Arrival vs departure delay scatter
# =============================================================================
# One point per airport × TAT_REGION_FLAG combination.
# Diagonal y = x: no transfer. Above = amplifier. Below = absorber.
# Sized by number of turnarounds.

plot_arr_vs_dep <- function(airport_summary) {

  # Axis range symmetric around origin for interpretability
  lim <- airport_summary |>
    summarise(across(c(mean_arr_delay, mean_dep_delay), ~ max(abs(.), na.rm = TRUE))) |>
    summarise(lim = max(mean_arr_delay, mean_dep_delay) * 1.15) |>
    pull(lim)

  ggplot(airport_summary,
         aes(x = mean_arr_delay, y = mean_dep_delay,
             colour = GEO_REGION, size = n, label = ICAO)) +
    # Diagonal reference: no transfer
    geom_abline(slope = 1, intercept = 0,
                colour = "grey50", linewidth = 0.4, linetype = "dashed") +
    # Quadrant annotations
    annotate("text", x = -Inf, y = Inf,  hjust = -0.1, vjust = 1.3,
             label = "Absorber", colour = clr_neg, size = 3, fontface = "italic") +
    annotate("text", x = Inf,  y = -Inf, hjust = 1.1,  vjust = -0.5,
             label = "Amplifier", colour = clr_pos, size = 3, fontface = "italic") +
    geom_point(alpha = 0.75) +
    ggrepel::geom_text_repel(size = 2.8, max.overlaps = 20, show.legend = FALSE) +
    scale_colour_manual(
      values = c("EUR" = clr_eur, "BRA" = clr_bra, "OTHER" = "grey50"),
      name   = "Region"
    ) +
    scale_size_continuous(name = "Turnarounds", range = c(2, 8),
                          labels = scales::comma) +
    scale_x_continuous(limits = c(-lim, lim)) +
    scale_y_continuous(limits = c(-lim, lim)) +
    coord_fixed() +
    facet_wrap(~ TAT_REGION_FLAG) +
    labs(
      title    = "Plot B — Arrival Delay vs Departure Delay",
      subtitle = "Points above the diagonal amplify delay · below absorb · diagonal = no transfer",
      x        = "Mean arrival delay (minutes)",
      y        = "Mean departure delay (minutes)"
    ) +
    theme_delay()
}

# =============================================================================
# 5.  PLOT C — TAT buffer vs delay transfer (flight level)
# =============================================================================
# Scatter at individual turnaround level.
# TAT buffer (planned slack) on x; delay transfer on y.
# A negative slope means: more buffer → more absorption. This is the mechanism.
# Smoothed trend per facet.

plot_buffer_vs_transfer <- function(flight_level,
                                    transfer_clip = 120,
                                    buffer_clip   = 60) {

  plot_data <- flight_level |>
    filter(
      between(DELAY_TRANSFER, -transfer_clip, transfer_clip),
      between(TAT_BUFFER,     -buffer_clip,   buffer_clip)
    )

  ggplot(plot_data,
         aes(x = TAT_BUFFER, y = DELAY_TRANSFER, colour = GEO_REGION)) +
    geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.4, linetype = "dashed") +
    geom_vline(xintercept = 0, colour = "grey50", linewidth = 0.4, linetype = "dashed") +
    geom_point(alpha = 0.05, size = 0.6, show.legend = FALSE) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 0.9,
                aes(fill = GEO_REGION), alpha = 0.15) +
    scale_colour_manual(
      values = c("EUR" = clr_eur, "BRA" = clr_bra, "OTHER" = "white" ), #, "OTHER" = "grey50"),
      name   = "Region"
    ) +
    scale_fill_manual(
      values = c("EUR" = clr_eur, "BRA" = clr_bra, "OTHER" = "grey50"),
      guide  = "none"
    ) +
    facet_wrap(~ TAT_REGION_FLAG) +
    labs(
      title    = "Plot C — TAT Buffer vs Delay Transfer",
      subtitle = paste0(
        "Flight level · clipped to ±", transfer_clip, " min transfer / ±", buffer_clip, " min buffer\n",
        "Negative slope: more planned slack → more delay absorbed"
      ),
      x = "TAT buffer  [scheduled − actual TAT, minutes]",
      y = "Delay transfer  [departure − arrival delay, minutes]"
    ) +
    theme_delay()
}

# =============================================================================
# 6.  RUNNER — call after building tat_list
# =============================================================================

run_delay_analysis <- function(tat_list,
                               eur_airports      = character(),
                               bra_airports      = character(),
                               region_flags_keep = NULL,
                               tat_min           = 10,
                               transfer_clip     = 120,
                               buffer_clip       = 60,
                               save_plots        = FALSE,
                               output_dir        = ".") {

  message("Preparing flight-level data ...")
  fl <- prepare_delay_data(tat_list, eur_airports, bra_airports,
                           region_flags_keep, tat_min)
  message(sprintf("  %s turnarounds across %s airports",
                  scales::comma(nrow(fl)), n_distinct(fl$ICAO)))

  message("Summarising at airport level ...")
  ap <- summarise_airports(fl)

  message("Building plots ...")
  pA <- plot_diverging_transfer(ap)
  pB <- plot_arr_vs_dep(ap)
  pC <- plot_buffer_vs_transfer(fl, transfer_clip, buffer_clip)

  if (save_plots) {
    ggsave(file.path(output_dir, "plotA_diverging_transfer.png"), pA,
           width = 12, height = 7, dpi = 150)
    ggsave(file.path(output_dir, "plotB_arr_vs_dep.png"),         pB,
           width = 10, height = 7, dpi = 150)
    ggsave(file.path(output_dir, "plotC_buffer_vs_transfer.png"), pC,
           width = 10, height = 7, dpi = 150)
    message("  Plots saved to ", output_dir)
  }

  list(
    flight_level     = fl,
    airport_summary  = ap,
    plot_diverging   = pA,
    plot_arr_vs_dep  = pB,
    plot_buffer      = pC
  )
}

# =============================================================================
# EXAMPLE USAGE
# =============================================================================
if (FALSE) {

  # Assumes match_turnarounds() has been run per airport:
  tat_list <- list(
    EDDF = tat_eddf,
    EHAM = tat_eham,
    EGLL = tat_egll,
    LFPG = tat_lfpg,
    LEMD = tat_lemd,
    LIRF = tat_lirf,
    LSZH = tat_lszh,
    EBBR = tat_ebbr,
    LOWW = tat_loww,
    LTFM = tat_ltfm,
    EPWA = tat_epwa,
    ENGM = tat_engm,
    SBGR = tat_sbgr,
    SBSP = tat_sbsp,
    SBRJ = tat_sbrj,
    SBGL = tat_sbgl,
    SBPA = tat_sbpa,
    SBSV = tat_sbsv,
    SBRF = tat_sbrf,
    SBCT = tat_sbct,
    SBCF = tat_sbcf,
    SBFZ = tat_sbfz,
    SBBE = tat_sbbe,
    SBMN = tat_sbmn
  )

  results <- run_delay_analysis(
    tat_list          = tat_list,
    eur_airports      = eur_airports,
    bra_airports      = bra_airports,
    region_flags_keep = NULL,        # NULL = all flags; or e.g. c("EUR-EUR")
    tat_min           = 10,
    transfer_clip     = 120,
    buffer_clip       = 60,
    save_plots        = TRUE,
    output_dir        = "outputs"
  )

  # Access individual outputs
  results$plot_diverging
  results$plot_arr_vs_dep
  results$plot_buffer

  # Further exploration: worst amplifiers
  results$airport_summary |>
    arrange(desc(mean_delay_transfer)) |>
    select(ICAO, GEO_REGION, TAT_REGION_FLAG, n, mean_arr_delay,
           mean_dep_delay, mean_delay_transfer, pct_absorbed)
}
```


final plot

```{r}
fb <- tat_list_eur |> run_delay_analysis()
```

```{r}
p_trade <- fb$plot_diverging
```

```{r}
library("ggtext")

our_title <- "<span style='font-size:12pt'><strong>Chart-Challenge - Day 14 - Trade</strong></span><br><strong>Are European study airports delay absorbers?</strong>"

our_subtitle <- "The following chart analyses the **delay transfer** for arrivals and successive departures of aircraft at European airports in 2025. We observe a general amplification of delay suggesting that arrival delays result in longer departure delays. Notable exception is Athens airport (LGAV) where a net absorption for aircraft operating European routes with an successive international leg take place (EUR-INTL). **On average the mean delay transfer accounts for less than 15 minutes.**"

our_caption <- "DECEA Performance Section & EUROCONTROL PRU"
```


```{r}
p_trade <- p_trade +
  labs(title = our_title, subtitle = our_subtitle, caption = our_caption) +
  
  theme(
    legend.position = "none"
    #------------------- title, subtitle, caption ----------------
    ,plot.title.position = "plot"
    ,plot.title = element_textbox_simple(
      margin = margin(b = 10)
    )
    
    ,plot.subtitle.position = "plot"
    ,plot.subtitle = element_textbox_simple(
      margin = margin(b = 10)
    )
    
    ,plot.caption.position = "plot"
    ,plot.caption = element_textbox_simple(
      margin = margin(t = 10)
    )
    
    ,axis.title.x = element_text(size = rel(0.8))
  ) 

p_trade +
  ggview::canvas(width = 8, height = 6, dpi = 300)
```

ggsave("./figures/day14-trade.png", width = 8, height = 6, dpi = 300)
